---
title: "MiSeq-239"
output:
  html_notebook:
    df_print: paged
    number_sections: yes
    theme: lumen
    toc: yes
    code_folding: hide
---

<style>
  html {
    font-size: 16pt;
  }
  
  body {
    font-size: 16pt;
  }
  
  h1 {
    font-size: 2.2rem;
  }
  
  h2 {
    font-size: 2rem;
  }
  
  h3 {
    font-size: 1.8rem;
  }
  
  h4 {
    font-size: 1.4rem;
  }
  
</style>

# FB Analysis -- Plots {.tabset}

## Setup
### Start Conda ENV
```{r}
startCondaEnv('alphadiv')
```

### Set Knitr Options
```{r}
library(rmarkdown)
library(knitr)

opts_chunk$set(
  echo=TRUE,
  dpi=300,
  fig.width=12
  )
```

### Load Libraries
```{r}
library(tidyverse)
library(magrittr)
library(ggplot2)
library(ggbeeswarm)
library(openxlsx)
library(DT)
library(pheatmap)
library(kableExtra)
library(vegan)
library(nlme)

### Custom libraries that can be loaded from GitHub
source('~/utilities/analysis-utilities/general_asv_data_analysis_utilities.R')
source('~/utilities/analysis-utilities/alpha_div_utilities.R')
```


### Set Paths
```{r}
source('./set_paths')
```

### Load Data
```{r}
all_master_table = read.xlsx(file.path(tables_dir, 'miseq-239_F_B_master_table.xlsx'))
age_lt_90_master_table = all_master_table %>% filter(Age<90)
amd_only_master_table = all_master_table %>% filter(CaseString=='AMD')

source('./observational_variables.R')

filename = file.path(tables_dir, 'miseq-239_fb_mann_whitney_unadjusted.xlsx')
pair_stats = read.xlsx(filename)
```

### Variables of interest
#### All Observational Variables
```{r}
print(observational_variables)
```

#### AMD Only Observational Variables
```{r}
amd_only_variables = observational_variables

amd_only_excluded_variables = c(
  'Gender',
  'CaseString',
  'Age'
)

for (varname in amd_only_excluded_variables)
{
  print(sprintf('Excluding %s', varname))
  amd_only_variables[varname] = NULL
}

```

**Excluding: ** `r amd_only_excluded_variables`

#### Raw Experimental Variables
```{r}
raw_exp_vars = c('Firmicutes', 'Bacteroidetes')
```

**Raw Experimental Variables: ** `r raw_exp_vars`


#### Calculated Experimental Variables

```{r}
calculated_exp_vars = c('F2B_Ratio', 'LogF2B_Ratio', 'FB_Index')
```
**Calculated Experimental Variables: ** `r calculated_exp_vars`

#### Subgroups
```{r}
subgroups = list(
  All=list(name='All Subgroups', prefix='all'),
  Age_lt_90=list(name='Age < 90', prefix='age_lt_90'),
  Amd_Only=list(name='AMD Only', prefix='amd_only')
)
```


### Set Metadata Factor Levels
```{r}

setFactors = function(df)
{
  for (var in observational_variables)
  {
    name = var$covariate_of_interest
    case = var$case
    control = var$control
    # print(sprintf('name: %s, case: %s, control: %s', name, case, control))
  
    if (case != '' & control != '')
    {
      # print('ok')
      df = mutate(df, !!name := factor(!!as.name(name), levels=c(control, case)))
    }
  
  }
  
  return(df)
}

all_master_table = setFactors(all_master_table)
age_lt_90_master_table = setFactors(age_lt_90_master_table)
amd_only_master_table = setFactors(amd_only_master_table)
```

### P-Value Annotationas
```{r}
pval_annotations = getPvalAnnotations(
  variables=c(raw_exp_vars, calculated_exp_vars),
  master_table=all_master_table,
  pval_height_factor=1.2,
  pair_stats=pair_stats
)
```


## Plots
### All Subgroups
```{r}
for (var in raw_exp_vars)
{
  makeAlphaDivPlot(
  all_master_table,
  indices=raw_exp_vars,
  aesthetics = list(x_var='index', color_var='index', facet_var='index')
) + 
  ggtitle("Firmicutes and Bacteroidetes") +
  ylab("Value") + 
  xlab("Abundance")
}
```

#### Abundance
```{r}
makeAlphaDivPlot(
  all_master_table,
  indices=raw_exp_vars,
  aesthetics = list(x_var='index', color_var='index', facet_var='index')
) + 
  ggtitle("Firmicutes and Bacteroidetes") +
  ylab("Value") + 
  xlab("Abundance")
```

#### Indices
```{r}
makeAlphaDivPlot(
  all_master_table,
  indices=calculated_exp_vars,
  aesthetics = list(x_var='index', color_var='index', facet_var='index')
) + 
  ggtitle("FB Indices") +
  ylab("Value") + 
  xlab("Index")
```

#### AMD vs. Control
```{r}
makePairPlot(
  master_table=all_master_table,
  indices=raw_exp_vars,
  var_data=observational_variables$CaseString,
  aesthetics=list(
    x_var='CaseString', 
    color_var='CaseString', 
    facet_var='index'
    ),
  annotation_data=pval_annotations %>% filter(TestGroup=='All'),
  title_template="FB Abundance: "
)

makePairPlot(
  master_table=all_master_table,
  indices=calculated_exp_vars,
  var_data=observational_variables$CaseString,
  aesthetics=list(
    x_var='CaseString', 
    color_var='CaseString', 
    facet_var='index'
    ),
  annotation_data=pval_annotations %>% filter(TestGroup=='All'),
  title_template="FB Index: "
)


```

### Age < 90
#### Abundance
```{r}
makeAlphaDivPlot(
  age_lt_90_master_table,
  indices=raw_exp_vars,
  aesthetics = list(x_var='index', color_var='index', facet_var='index')
) + 
  ggtitle("Firmicutes and Bacteroidetes") +
  ylab("Value") + 
  xlab("Abundance")
```

#### Indices
```{r}
makeAlphaDivPlot(
  age_lt_90_master_table,
  indices=calculated_exp_vars,
  aesthetics = list(x_var='index', color_var='index', facet_var='index')
) + 
  ggtitle("FB Indices") +
  ylab("Value") + 
  xlab("Index")
```

#### AMD vs. Control
```{r}
makePairPlot(
  master_table=age_lt_90_master_table,
  indices=raw_exp_vars,
  var_data=observational_variables$CaseString,
  aesthetics = list(x_var='CaseString', color_var='CaseString', facet_var='index'),
  annotation_data=pval_annotations %>% filter(TestGroup=='All'),
  title_template="FB Abundance: "
)

makePairPlot(
  master_table=age_lt_90_master_table,
  indices=calculated_exp_vars,
  var_data=observational_variables$CaseString,
  aesthetics = list(x_var='CaseString', color_var='CaseString', facet_var='index'),
  annotation_data=pval_annotations %>% filter(TestGroup=='All'),
  title_template="FB Index: "
)

```


### AMD Only

#### Abundance
```{r}
makeAlphaDivPlot(
  age_lt_90_master_table,
  indices=raw_exp_vars,
  aesthetics = list(x_var='index', color_var='index', facet_var='index')
) + 
  ggtitle("Firmicutes and Bacteroidetes") +
  ylab("Value") + 
  xlab("Abundance")
```

#### Indices
```{r}
makeAlphaDivPlot(
  age_lt_90_master_table,
  indices=calculated_exp_vars,
  aesthetics = list(x_var='index', color_var='index', facet_var='index')
) + 
  ggtitle("FB Indices") +
  ylab("Value") + 
  xlab("Index")
```

#### Covariates

##### Abundance
```{r}
plotPairTests(
  variables=amd_only_variables,
  master_table=all_master_table,
  indices=raw_exp_vars,
  title_template="FB Abundance: ",
  annotation_data=pval_annotations %>% filter(TestGroup=='AMD_Only')
)

```

##### Index
```{r}
plotPairTests(
  variables=amd_only_variables,
  master_table=all_master_table,
  indices=calculated_exp_vars,
  title_template="FB Abundance: ",
  annotation_data=pval_annotations %>% filter(TestGroup=='AMD_Only')
)

```
