---
title: "MiSeq-239"
output:
  html_notebook:
    df_print: paged
    number_sections: yes
    theme: lumen
    toc: yes
    code_folding: hide
---

<style>
  html {
    font-size: 16pt;
  }
  
  body {
    font-size: 16pt;
  }
  
  h1 {
    font-size: 2.2rem;
  }
  
  h2 {
    font-size: 2rem;
  }
  
  h3 {
    font-size: 1.8rem;
  }
  
  h4 {
    font-size: 1.4rem;
  }
  
</style>

# F B Analysis; Pair Tests {.tabset}
## Clustering Level
Data clustered at the Genus level for taxa abundances.

## Setup

### Start Conda ENV
```{r}
startCondaEnv('alphadiv')
```

### Set Knitr Options
```{r}
library(rmarkdown)
library(knitr)

opts_chunk$set(
  echo=TRUE,
  dpi=300,
  fig.width=12
  )
```

### Load Libraries
```{r}
library(tidyverse)
library(magrittr)
library(ggplot2)
library(ggbeeswarm)
library(openxlsx)
library(DT)
library(pheatmap)
library(kableExtra)
library(vegan)
library(nlme)

### Custom libraries that can be loaded from GitHub
source('~/utilities/analysis-utilities/general_asv_data_analysis_utilities.R')
source('~/utilities/analysis-utilities/alpha_div_utilities.R')
```
### Set Paths
```{r}
source('./set_paths.R')
```

### Load Data
```{r}
all_master_table = read.xlsx(file.path(tables_dir, 'miseq-239_F_B_master_table.xlsx'))

source(file.path(metadata_dir, 'observational_variables.R'))

age_lt_90_master_table = all_master_table %>% filter(Age<90)
amd_only_master_table = all_master_table %>% filter(CaseString=='AMD')
```


### Variables of interest
#### All Observational Variables
```{r}
# print(observational_variables)
```

#### AMD Only Observational Variables
```{r}
amd_only_variables = observational_variables

amd_only_excluded_variables = c(
  'Gender',
  'CaseString',
  'Age'
)

for (varname in amd_only_excluded_variables)
{
  print(sprintf('Excluding %s', varname))
  amd_only_variables[varname] = NULL
}

```

**Excluding: ** `r amd_only_excluded_variables`

#### Raw Experimental Variables
```{r}
raw_exp_vars = c('Firmicutes', 'Bacteroidetes')
```

**Raw Experimental Variables: ** `r raw_exp_vars`


#### Calculated Experimental Variables

```{r}
calculated_exp_vars = c('F2B_Ratio', 'LogF2B_Ratio', 'FB_Index')
```
**Calculated Experimental Variables: ** `r calculated_exp_vars`

#### Subgroups
```{r}
subgroups = list(
  All=list(name='All Subgroups', prefix='all'),
  Age_lt_90=list(name='Age < 90', prefix='age_lt_90'),
  Amd_Only=list(name='AMD Only', prefix='amd_only')
)
```


### Set Metadata Factor Levels
```{r}
all_master_table = setFactorsLevels(
  all_master_table, 
  observational_variables
  )
age_lt_90_master_table = setFactorsLevels(
  age_lt_90_master_table, 
  observational_variables
  )
amd_only_master_table = setFactorsLevels(
  amd_only_master_table, 
  observational_variables
  )
```


## Pair Tests

#### Initialize Table

```{r}
contrast_names = makeContrastNames(observational_variables)
```

```{r}
all_pair_stats = data.frame(
  TestGroup=character(),
  TestVariable=character(),
  Reference=character(),
  Comparison=character(),
  Firmicutes=numeric(),
  Bacteroidetes=numeric(),
  F2B_Ratio=numeric(),
  LogF2B_Ratio=numeric(),
  FB_Index=numeric()
  )

```


#### All
```{r}
pval_list = doPairTests(
  exp_vars = c(raw_exp_vars, calculated_exp_vars),
  test_var = observational_variables$CaseString,
  master_table = all_master_table,
  test_group = 'All'
)

all_pair_stats %<>%
  add_row(!!!pval_list)
```

#### Age < 90
```{r}
pval_list = doPairTests(
  exp_vars = c(raw_exp_vars, calculated_exp_vars),
  test_var = observational_variables$CaseString,
  master_table = age_lt_90_master_table,
  test_group = 'Age_lt_90'
)

all_pair_stats %<>%
  add_row(!!!pval_list)

```
``


```{r}
amd_only_pair_stats = data.frame(
  TestGroup=character(),
  TestVariable=character(),
  Reference=character(),
  Comparison=character(),
  Firmicutes=numeric(),
  Bacteroidetes=numeric(),
  F2B_Ratio=numeric(),
  LogF2B_Ratio=numeric(),
  FB_Index=numeric()
  )

exp_vars = c(raw_exp_vars, calculated_exp_vars)
# test_var = amd_only_variables$AREDS
master_table = amd_only_master_table
test_group = 'AMD_Only'
  
for (test_var in amd_only_variables)
{
  
  name = test_var$covariate_of_interest
  print(name)
  
  pval_list = doPairTests(
    exp_vars = c(raw_exp_vars, calculated_exp_vars),
    test_var = test_var,
    master_table = amd_only_master_table,
    test_group = 'AMD_Only'
  )
  
  amd_only_pair_stats %<>%
    add_row(!!!pval_list)
}
```

```{r}
all_pair_stats =
  all_pair_stats %>%
  add_row(!!!amd_only_pair_stats)
```



```{r}
filename = file.path(tables_dir, 'miseq-239_fb_mann_whitney_unadjusted.xlsx')
write.xlsx(
  all_pair_stats,
  file=filename,
  quote=F
)
```

