---
title: "MiSeq-239"
output:
  html_notebook:
    df_print: paged
    number_sections: yes
    theme: lumen
    toc: yes
    code_folding: hide
---

<style>
  html {
    font-size: 16pt;
  }
  
  body {
    font-size: 16pt;
  }
  
  h1 {
    font-size: 2.2rem;
  }
  
  h2 {
    font-size: 2rem;
  }
  
  h3 {
    font-size: 1.8rem;
  }
  
  h4 {
    font-size: 1.4rem;
  }
  
</style>

# F B Analysis: Data Prep {.tabset}
## Clustering Level
Data clustered at the Genus level for taxa abundances.

## Setup 

### [Close]

### Start Conda ENV
```{r}
startCondaEnv('alphadiv')
```

### Set Knitr Options
```{r}
library(rmarkdown)
library(knitr)

opts_chunk$set(
  echo=TRUE,
  dpi=300,
  fig.width=12
  )
```

### Load Libraries
```{r}
library(tidyverse)
library(magrittr)
library(ggplot2)
library(ggbeeswarm)
library(openxlsx)
library(DT)
library(pheatmap)
library(kableExtra)
library(vegan)
library(nlme)

### Custom libraries that can be loaded from GitHub
source('~/utilities/analysis-utilities/general_asv_data_analysis_utilities.R')
source('~/utilities/analysis-utilities/alpha_div_utilities.R')
```

### Set Paths
```{r}
source('./set_paths.R')
```

### Load Data
```{r}
asv_table = 
  read.delim(
    file.path(dada2_table_dir, 'miseq-239_tall_asv_table.tsv'), 
    header=T, 
    sep='\t',
    stringsAsFactors=F
    )

taxonomy_table = 
  read.delim(
    file.path(dada2_table_dir, 'miseq-239_taxonomy_table.tsv'), 
    header=T, 
    sep='\t',
    stringsAsFactors=F
    ) %>%
  mutate_all(as.character())

sample_metadata = 
  read.delim(
    file.path(project_dir, 'metadata', 'miseq-239_metadata_CNV_GA.tsv'), 
    header=T, 
    sep='\t',
    stringsAsFactors=F
    ) %>%
  filter(SampleID!='Undetermined')

metadata_columns = colnames(sample_metadata)
sample_names = sample_metadata$SampleName

asv_table_cols = 
  data.frame(
    SampleID=colnames(asv_table)
  ) %>%
  inner_join(sample_metadata, by='SampleID') %>%
  pull(SampleName)

colnames(asv_table) = c('ASVs', asv_table_cols)

pval_names = list(
  CaseString='CaseStringAMD',
  Gender='GenderM',
  AREDS='AREDSY',
  Age='Age'
)
```

### Get Filtered Taxa Abundance
```{r}
genus_taxa_abundance = getFilteredTaxaCounts(
  asv_table,
  taxonomy_table,
  sample_metadata, 
  lowest_rank='Genus',
  relative_abundance_cutoff=0.002,
  prevalence_cutoff=0,
  id_col="SampleName",
  normalize=T
  )

phylum_taxa_abundance = getFilteredTaxaCounts(
  asv_table,
  taxonomy_table,
  sample_metadata, 
  lowest_rank='Phylum',
  relative_abundance_cutoff=0.002,
  prevalence_cutoff=0,
  id_col="SampleName",
  normalize=T
  )

filtered_taxa = genus_taxa_abundance$short_glommed_taxa

fb_taxa_abundance = 
  genus_taxa_abundance %>%
  select(Phylum, !!sample_names) %>%
  filter(Phylum %in% c('Bacteroidetes', 'Firmicutes')) %>%
  group_by(Phylum) %>%
  summarise_all(sum)
```

# Firmicute and Bacteroidetes Table

## Create FB Stats Table

Final table has columns for

* Firmicutes abundance
* Bacteroidets abundance
* F to B ratio
  $$ f2br = \frac{[Firmicutes]}{[Bacteroidetes]} $$
* Log F to B ratio $logf2br$
* FB index
  $$ fbidx = \frac{[Firmicutes] - [Bacteroidetes]}{[Firmicutes] + [Bacteroidetes]}$$
* All metadata columns (`SampleName`, `CaseString`, `AREDS`, `Gender`, etc.)

Each row represents a different sample.

The three measures are different ways of characterize the relationship between Firmicute and Bacteroidetes abundance. The $fbidx$ varies from -1 (only Firmicutes present) to 1 (only Bacteroidetes present) for that sample.

```{r}
fb_master_table = getMasterTable(
  fb_taxa_abundance,
  sample_metadata,
  'SampleName',
  'Phylum'
)

genus_master_table = getMasterTable(
  genus_taxa_abundance,
  sample_metadata,
  'SampleName',
  'short_glommed_taxa'
)

master_table = 
  fb_master_table %>% 
  select(SampleName, Firmicutes, Bacteroidetes) %>%
  inner_join(genus_master_table, by='SampleName')

choice_metadata_columns = setdiff(metadata_columns, 'SampleName')

fb_master_table = 
  master_table %>%
  mutate(
    F2B_Ratio = Firmicutes/Bacteroidetes,
    LogF2B_Ratio = log(F2B_Ratio),
    FB_Index = (Firmicutes - Bacteroidetes)/(Firmicutes + Bacteroidetes),
    FxB = Firmicutes*Bacteroidetes
  ) %>%
  select(
    SampleName, 
    Firmicutes, 
    Bacteroidetes, 
    F2B_Ratio, 
    LogF2B_Ratio, 
    FB_Index, 
    metadata_columns,
    everything()
    )

```

## Write Table
```{r}
filename = file.path(tables_dir, 'miseq-239_F_B_master_table.xlsx')

write.xlsx(
  fb_master_table,
  file=filename,
  quote=F
)
```

