---
title: "MiSeq-239"
output:
  html_notebook:
    df_print: paged
    number_sections: yes
    theme: lumen
    toc: yes
    code_folding: hide
---

<style>
  html {
    font-size: 16pt;
  }
  
  body {
    font-size: 16pt;
  }
  
  h1 {
    font-size: 2.2rem;
  }
  
  h2 {
    font-size: 2rem;
  }
  
  h3 {
    font-size: 1.8rem;
  }
  
  h4 {
    font-size: 1.4rem;
  }
  
</style>

# F B Analysis: Regression {.tabset}

## Clustering Level
Data clustered at the Genus level for taxa abundances.

## Setup {.tabset}

### [Close]

### Start Conda ENV
```{r}
startCondaEnv('alphadiv')
```

### Set Knitr Options
```{r}
library(rmarkdown)
library(knitr)

opts_chunk$set(
  echo=TRUE,
  dpi=300,
  fig.width=12
  )
```

### Load Libraries
```{r}
library(tidyverse)
library(magrittr)
library(ggplot2)
library(ggbeeswarm)
library(openxlsx)
library(DT)
library(pheatmap)
library(kableExtra)
library(vegan)
library(nlme)

### Custom libraries that can be loaded from GitHub
source('~/utilities/analysis-utilities/general_asv_data_analysis_utilities.R')
source('~/utilities/analysis-utilities/alpha_div_utilities.R')
```


### Set Paths
```{r}
source('./set_paths.R')
```

### Load Data
```{r}
all_master_table = read.xlsx(file.path(tables_dir, 'miseq-239_F_B_master_table.xlsx'))
age_lt_90_master_table = all_master_table %>% filter(Age<90)
amd_only_master_table = all_master_table %>% filter(CaseString=='AMD')

source(file.path(metadata_dir, 'observational_variables.R'))

sample_metadata = read.delim(
  file.path(metadata_dir, 'miseq-239_metadata_CNV_GA.tsv'), 
  stringsAsFactors=F
  )
```

### Variables of interest
#### All Observational Variables
```{r}
# print(observational_variables)
```

#### AMD Only Observational Variables
```{r}
amd_only_variables = observational_variables

amd_only_excluded_variables = c(
  'Gender',
  'CaseString',
  'Age'
)

for (varname in amd_only_excluded_variables)
{
  print(sprintf('Excluding %s', varname))
  amd_only_variables[varname] = NULL
}

```

**Excluding: ** `r amd_only_excluded_variables`

#### Raw Experimental Variables
```{r}
raw_exp_vars = c('Firmicutes', 'Bacteroidetes')
```

**Raw Experimental Variables: ** `r raw_exp_vars`


#### Calculated Experimental Variables

```{r}
calculated_exp_vars = c('F2B_Ratio', 'LogF2B_Ratio', 'FB_Index')
```
**Calculated Experimental Variables: ** `r calculated_exp_vars`

#### Subgroups
```{r}
subgroups = list(
  All=list(name='All Subgroups', prefix='all'),
  Age_lt_90=list(name='Age < 90', prefix='age_lt_90'),
  Amd_Only=list(name='AMD Only', prefix='amd_only')
)
```


### Set Metadata Factor Levels
```{r}
all_master_table = setFactorsLevels(all_master_table, observational_variables)
age_lt_90_master_table = setFactorsLevels(age_lt_90_master_table, observational_variables)
amd_only_master_table = setFactorsLevels(amd_only_master_table, observational_variables)
```


### Determine which covariates have sufficient samples with each value
```{r}
num_samples_threshold = 5

all_variables = filterVarsBySampleThreshold(
  all_master_table, 
  num_samples_threshold,
  observational_variables
  )

age_lt_90_variables = filterVarsBySampleThreshold(
  age_lt_90_master_table, 
  num_samples_threshold,
  observational_variables
  )

amd_only_variables = filterVarsBySampleThreshold(
  amd_only_master_table %>% select(-CaseString), 
  num_samples_threshold,
  observational_variables
  )


all_formula_rhs = paste0(names(all_variables), collapse=' + ')
age_lt_90_formula_rhs = paste0(names(age_lt_90_variables), collapse=' + ')
amd_only_formula_rhs = paste0(names(amd_only_variables), collapse=' + ')
```


### Create p-val names for contrasts

```{r}
contrast_names = makeContrastNames(all_variables)
```

### Initialize regression stats table

```{r}
regression_pvals = makeRegressionStatsTemplate(all_variables)
regression_effect_sizes = makeRegressionStatsTemplate(all_variables)
```

## Linear Regression {.tabset}

### All

```{r}
exp_vars = c(raw_exp_vars, calculated_exp_vars)
predictors = all_variables
master_table = all_master_table
index_name_template = 'All_'

# regression_pvals = regression_pvals_template
# regression_effect_sizes = regression_pvals_template

for (var in exp_vars)
{
  index_name = paste0(index_name_template, var)
  print(index_name)
  
  fit = doMultipleRankRegression(
    var,
    predictors,
    master_table
  )
  
  pvals = getRegressionPvals(
    fit,
    predictors,
    contrast_names,
    index_name
  )
  
  regression_pvals = regression_pvals %>% add_row(!!!pvals)

  effect_sizes = getRegressionEffectSizes(
    fit,
    predictors,
    contrast_names,
    index_name
  )

  regression_effect_sizes = regression_effect_sizes %>% add_row(!!!effect_sizes)

}
```

### Age < 90

```{r}
exp_vars = c(raw_exp_vars, calculated_exp_vars)
predictors = age_lt_90_variables
master_table = age_lt_90_master_table
index_name_template = 'Age_lt_90_'

# regression_pvals = regression_pvals_template
# regression_effect_sizes = regression_pvals_template

for (var in exp_vars)
{
  index_name = paste0(index_name_template, var)
  print(index_name)
  
  fit = doMultipleRankRegression(
    var,
    predictors,
    master_table
  )
  
  pvals = getRegressionPvals(
    fit,
    predictors,
    contrast_names,
    index_name
  )
  
  regression_pvals = regression_pvals %>% add_row(!!!pvals)

  effect_sizes = getRegressionEffectSizes(
    fit,
    predictors,
    contrast_names,
    index_name
  )

  regression_effect_sizes = regression_effect_sizes %>% add_row(!!!effect_sizes)

}
```


### AMD ONly

```{r}
exp_vars = c(raw_exp_vars, calculated_exp_vars)
predictors = amd_only_variables
master_table = amd_only_master_table
index_name_template = 'AMD_Only_'

# regression_pvals = regression_pvals_template
# regression_effect_sizes = regression_pvals_template

for (var in exp_vars)
{
  index_name = paste0(index_name_template, var)
  print(index_name)
  
  fit = doMultipleRankRegression(
    var,
    predictors,
    master_table
  )
  
  pvals = getRegressionPvals(
    fit,
    predictors,
    contrast_names,
    index_name
  )
  
  regression_pvals = regression_pvals %>% add_row(!!!pvals)

  effect_sizes = getRegressionEffectSizes(
    fit,
    predictors,
    contrast_names,
    index_name
  )

  regression_effect_sizes = regression_effect_sizes %>% add_row(!!!effect_sizes)

}
```


### Write Tables
```{r}
regression_stats = createWorkbook()

addWorksheet(
  regression_stats, 
  sheetName="Unadjusted P-Values"
  )

openxlsx::writeData(
  regression_stats, 
  sheet="Unadjusted P-Values", 
  regression_pvals
  )

addWorksheet(
  regression_stats, 
  sheetName="Effect Sizes"
  )

openxlsx::writeData(
  regression_stats,
  sheet="Effect Sizes", 
  regression_effect_sizes
  )

# addWorksheet(regression_pvals, sheetName="Adjusted P-Values")
# openxlsx::writeData(regression_pvals, sheet="Adjusted P-Values", adjusted_regression_pvals)

filename = file.path(tables_dir, 'miseq-239_FB_rank_regression_stats.xlsx')

saveWorkbook(regression_stats, file=filename, overwrite=T)
```



