---
title: "MiSeq-239"
output:
  html_notebook:
    df_print: paged
    number_sections: yes
    theme: lumen
    toc: yes
    code_folding: hide
---

<style>
  html {
    font-size: 16pt;
  }
  
  body {
    font-size: 16pt;
  }
  
  h1 {
    font-size: 2.2rem;
  }
  
  h2 {
    font-size: 2rem;
  }
  
  h3 {
    font-size: 1.8rem;
  }
  
  h4 {
    font-size: 1.4rem;
  }
  
</style>

# Alpha Diversity: Data Prep {.tabset}

## Clustering Level
Data Clustered at the Genus level.

## Setup 

### [Close]

### Start Conda ENV
```{r}
startCondaEnv('alphadiv')
```

### Set Knitr Options
```{r}
library(rmarkdown)
library(knitr)

opts_chunk$set(
  echo=TRUE,
  dpi=300,
  fig.width=12
  )
```

### Load Libraries
```{r}
library(tidyverse)
library(magrittr)
library(ggplot2)
library(ggbeeswarm)
library(openxlsx)
library(DT)
library(pheatmap)
library(kableExtra)
library(vegan)
library(nlme)

### Custom libraries that can be loaded from GitHub
source('~/utilities/analysis-utilities/general_asv_data_analysis_utilities.R')
source('~/utilities/analysis-utilities/alpha_div_utilities.R')
```

### Set Paths
```{r}
source('./set_paths.R')
```

### Load Data
```{r}
asv_table = 
  read.delim(
    file.path(dada2_table_dir, 'miseq-239_tall_asv_table.tsv'), 
    header=T, 
    sep='\t',
    stringsAsFactors=F
    )

taxonomy_table = 
  read.delim(
    file.path(dada2_table_dir, 'miseq-239_taxonomy_table.tsv'), 
    header=T, 
    sep='\t',
    stringsAsFactors=F
    ) %>%
  mutate_all(as.character())

sample_metadata = 
  read.delim(
    file.path(project_dir, 'metadata', 'miseq-239_metadata_CNV_GA.tsv'), 
    header=T, 
    sep='\t',
    stringsAsFactors=F
    ) %>%
  filter(SampleID!='Undetermined')

metadata_columns = colnames(sample_metadata) %>% setdiff('SampleName')
sample_names = sample_metadata$SampleName

asv_table_cols = 
  data.frame(
    SampleID=colnames(asv_table)
  ) %>%
  inner_join(sample_metadata, by='SampleID') %>%
  pull(SampleName)

colnames(asv_table) = c('ASVs', asv_table_cols)

pval_names = list(
  CaseString='CaseStringAMD',
  Gender='GenderM',
  AREDS='AREDSY',
  Age='Age'
)
```

### Get Filtered Taxa Abundance
```{r}
genus_taxa_abundance = getFilteredTaxaCounts(
  asv_table,
  taxonomy_table,
  sample_metadata, 
  lowest_rank='Genus',
  relative_abundance_cutoff=0.002,
  prevalence_cutoff=0,
  id_col="SampleName",
  normalize=T
  )
```

# Firmicute and Bacteroidetes Table

## Create Master TAble

```{r}
master_table = getMasterTable(
  genus_taxa_abundance,
  sample_metadata,
  'SampleName',
  'short_glommed_taxa'
)
```

## Write TMaster able
```{r}
# filename = file.path(tables_dir, 'miseq-239_alpha_diversity_genus_master_table.xlsx')
# 
# write.xlsx(
#   master_table,
#   file=filename,
#   quote=F
# )
```

## Calculate Diversity Indices {.tabset}

### [Close]  

### All
```{r}
all_alpha_diversity = calculateAlphaDiversity2(
  master_table,
  genus_taxa_abundance$short_glommed_taxa,
  c('observed', 'shannon', 'simpson')
)

amd_only_alpha_diversity = calculateAlphaDiversity2(
  master_table %>% filter(CaseString=='AMD'),
  genus_taxa_abundance$short_glommed_taxa,
  c('observed', 'shannon', 'simpson')
)

age_lt_90_alpha_diversity = calculateAlphaDiversity2(
  master_table %>% filter(Age<90),
  genus_taxa_abundance$short_glommed_taxa,
  c('observed', 'shannon', 'simpson')
)

all_master_table =
  inner_join(all_alpha_diversity, master_table, by='SampleName') %>%
  select(SampleName, Observed=observed, Shannon=shannon, Simpson=simpson, metadata_columns, everything())

amd_only_master_table =
  inner_join(amd_only_alpha_diversity, master_table, by='SampleName') %>%
  select(SampleName, Observed=observed, Shannon=shannon, Simpson=simpson, metadata_columns, everything())


age_lt_90_master_table =
  inner_join(age_lt_90_alpha_diversity, master_table, by='SampleName') %>%
  select(SampleName, Observed=observed, Shannon=shannon, Simpson=simpson, metadata_columns, everything())



alpha_div_indices = createWorkbook()

addWorksheet(alpha_div_indices, sheetName="All Subgroups")
openxlsx::writeData(alpha_div_indices, sheet="All Subgroups", all_master_table)
addWorksheet(alpha_div_indices, sheetName="Age < 90")
openxlsx::writeData(alpha_div_indices, sheet="Age < 90", age_lt_90_master_table)
addWorksheet(alpha_div_indices, sheetName="AMD Only")
openxlsx::writeData(alpha_div_indices, sheet="AMD Only", amd_only_master_table)

filename = file.path(tables_dir, 'miseq-239_genus_clust_alpha_diversity_master_tables.xlsx')

saveWorkbook(alpha_div_indices, file=filename, overwrite=T)

```
