---
title: "MiSeq-239"
output:
  html_notebook:
    df_print: paged
    number_sections: yes
    theme: lumen
    toc: yes
    code_folding: hide
---

<style>
  html {
    font-size: 16pt;
  }
  
  body {
    font-size: 16pt;
  }
  
  h1 {
    font-size: 2.2rem;
  }
  
  h2 {
    font-size: 2rem;
  }
  
  h3 {
    font-size: 1.8rem;
    
  }
  
  h4 {
    font-size: 1.4rem;
  }
  
</style>

# Alapha Diversity -- Plots {.tabset}
## Clustering Level
Data Clustered at the Genus level.

## Setup
### Start Conda ENV
```{r}
startCondaEnv('alphadiv')
```

### Set Knitr Options
```{r}
library(rmarkdown)
library(knitr)

opts_chunk$set(
  echo=TRUE,
  dpi=300,
  fig.width=12
  )
```

### Load Libraries
```{r}
library(tidyverse)
library(magrittr)
library(ggplot2)
library(ggbeeswarm)
library(openxlsx)
library(DT)
library(pheatmap)
library(kableExtra)
library(vegan)
library(nlme)

### Custom libraries that can be loaded from GitHub
source('~/utilities/analysis-utilities/general_asv_data_analysis_utilities.R')
source('~/utilities/analysis-utilities/alpha_div_utilities.R')
```


### Set Paths
```{r}
source('./set_paths.R')
```

### Load Data
```{r}
filename = file.path(tables_dir, 'miseq-239_genus_clust_alpha_diversity_master_tables.xlsx')

all_master_table = readWorkbook(filename, sheet='All Subgroups')
age_lt_90_master_table = readWorkbook(filename, sheet='Age < 90')
amd_only_master_table = readWorkbook(filename, sheet='AMD Only')

source('./observational_variables.R')

sample_metadata = read.delim(
  file.path(metadata_dir, 'miseq-239_metadata_CNV_GA.tsv'), 
  stringsAsFactors=F
  )

alpha_diversity = 
  all_master_table %>%
  select(SampleName, Observed, Shannon, Simpson)

filename = file.path(tables_dir, 'miseq-239_alpha_div_genus_clust_mann_whitney_unadjusted.xlsx')

pair_stats = 
  read.xlsx(filename)

```

### Variables of interest
#### All Observational Variables
```{r}
# print(observational_variables)
```

#### AMD Only Observational Variables
```{r}
amd_only_variables = observational_variables

amd_only_excluded_variables = c(
  'Gender',
  'CaseString',
  'Age'
)

for (varname in amd_only_excluded_variables)
{
  print(sprintf('Excluding %s', varname))
  amd_only_variables[varname] = NULL
}

```

**Excluding: ** `r amd_only_excluded_variables`


#### Calculated Experimental Variables

```{r}
calculated_exp_vars = c('Observed', 'Shannon', 'Simpson')
```
**Calculated Experimental Variables: ** `r calculated_exp_vars`

#### Subgroups
```{r}
subgroups = list(
  All=list(name='All Subgroups', prefix='all'),
  Age_lt_90=list(name='Age < 90', prefix='age_lt_90'),
  Amd_Only=list(name='AMD Only', prefix='amd_only')
)
```


### Set Metadata Factor Levels
```{r}
all_master_table = setFactorsLevels(
  all_master_table, 
  observational_variables
  )
age_lt_90_master_table = setFactorsLevels(
  age_lt_90_master_table, 
  observational_variables
  )
amd_only_master_table = setFactorsLevels(
  amd_only_master_table, 
  observational_variables
  )
```

### P-Value Annotationas
```{r}
pval_annotations = getPvalAnnotations(
  variables=c(calculated_exp_vars),
  master_table=all_master_table,
  pval_height_factor=1.2,
  pair_stats=pair_stats
)
```


## Plots
### All Subgroups
#### Indices
```{r}
makeAlphaDivPlot(
  all_master_table,
  indices=calculated_exp_vars,
  aesthetics = list(x_var='index', color_var='index', facet_var='index')
) + 
  ggtitle("Alpha Diversity") +
  ylab("Value")
```

#### AMD vs. Control
```{r}
makePairPlot(
  master_table=all_master_table,
  indices=calculated_exp_vars,
  var_data=observational_variables$CaseString,
  aesthetics=list(
    x_var='CaseString', 
    color_var='CaseString', 
    facet_var='index'
    ),
  annotation_data=pval_annotations %>% filter(TestGroup=='All'),
  title_template="Alpha Diversity -- All Subgroups: "
)


```

### Age < 90
#### Indices
```{r}
makeAlphaDivPlot(
  age_lt_90_master_table,
  indices=calculated_exp_vars,
  aesthetics = list(x_var='index', color_var='index', facet_var='index')
) + 
  ggtitle("Alpha Diversity: Age < 90") +
  ylab("Value") 
```

#### AMD vs. Control
```{r}

makePairPlot(
  master_table=age_lt_90_master_table,
  indices=calculated_exp_vars,
  var_data=observational_variables$CaseString,
  aesthetics = list(x_var='CaseString', color_var='CaseString', facet_var='index'),
  annotation_data=pval_annotations %>% filter(TestGroup=='All'),
  title_template="Alpha Diversity -- Age < 90: "
)

```


### AMD Only

#### Indices
```{r}
makeAlphaDivPlot(
  age_lt_90_master_table,
  indices=calculated_exp_vars,
  aesthetics = list(x_var='index', color_var='index', facet_var='index')
) + 
  ggtitle("Alpha Diversity -- AMD Only") +
  ylab("Value")
```

#### Covariates

```{r}
plotPairTests(
  variables=amd_only_variables,
  master_table=all_master_table,
  indices=calculated_exp_vars,
  title_template="Alpha Diversity AMD Only: ",
  annotation_data=pval_annotations %>% filter(TestGroup=='AMD_Only')
)

```
