---
title: "MiSeq"
output:
  html_notebook:
  df_print: paged
number_sections: yes
theme: lumen
toc: yes
code_folding: hide
---

  <style>
  html {
    font-size: 16pt;
  }

body {
  font-size: 16pt;
}

h1 {
  font-size: 2.2rem;
}

h2 {
  font-size: 2rem;
}

h3 {
  font-size: 1.8rem;
}

h4 {
  font-size: 1.4rem;
}

</style>


```{r}
source('/home/balter/miseq-0278/metadata/project_metadata.R')
analysis_type = "igaseq"
clustering_level = "asv"
tables_dir = file.path(analysis_dir, analysis_type, clustering_level, 'tables')
analysis_title = "Igaseq"
```

# MiSeq-0278 `r analysis_title` Two Group Tests {.tabset}
## Clustering Level
Data clustered at the `r toupper(clustering_level)` Level.

## Setup

### Start Conda ENV
```{r}
startCondaEnv('alphadiv')
```

### Set Knitr Options
```{r}
library(rmarkdown)
library(knitr)

${knitr_options}

```

### Load Libraries
```{r}
library(tidyverse)
library(magrittr)
library(DT)
library(openxlsx)
library(ggplot2)
library(ggbeeswarm)

### Custom libraries that can be loaded from GitHub
source('~/utilities/analysis-utilities/general_asv_data_analysis_utilities.R')
source('~/utilities/analysis-utilities/alpha_div_utilities.R')
source('~/utilities/analysis-utilities/amd_project_utilities.R')
```

### Load Data
```{r}
filename = makeDataFileName(
  'master_tables.xlsx',
  "${tables_dir}",
  "${analysis_type}",
  "${miseq_project_prefix}",
  "${clustering_level}"
)

iga_pos_master_table = readWorkbook(filename, sheet='Iga Pos')
iga_neg_master_table = readWorkbook(filename, sheet='Iga Neg')
iga_index_master_table = readWorkbook(filename, sheet='Iga Index')
ici_score_master_table = readWorkbook(filename, sheet='Ici Score')
alt_index_master_table = readWorkbook(filename, sheet='Alt Index')
top_n_features = readWorkbook(filename, sheet='Taxa') %>%
  unlist() %>%
  unname()

filename = file.path(
  "${project_root}", 
  "${metadata_dir}", 
  'observational_variables.R'
  )
source(filename)
```

**Included Taxa: `print(top_n_features)`


### Set Factor Levels
```{r}
iga_index_master_table = setFactorsLevels(
  iga_index_master_table, 
  observational_variables
  )
ici_score_master_table = setFactorsLevels(
  ici_score_master_table, 
  observational_variables
  )
alt_index_master_table = setFactorsLevels(
  alt_index_master_table, 
  observational_variables
  )
# sample_metadata = setFactorsLevels(sample_metadata, categorical_variables)
```



## Preditor Variables (observational)
```{r}
# print(observational_variables)
```
**Observational Variables of Interest:** `r observational_variables`


### Categorical Variables
```{r}
categorical_variable_names = c(
  'Gender',
  'CaseString',
  'ARMS2rs10490924', 
  'CFHrs1061170', 
  'CFHrs10737680', 
  'SKIV2Lrs429608', 
  'CNV_Either_Eye', 
  'GA_No_CNV_Either_Eye', 
  'AREDS'
)
categorical_variables = 
  observational_variables[categorical_variable_names]
observational_variable_names = observational_variables %>% names() %>% unname()
```

### Determine which covariates have sufficient samples for each value
```{r}
num_samples_threshold = 5

all_variables = filterVarsBySampleThreshold(
  iga_index_master_table,
  num_samples_threshold,
  observational_variables
  )

age_lt_90_variables = filterVarsBySampleThreshold(
  iga_index_master_table,
  num_samples_threshold,
  observational_variables
  )

amd_only_variables = filterVarsBySampleThreshold(
  iga_index_master_table %>% select(-CaseString),
  num_samples_threshold,
  observational_variables
  )

control_only_variables = filterVarsBySampleThreshold(
  iga_index_master_table %>% filter(CaseString=='Control') %>% select(-CaseString),
  num_samples_threshold,
  observational_variables
  )

```


```
#### AMD Only Observational Variables


amd_only_variables = observational_variables

amd_only_excluded_variables = c(
  'Gender',
  'CaseString',
  'Age'
)

for (varname in amd_only_excluded_variables)
{
  print(sprintf('Excluding %s', varname))
  amd_only_variables[varname] = NULL
}


**Excluding: ** `r amd_only_excluded_variables`
```


## Response Variables

```{r}
raw_exp_vars = ${raw_exp_vars}
calculated_exp_vars = ${calculated_exp_vars}
all_experimental_variables = c(raw_exp_vars, calculated_exp_vars)
```
**Raw Experimental Variables: ** `r raw_exp_vars`
**Calculated Experimental Variables: ** `r calculated_exp_vars`

## Regression analysis

### General Parameters
```{r}
stats = ${stats}
use_parametric = ${use_parametric}
adjustment_method = "${adjustment_method}"
response_var_name = "${response_var_name}"
```


### All

```{r}
all_iga_index_regression_stats = getMultipleRegressionStats(
  predictor_vars=all_variables,
  response_vars=top_n_features,
  master_table=iga_index_master_table,
  stats=stats,
  response_var_name=response_var_name,
  adjustment_method=adjustment_method,
  parametric=use_parametric
)

all_ici_score_regression_stats = getMultipleRegressionStats(
  predictor_vars=all_variables,
  response_vars=top_n_features,
  master_table=ici_score_master_table,
  stats=stats,
  response_var_name=response_var_name,
  adjustment_method=adjustment_method,
  parametric=use_parametric 
)

```


### Age < 90
```{r}
age_lt_90_iga_index_regression_stats = getMultipleRegressionStats(
  predictor_vars=age_lt_90_variables,
  response_vars=top_n_features,
  master_table=iga_index_master_table %>% filter(Age < 90),
  stats=stats,
  response_var_name=response_var_name,
  adjustment_method=adjustment_method,
  parametric=use_parametric
)

age_lt_90_ici_score_regression_stats = getMultipleRegressionStats(
  predictor_vars=age_lt_90_variables,
  response_vars=top_n_features,
  master_table=ici_score_master_table %>% filter(Age < 90),
  stats=stats,
  response_var_name=response_var_name,
  adjustment_method=adjustment_method,
  parametric=use_parametric 
)

```

### AMD Only
```{r}
amd_only_iga_index_regression_stats = getMultipleRegressionStats(
  predictor_vars=amd_only_variables,
  response_vars=top_n_features,
  master_table=iga_index_master_table %>% filter(CaseString=='AMD'),
  stats=stats,
  response_var_name=response_var_name,
  adjustment_method=adjustment_method,
  parametric=use_parametric
)

amd_only_ici_score_regression_stats = getMultipleRegressionStats(
  predictor_vars=amd_only_variables,
  response_vars=top_n_features,
  master_table=ici_score_master_table %>% filter(CaseString=='AMD'),
  stats=stats,
  response_var_name=response_var_name,
  adjustment_method=adjustment_method,
  parametric=use_parametric 
)

```


## Write Regression Stats
```{r}

regression_stats_table_names = c(
  "all_iga_index", 
  "age_lt_90_iga_index",
  "amd_only_iga_index",
  "all_ici_score",
  "age_lt_90_ici_score",
  "amd_only_ici_score"
  )


for (tab_name in regression_stats_table_names)
{
  contrast_regression_stats = createWorkbook()
  
  print(tab_name)
  stats_table_name = paste0(tab_name, "_regression_stats")
  print(stats_table_name)
  stats_table = get(stats_table_name)
  for (contrast in contrast_names)
  {
    print(contrast)
    addWorksheet(contrast_regression_stats, sheetName=contrast)
    openxlsx::writeData(
    wb=contrast_regression_stats, 
    sheet=contrast, 
    x=stats_table[[contrast]]
    )
  }
  
  filename = makeDataFileName(
    paste0(stats_table_name, '_regression_stats_tables.xlsx'),
    "${tables_dir}",
    "${analysis_type}",
    "${miseq_project_prefix}",
    "${clustering_level}"
  )
  print(filename)
  
  saveWorkbook(contrast_regression_stats, file=filename, overwrite=T)
  
}

```
