---
title: "MiSeq-239"
output:
  html_notebook:
    df_print: paged
    number_sections: yes
    theme: lumen
    toc: yes
    code_folding: hide
---

<style>
  html {
    font-size: 16pt;
  }
  
  body {
    font-size: 16pt;
  }
  
  h1 {
    font-size: 2.2rem;
  }
  
  h2 {
    font-size: 2rem;
  }
  
  h3 {
    font-size: 1.8rem;
  }
  
  h4 {
    font-size: 1.4rem;
  }
  
</style>

# Alpha Diversity: Regression Analysis {.tabset}

## Clustering Level
Data Clustered at the Genus level.

## Setup {.tabset}

### [Close]

### Sourcing setup file
```{r}
source('setup.R')
```

### Start Conda ENV
```{r}
startCondaEnv('alphadiv')
```

### Set Knitr Options
```{r}
library(rmarkdown)
library(knitr)

opts_chunk$set(
  echo=TRUE,
  dpi=300,
  fig.width=12
  )
```

### Load Libraries
```{r}
library(tidyverse)
library(magrittr)
library(ggplot2)
library(ggbeeswarm)
library(openxlsx)
library(DT)
library(pheatmap)
library(kableExtra)
library(vegan)
library(nlme)

### Custom libraries that can be loaded from GitHub
source('~/utilities/analysis-utilities/general_asv_data_analysis_utilities.R')
source('~/utilities/analysis-utilities/alpha_div_utilities.R')
source('~/utilities/analysis-utilities/project_utilities.R')
```


### Set Paths
```{r}
source('./set_paths.R')
```

### Load Data
```{r}
filename = file.path(tables_dir, 'miseq-239_genus_clust_alpha_diversity_master_tables.xlsx')

all_master_table = readWorkbook(filename, sheet='All Subgroups')
age_lt_90_master_table = readWorkbook(filename, sheet='Age < 90')
amd_only_master_table = readWorkbook(filename, sheet='AMD Only')

sample_metadata = read.delim(
  file.path(metadata_dir, 'miseq-239_metadata_CNV_GA.tsv'), 
  stringsAsFactors=F
  )

source(file.path(metadata_dir, 'observational_variables.R'))
```

### Variables of interest
#### All Observational Variables
```{r}
# print(observational_variables)
```

#### Categorical Variables
```{r}
categorical_variable_names = c(
  'Gender',
  'CaseString',
  'ARMS2rs10490924', 
  'CFHrs1061170', 
  'CFHrs10737680', 
  'SKIV2Lrs429608', 
  'CNV_Either_Eye', 
  'GA_No_CNV_Either_Eye', 
  'AREDS'
)
categorical_variables = 
  observational_variables[categorical_variable_names]
```

#### Filter by Sample Number Threshold
```
num_samples_threshold = 5

all_variables = filterVarsBySampleThreshold(
  all_master_table, 
  num_samples_threshold,
  observational_variables
  )

age_lt_90_variables = filterVarsBySampleThreshold(
  age_lt_90_master_table, 
  num_samples_threshold,
  observational_variables
  )

amd_only_variables = filterVarsBySampleThreshold(
  amd_only_master_table %>% select(-CaseString), 
  num_samples_threshold,
  observational_variables
  )


all_formula_rhs = paste0(names(all_variables), collapse=' + ')
age_lt_90_formula_rhs = paste0(names(age_lt_90_variables), collapse=' + ')
amd_only_formula_rhs = paste0(names(amd_only_variables), collapse=' + ')
```


#### AMD Only Observational Variables
```{r}
amd_only_variables = observational_variables

amd_only_excluded_variables = c(
  'Gender',
  'CaseString',
  'Age'
)

for (varname in amd_only_excluded_variables)
{
  print(sprintf('Excluding %s', varname))
  amd_only_variables[varname] = NULL
}

```

**Excluding: ** `r amd_only_excluded_variables`

#### Calculated Experimental Variables

```{r}
calculated_exp_vars = c('Observed', 'Shannon', 'Simpson')
```
**Calculated Experimental Variables: ** `r calculated_exp_vars`

#### Subgroups
```{r}
subgroups = list(
  All=list(name='All Subgroups', prefix='all'),
  Age_lt_90=list(name='Age < 90', prefix='age_lt_90'),
  Amd_Only=list(name='AMD Only', prefix='amd_only')
)
```


### Set Metadata Factor Levels
```{r}
all_master_table = setFactorsLevels(all_master_table, categorical_variables)
age_lt_90_master_table = setFactorsLevels(age_lt_90_master_table, categorical_variables)
amd_only_master_table = setFactorsLevels(amd_only_master_table, categorical_variables)
sample_metadata = setFactorsLevels(sample_metadata, categorical_variables)
```


### Determine which covariates have sufficient samples with each value
```{r}
num_samples_threshold = 3

all_variables = filterVarsBySampleThreshold(
  all_master_table,
  num_samples_threshold,
  observational_variables
  )

age_lt_90_variables = filterVarsBySampleThreshold(
  age_lt_90_master_table,
  num_samples_threshold,
  observational_variables
  )

amd_only_variables = filterVarsBySampleThreshold(
  amd_only_master_table %>% select(-CaseString),
  num_samples_threshold,
  observational_variables
  )

control_only_variables = filterVarsBySampleThreshold(
  all_master_table %>% filter(CaseString=='Control') %>% select(-CaseString),
  num_samples_threshold,
  observational_variables
  )

# all_variables = observational_variables
# age_lt_90_variables = all_variables

```


### Create p-val names for contrasts

```{r}
all_contrast_names = makeContrastNames(
  all_variables, 
  all_master_table
  )

age_lt_90_contrast_names = makeContrastNames(
  age_lt_90_variables, 
  all_master_table
  )

amd_only_contrast_names = makeContrastNames(
  amd_only_variables, 
  all_master_table
  )

control_only_contrast_names = makeContrastNames(
  control_only_variables, 
  all_master_table
  )
```

### Initialize regression stats table

```{r}
regression_pvals = makeRegressionStatsContrastTemplate(flatten_chr(all_contrast_names))

regression_effect_sizes = makeRegressionStatsContrastTemplate(flatten_chr(all_contrast_names))
```

## Linear Regression {.tabset}

### All

```{r}
exp_vars = calculated_exp_vars
predictors = control_only_variables
master_table = all_master_table
contrast_names = control_only_contrast_names %>% flatten_chr()
index_name_template = 'All_'

# regression_pvals = regression_pvals_template
# regression_effect_sizes = regression_pvals_template

for (var in exp_vars)
{
  index_name = paste0(index_name_template, var)
  print(index_name)
  
  fit = doMultipleRankRegression(
    var,
    predictors,
    master_table
  )
  
  # print(vif(fit))
  
  pvals = getRegressionPvals(
    fit,
    contrast_names,
    index_name
  )
  
  regression_pvals = regression_pvals %>% add_row(!!!pvals)

  effect_sizes = getRegressionEffectSizes(
    fit,
    contrast_names,
    index_name
  )

  regression_effect_sizes = regression_effect_sizes %>% add_row(!!!effect_sizes)

}
```

### Age < 90

```{r}
exp_vars = calculated_exp_vars
predictors = age_lt_90_variables
master_table = age_lt_90_master_table
contrasts = age_lt_90_contrast_names %>% unlist()
index_name_template = 'Age_lt_90_'

# regression_pvals = regression_pvals_template
# regression_effect_sizes = regression_pvals_template

for (var in exp_vars)
{
  index_name = paste0(index_name_template, var)
  print(index_name)
  
  fit = doMultipleRankRegression(
    var,
    predictors,
    master_table
  )
  
  pvals = getRegressionPvals(
    fit,
    unlist(contrast_names),
    index_name
  )
  
  regression_pvals = regression_pvals %>% add_row(!!!pvals)

  effect_sizes = getRegressionEffectSizes(
    fit,
    unlist(contrast_names),
    index_name
  )

  regression_effect_sizes = regression_effect_sizes %>% add_row(!!!effect_sizes)

}
```


### AMD Only

```{r}
exp_vars = calculated_exp_vars
predictors = amd_only_variables
master_table = amd_only_master_table
contrasts = amd_only_contrast_names %>% unlist()
index_name_template = 'AMD_Only_'

# regression_pvals = regression_pvals_template
# regression_effect_sizes = regression_pvals_template

for (var in exp_vars)
{
  index_name = paste0(index_name_template, var)
  print(index_name)
  
  fit = doMultipleRankRegression(
    var,
    predictors,
    master_table
  )
  
  pvals = getRegressionPvals(
    fit,
    contrast_names,
    index_name
  )
  
  regression_pvals = regression_pvals %>% add_row(!!!pvals)

  effect_sizes = getRegressionEffectSizes(
    fit,
    contrast_names,
    index_name
  )

  regression_effect_sizes = regression_effect_sizes %>% add_row(!!!effect_sizes)

}
```
### Visualize
```{r}
pheatmap(regression_pvals %>% column_to_rownames('Index'), cluster_cols=F, cluster_rows=F)

pheatmap(regression_effect_sizes %>% column_to_rownames('Index'), cluster_cols=F, cluster_rows=F)
```



### Write Tables
```{r}
regression_stats = createWorkbook()

addWorksheet(
  regression_stats, 
  sheetName="Unadjusted P-Values"
  )

openxlsx::writeData(
  regression_stats, 
  sheet="Unadjusted P-Values", 
  regression_pvals
  )

addWorksheet(
  regression_stats, 
  sheetName="Effect Sizes"
  )

openxlsx::writeData(
  regression_stats,
  sheet="Effect Sizes", 
  regression_effect_sizes
  )

# addWorksheet(regression_pvals, sheetName="Adjusted P-Values")
# openxlsx::writeData(regression_pvals, sheet="Adjusted P-Values", adjusted_regression_pvals)

filename = file.path(tables_dir, 'miseq-239_genus_clust_alpha_div_regression_stats.xlsx')

saveWorkbook(regression_stats, file=filename, overwrite=T)
```



```{r}
categorical_data = 
  sample_metadata %>%
  select(
    Gender,
    CaseString,
    ARMS2rs10490924, 
    CFHrs1061170, 
    CFHrs10737680, 
    SKIV2Lrs429608, 
    CNV_Either_Eye, 
    GA_No_CNV_Either_Eye, 
    AREDS
  ) 

# %>%
#   mutate_all(list(~ifelse(is.na(.), 0, 1))) 
```


```{r}
num_vars = length(categorical_variable_names)

categorical_chisq_pvals = matrix(rep(0, num_vars*num_vars), ncol=num_vars, nrow=num_vars)
colnames(categorical_chisq_pvals) = categorical_variable_names
rownames(categorical_chisq_pvals) = categorical_variable_names

for (var in categorical_variable_names)
{
  for(other_var in categorical_variable_names)
  {
    print(sprintf('var: %s, other_var: %s', var, other_var))
    
    if (var != other_var)
    {
      test =
        sample_metadata %>%
        select(var, other_var) %>%
        table() %>%
        chisq.test()

      categorical_chisq_pvals[var, other_var] = round(test$p.value,3)
    }

  }
}
  
```

