---
title: "MiSeq"
output:
  html_notebook:
    df_print: paged
    number_sections: yes
    theme: lumen
    toc: yes
    code_folding: hide
---

<style>
  html {
    font-size: 16pt;
  }
  
  body {
    font-size: 16pt;
  }
  
  h1 {
    font-size: 2.2rem;
  }
  
  h2 {
    font-size: 2rem;
  }
  
  h3 {
    font-size: 1.8rem;
  }
  
  h4 {
    font-size: 1.4rem;
  }
  
</style>

```{r}
source('${project_metadata_file}')
analysis_type = "${analysis_type}"
clustering_level = "${clustering_level}"
tables_dir = "${tables_dir}"
use_allbac = ${use_allbac}
```

# MiSeq-${miseq_run_number} FB Analysis Data Prep{.tabset}

## Setup 

### [Close]

### Start Conda ENV
```{r}
startCondaEnv('${conda_env}')
```

### Set Knitr Options
```{r}
library(rmarkdown)
library(knitr)

${knitr_options}
```

### Load Libraries
```{r}
library(tidyverse)
library(magrittr)
library(ggplot2)
library(ggbeeswarm)
library(openxlsx)
library(DT)
library(pheatmap)
library(kableExtra)

### Custom libraries that can be loaded from GitHub
source('~/utilities/analysis-utilities/general_asv_data_analysis_utilities.R')
source('~/utilities/analysis-utilities/alpha_div_utilities.R')
source('~/utilities/analysis-utilities/amd_project_utilities.R')
```

### Load Data
```{r}
asv_table = 
  read.delim(
    '${asv_table_file}', 
    header=T, 
    sep='\t',
    stringsAsFactors=F
    )

taxonomy_table = 
  read.delim(
    '${taxonomy_table_file}', 
    header=T, 
    sep='\t',
    stringsAsFactors=F
    ) %>%
  mutate_all(as.character())

sample_data = 
  read.delim(
    '${sample_data_file}', 
    header=T, 
    sep='\t',
    stringsAsFactors=F
    ) %>%
  filter(SampleID!='Undetermined')


if ("IGA" %in% colnames(sample_data) & use_allbac)
{
  sample_data = 
    sample_data %>%
    filter(IGA=='AllBac')

  sample_names = sample_data$SampleName
  
  asv_table = 
    asv_table %>%
    select('ASVs', ends_with('AllBac'))
}

sample_data_columns = colnames(sample_data)
sample_names = sample_data$SampleName

```

### Set Filtering
```{r}
relative_abundance_cutoff = ${relative_abundance_cutoff}
prevalence_cutoff = ${prevalence_cutoff}
```
**relative abundance cutoff:** `r relative_abundance_cutoff`
**prevalence cutoff:** `r prevalence_cutoff`

### Get Filtered Taxa Abundance
```{r}
genus_taxa_abundance = getFilteredTaxaCounts(
  asv_table,
  taxonomy_table,
  sample_data, 
  lowest_rank='Genus',
  relative_abundance_cutoff=relative_abundance_cutoff,
  prevalence_cutoff=prevalence_cutoff,
  id_col="SampleName",
  normalize=T
  )

filtered_taxa = genus_taxa_abundance$short_glommed_taxa
```

## Create Master Table
```{r}
genus_master_table = getMasterTable(
  genus_taxa_abundance,
  sample_data,
  'SampleName',
  'short_glommed_taxa'
)
```


## Write Table
```{r}
logistic_regression_prep = createWorkbook()

addWorksheet(
  wb=logistic_regression_prep, 
  sheetName="Master Table"
  )

openxlsx::writeData(
  wb=logistic_regression_prep, 
  sheet="Master Table", 
  x=genus_master_table
  )

addWorksheet(
  wb=logistic_regression_prep, 
  sheetName="Filtered Taxa"
  )

openxlsx::writeData(
  wb=logistic_regression_prep,
  sheet="Filtered Taxa", 
  x=filtered_taxa
  )

filename = makeDataFileName(
  'master_table.xlsx',
  "${tables_dir}",
  "${analysis_type}",
  "${miseq_project_prefix}",
  "${clustering_level}"
)

print(filename)

saveWorkbook(logistic_regression_prep, file=filename, overwrite=T)
```


