---
title: "MiSeq-${miseq_run_number}"
output:
  html_notebook:
    df_print: paged
    number_sections: yes
    theme: lumen
    toc: yes
    code_folding: hide
---

<style>
  html {
    font-size: 16pt;
  }
  
  body {
    font-size: 16pt;
  }
  
  h1 {
    font-size: 2.2rem;
  }
  
  h2 {
    font-size: 2rem;
  }
  
  h3 {
    font-size: 1.8rem;
  }
  
  h4 {
    font-size: 1.4rem;
  }
  
</style>



```{r}
source('${project_metadata_file}')
analysis_type = "${analysis_type}"
clustering_level = "${clustering_level}"
tables_dir = "${tables_dir}"
analysis_title = "${analysis_title}"
use_allbac = ${use_allbac}

```

# MiSeq-${miseq_run_number} `r analysis_title` Data Prep {.tabset}
## Clustering Level
Data clustered at the `r toupper('${clustering_level}')` Level.

## Filtering
* **Relative Abundance Cutoff:** ${relative_abundance_cutoff}
* **Prevalence Cutoff:** ${prevalence_cutoff}
* **Min Count Cutoff:** ${min_count_cutoff}

## Setup 
### Start Conda ENV
```{r}
startCondaEnv('alphadiv')
```

### [Close]

### Load Libraries
```{r}
library(rmarkdown)
library(knitr)

library(tidyverse)
library(magrittr)
library(DT)
library(openxlsx)

### Custom libraries that can be loaded from GitHub
source('~/utilities/analysis-utilities/general_asv_data_analysis_utilities.R')
source('~/utilities/analysis-utilities/alpha_div_utilities.R')
source('~/utilities/amd_templates/setup/amd_project_utilities.R')
```

### Set Knitr Options
```{r}

${knitr_options}
```

### Load Data
```{r}
asv_table = 
  read.delim(
    '${asv_table_file}', 
    header=T, 
    sep='\t',
    stringsAsFactors=F
    )

taxonomy_table = 
  read.delim(
    '${taxonomy_table_file}', 
    header=T, 
    sep='\t',
    stringsAsFactors=F
    ) %>%
  mutate_all(as.character())

sample_data = 
  read.delim(
    '${sample_data_file}', 
    header=T, 
    sep='\t',
    stringsAsFactors=F
    ) %>%
  filter(SampleID!='Undetermined')

if ("IGA" %in% colnames(sample_data) & use_allbac)
{
  sample_data = 
    sample_data %>%
    filter(IGA=='AllBac')

  sample_names = sample_data$SampleName
  
  asv_table = 
    asv_table %>%
    select('ASVs', ends_with('AllBac'))
}

sample_data_columns = colnames(sample_data)

```

### Set Filtering
```{r}
relative_abundance_cutoff = ${relative_abundance_cutoff}
prevalence_cutoff = ${prevalence_cutoff}
```
**relative abundance cutoff:** `r relative_abundance_cutoff`
**prevalence cutoff:** `r prevalence_cutoff`

### Get Filtered Taxa Abundance
```{r}
genus_taxa_abundance = getFilteredTaxaCounts(
  asv_table,
  taxonomy_table,
  sample_data, 
  lowest_rank="${clustering_level}",
  relative_abundance_cutoff=relative_abundance_cutoff,
  prevalence_cutoff=prevalence_cutoff,
  id_col="SampleName",
  normalize=T
  )
```

### Create Master Table

```{r}
master_table = getMasterTable(
  genus_taxa_abundance,
  sample_data,
  'SampleName',
  'short_glommed_taxa'
)
```

## Calculate Alpha Diversity {.tabset}

### [Close]  

### Subgroups
```{r}
all_alpha_diversity = calculateAlphaDiversity2(
  master_table,
  genus_taxa_abundance$short_glommed_taxa,
  c('observed', 'shannon', 'simpson')
)

amd_only_alpha_diversity = calculateAlphaDiversity2(
  master_table %>% filter(CaseString=='AMD'),
  genus_taxa_abundance$short_glommed_taxa,
  c('observed', 'shannon', 'simpson')
)

age_lt_90_alpha_diversity = calculateAlphaDiversity2(
  master_table %>% filter(Age<90),
  genus_taxa_abundance$short_glommed_taxa,
  c('observed', 'shannon', 'simpson')
)

all_master_table =
  inner_join(all_alpha_diversity, master_table, by='SampleName') %>%
  select(SampleName, Observed=observed, Shannon=shannon, Simpson=simpson, sample_data_columns, everything())

amd_only_master_table =
  inner_join(amd_only_alpha_diversity, master_table, by='SampleName') %>%
  select(SampleName, Observed=observed, Shannon=shannon, Simpson=simpson, sample_data_columns, everything())


age_lt_90_master_table =
  inner_join(age_lt_90_alpha_diversity, master_table, by='SampleName') %>%
  select(SampleName, Observed=observed, Shannon=shannon, Simpson=simpson, sample_data_columns, everything())
```

### Write Alpha Diversity Tables
```{r}
wb = createWorkbook()

all_ranks = c('Phylum', 'Class', 'Order', 'Family', 'Genus')

if (clustering_level == 'ASV')
{
  ranks_to_keep = c(all_ranks, 'ASVs')
} else
{
  ranks_to_keep = all_ranks[1:match("${clustering_level}", all_ranks)]
}

addWorksheet(wb, sheetName="Master Table")
openxlsx::writeData(wb, sheet="Master Table", all_master_table)

just_taxa = genus_taxa_abundance %>% select(ranks_to_keep, glommed_taxa, short_glommed_taxa)
addWorksheet(wb, sheetName="${tools::toTitleCase(clustering_level)}")
openxlsx::writeData(wb, sheet="${tools::toTitleCase(clustering_level)}", just_taxa)

filename = makeDataFileName(
  'master_table.xlsx',
  "${tables_dir}",
  "${analysis_type}",
  "${miseq_project_prefix}",
  "${clustering_level}"
)
print(sprintf("master table file: %s", filename))

saveWorkbook(wb, file=filename, overwrite=T)
```
