---
title: "${miseq_project_prefix}: ${analysis_type}"
output:
  html_notebook:
    df_print: paged
    number_sections: yes
    theme: lumen
    toc: yes
    code_folding: hide
---

<style>
  html {
    font-size: 16pt;
  }
  
  body {
    font-size: 16pt;
  }
  
  h1 {
    font-size: 2.2rem;
  }
  
  h2 {
    font-size: 2rem;
  }
  
  h3 {
    font-size: 1.8rem;
  }
  
  h4 {
    font-size: 1.4rem;
  }
  
</style>

```{r}
source('${project_metadata_file}')
analysis_type = "${analysis_type}"
clustering_level = "${clustering_level}"
filter_by = "${filter_by}"
tables_dir = "${tables_dir}"
analysis_title = tools::toTitleCase(gsub("_", " ", analysis_type))
```

# ${miseq_project_prefix} Beta Diversity {.tabset}

**Clustering**
Data clustered at the `r toupper('${clustering_level}')` Level.

**Subgroups**
```{r, results='asis', echo=FALSE}
for (sg_filter in c(${subgroup_filters}))
{
  cat(sg_filter, '\n\n')
}
```

**Filtering**
Data filtered by `r toupper('${filter_by}')`

* **Relative Abundance cutoff:** ${relative_abundance_cutoff}
* **Prevalence Cutoff:** ${prevalence_cutoff}
* **Min Count Cutoff:** ${min_count_cutoff}
```{r}
relative_abundance_cutoff = ${relative_abundance_cutoff}
prevalence_cutoff = ${prevalence_cutoff}
min_count_cutoff = ${min_count_cutoff}
```

## Setup

### Start Conda ENV
```{r}
startCondaEnv('betadiv', lib='~/R35')
```

### Load Libraries
```{r, echo=F}
library(tidyverse)
library(magrittr)
library(ggplot2)
library(ggbeeswarm)
library(openxlsx)

library(pheatmap)
library(kableExtra)
library(vegan)
library(plotly)

### Custom libraries that can be loaded from GitHub
source('${utilities_dir}/analysis-utilities/general_asv_data_analysis_utilities.R')
source('${utilities_dir}/amd_templates/setup/amd_project_utilities.R')
source('${utilities_dir}/analysis-utilities/beta_diversity_utilities.R')
```

### Set Knitr Options
```{r}
${knitr_options}
```

### Load Data
```{r}
asv_table = 
  read.delim(
    '${asv_table_file}', 
    header=T, 
    sep='\t',
    stringsAsFactors=F
    )

taxonomy_table = 
  read.delim(
    '${taxonomy_table_file}', 
    header=T, 
    sep='\t',
    stringsAsFactors=F
    ) %>%
  mutate_all(as.character())

sample_data = 
  read.xlsx('${sample_data_file}') %>%
  mutate(
    CNV_Either_Eye = ifelse(CNV_Either_Eye == 1, '1', '0'),
    GA_No_CNV_Either_Eye = ifelse(GA_No_CNV_Either_Eye == 1, '1', '0')
  )

sample_data_columns = colnames(sample_data)
sample_names = sample_data$SampleName


source(file.path(
  "${project_root}", 
  "${metadata_dir}", 
  'observational_variables.R'
  ))

```

### Extract Subgroups
```{r}
if (!is.null(c(${subgroup_filters})))
{
  subgroup_filter_string = paste0(${subgroup_filters}, collapse=" & ")
  print(sprintf("Subgroup Filter String: %s", subgroup_filter_string))
  subgroup_filter_expression = str2expression(subgroup_filter_string)
  sample_data = sample_data %>% filter(eval(subgroup_filter_expression))
  sample_names = sample_data$SampleName

  asv_table = asv_table %>% select('ASVs', sample_names)
}

```

### Set Filtering
```{r}
relative_abundance_cutoff = ${relative_abundance_cutoff}
prevalence_cutoff = ${prevalence_cutoff}
```
**relative abundance cutoff:** `r relative_abundance_cutoff`
**prevalence cutoff:** `r prevalence_cutoff`

### Variables of interest
#### Observational Variables
```{r}
case = ${case} %>% setdiff('')
genotypes = ${genotypes} %>% setdiff('')
phenotypes = ${phenotypes} %>% setdiff('')
key_covariates = ${key_covariates} %>% setdiff('')
treatments = ${treatments} %>% setdiff('')
categorical_variables = ${categorical_variables} %>% setdiff('')

variables = c(case, key_covariates, treatments, phenotypes, genotypes)
amd_only_variables = c(key_covariates, treatments, phenotypes, genotypes)
```

* **Case:** `r case`
* **Genotype variables:** `r genotypes`
* **Phenotype variables:** `r phenotypes`
* **Treatment variables:** `r treatments`
* **Key covariates:**  `r key_covariates`


#### Remove Age
```{r}
variables %<>% setdiff('Age')
```

* **Treatment variables:** ${treatments}
* **Key covariates:**  ${key_covariates}
* **Variables Used:** `r variables`

### Get Filtered Taxa Abundance
```{r}
taxa_abundance = getFilteredTaxaCounts(
  asv_table,
  taxonomy_table,
  sample_data, 
  cluster_by="${clustering_level}",
  filter_by="${filter_by}",
  relative_abundance_cutoff=${relative_abundance_cutoff},
  prevalence_cutoff=${prevalence_cutoff},
  id_col="SampleName",
  normalize=T
  )

filtered_taxa = taxa_abundance$short_glommed_taxa

```

### Build Master Table
```{r}
master_table = getMasterTable(
  taxa_abundance,
  sample_data,
  'SampleName',
  'short_glommed_taxa'
)
```

### Axes Sets
```{r}
axes_sets = list(
  c(1,2),
  c(2,3),
  c(1,3)
)
```


## MDS Plots {.tabset}
### Scree Plot
```{r}
# plotNMDSScree <- function(x) 
# {
#   plot(
#     rep(1, 10), 
#     replicate(10, metaMDS(x, autotransform = F, k = 1)$stress), 
#     xlim = c(1, 10),
#     ylim = c(0, 0.30), 
#     xlab = "# of Dimensions", 
#     ylab = "Stress", 
#     main = "NMDS stress plot"
#     )
#   for (i in 1:10)
#   {
#     points(
#       rep(i + 1,10),
#       replicate(10, metaMDS(x, autotransform = F, k = i + 1)$stress)
#       )
#   }
# }
# 
# plotNMDSScree(master_table %>% select(filtered_taxa))
```


### 2D
```{r, echo=F}

for (var in variables)
{
  var_data = observational_variables[[var]]
  print(var_data)
  
  name = var_data$covariate_of_interest
  case = var_data$case
  control = var_data$control
  
  for (axes in axes_sets)
  {
    makeNMDSPlot(
      master_table %>% filter(!!as.symbol(var) %in% c(case, control)),
      features=filtered_taxa,
      axes=axes,
      color=var,
      title_extra=paste(name, case, 'vs', control),
      subtitle="${analysis_subtitle}",
      elipses=F
    )
  }

}

```

### 3D
```{r}
pc = getNMDS(master_table, filtered_taxa, axes=c(1:3))
pcdf = data.frame(pc$points) %>% rownames_to_column('SampleName')
pcmt = inner_join(pcdf, master_table, by='SampleName')

plotOrd3D(
  pcmt, 
  axes=c('MDS1', 'MDS2', 'MDS3'), 
  color="CaseString",
  title_extra="nMDS"
  )
```


## PCoA Plots {.tabset}
### Scree Plot
```{r}
pc = getPrinceComps(master_table, filtered_taxa, 'SampleName')
screeplot(pc)
```

### 2D
```{r}
for (var in variables)
{
  var_data = observational_variables[[var]]
  print(var_data)
  
  name = var_data$covariate_of_interest
  case = var_data$case
  control = var_data$control
  
  for (axes in axes_sets)
  {
    makePCOAPlot(
      master_table %>% filter(!!as.symbol(var) %in% c(case, control)),
      method='bray',
      features=filtered_taxa,
      axes=axes,
      color=var,
      title_extra=paste(name, case, 'vs', control),
      subtitle="${analysis_subtitle}",
      elipses=F
    ) 
  }

}
# + aes(x=PC1, y=PC2, color=IGA, shape=CaseString)
```

### 3D
```{r}
pc = getPrinceComps(master_table, filtered_taxa, 'SampleName')
pcdf = data.frame(pc$x) %>% rownames_to_column('SampleName')
pcmt = inner_join(pcdf, master_table, by='SampleName')

plotOrd3D(
  princomp_master_table=pcmt,
  axes=c('PC1', 'PC2', 'PC3'),
  id_col="SampleName",
  color="CaseString",
  title_extra="Bray Distance PCoA"
)
```







