---
title: "MiSeq-${miseq_run_number}"
output:
  html_notebook:
    df_print: paged
    number_sections: yes
    theme: lumen
    toc: yes
    code_folding: hide
---

<style>
  html {
    font-size: 16pt;
  }
  
  body {
    font-size: 16pt;
  }
  
  h1 {
    font-size: 2.2rem;
  }
  
  h2 {
    font-size: 2rem;
  }
  
  h3 {
    font-size: 1.8rem;
  }
  
  h4 {
    font-size: 1.4rem;
  }
  
</style>

```{r}
source('${project_metadata_file}')
analysis_type = "${analysis_type}"
clustering_level = "${clustering_level}"
tables_dir = "${tables_dir}"
use_allbac = ${use_allbac}
analysis_title = tools::toTitleCase(gsub("_", " ", analysis_type))
```

# MiSeq-${miseq_run_number} Beta Diversity {.tabset}

## Clustering Level
Data clustered at the `r toupper('${clustering_level}')` Level.

## Filtering
* **Relative Abundance Cutoff:** ${relative_abundance_cutoff}
* **Prevalence Cutoff:** ${prevalence_cutoff}
* **Min Count Cutoff:** ${min_count_cutoff}

## Setup

### Start Conda ENV
```{r}
startCondaEnv('betadiv', lib='~/R35')
```

### Load Libraries
```{r, echo=F}
library(tidyverse)
library(magrittr)
library(ggplot2)
library(ggbeeswarm)
library(openxlsx)
library(DT)
library(pheatmap)
library(kableExtra)
library(vegan)
library(plotly)

### Custom libraries that can be loaded from GitHub
source('~/utilities/analysis-utilities/general_asv_data_analysis_utilities.R')
source('~/utilities/analysis-utilities/beta_diversity_utilities.R')
```

### Set Knitr Options
```{r}
${knitr_options}
```

### Load Data
```{r}
asv_table = 
  read.delim(
    '${asv_table_file}', 
    header=T, 
    sep='\t',
    stringsAsFactors=F
    )

taxonomy_table = 
  read.delim(
    '${taxonomy_table_file}', 
    header=T, 
    sep='\t',
    stringsAsFactors=F
    ) %>%
  mutate_all(as.character())

sample_data = read.xlsx('${sample_data_file}')

if ("IGA" %in% colnames(sample_data) & use_allbac)
{
  sample_data = 
    sample_data %>%
    filter(IGA=='AllBac')

  sample_names = sample_data$SampleName
  
  asv_table = 
    asv_table %>%
    select('ASVs', ends_with('AllBac'))
}

sample_data_columns = colnames(sample_data)
sample_names = sample_data$SampleName


source(file.path(
  "${project_root}", 
  "${metadata_dir}", 
  'observational_variables.R'
  ))

```

### Set Filtering
```{r}
relative_abundance_cutoff = ${relative_abundance_cutoff}
prevalence_cutoff = ${prevalence_cutoff}
```
**relative abundance cutoff:** `r relative_abundance_cutoff`
**prevalence cutoff:** `r prevalence_cutoff`

### Variables of interest
#### Observational Variables
```{r}
key_covariates = ${key_covariates}
treatments = ${treatments}
variables = c('CaseString', key_covariates, treatments) %>% unique()
```

#### Remove Age
```{r}
variables %<>% setdiff('Age')
```

* **Treatment variables:** ${treatments}
* **Key covariates:**  ${key_covariates}
* **Variables Used:** `r variables`

### Get Filtered Taxa Abundance
```{r}
taxa_abundance = getFilteredTaxaCountsDev(
  asv_table,
  taxonomy_table,
  sample_data, 
  cluster_by="${clustering_level}",
  relative_abundance_cutoff=${relative_abundance_cutoff},
  prevalence_cutoff=${prevalence_cutoff},
  id_col="SampleName",
  normalize=T
  )

filtered_taxa = taxa_abundance$short_glommed_taxa

```

### Build Master Table
```{r}
master_table = getMasterTable(
  taxa_abundance,
  sample_data,
  'SampleName',
  'short_glommed_taxa'
)
```

### Axes Sets
```{r}
axes_sets = list(
  c(1,2),
  c(2,3),
  c(1,3)
)
```


## MDS Plots {.tabset}
### Scree Plot
```{r}
# plotNMDSScree <- function(x) 
# {
#   plot(
#     rep(1, 10), 
#     replicate(10, metaMDS(x, autotransform = F, k = 1)$stress), 
#     xlim = c(1, 10),
#     ylim = c(0, 0.30), 
#     xlab = "# of Dimensions", 
#     ylab = "Stress", 
#     main = "NMDS stress plot"
#     )
#   for (i in 1:10)
#   {
#     points(
#       rep(i + 1,10),
#       replicate(10, metaMDS(x, autotransform = F, k = i + 1)$stress)
#       )
#   }
# }
# 
# plotNMDSScree(master_table %>% select(filtered_taxa))
```


### 2D
```{r, echo=F}

for (var in variables)
{
  var_data = observational_variables[[var]]
  print(var_data)
  
  name = var_data$covariate_of_interest
  case = var_data$case
  control = var_data$control
  
  for (axes in axes_sets)
  {
    makeNMDSPlot(
      master_table,
      features=filtered_taxa,
      axes=axes,
      color=var,
      title_extra=paste(name, case, 'vs', control),
      elipses=F
    )
  }

}

```

### 3D
```{r}
pc = getNMDS(master_table, filtered_taxa, axes=c(1:3))
pcdf = data.frame(pc$points) %>% rownames_to_column('SampleName')
pcmt = inner_join(pcdf, master_table, by='SampleName')

plotOrd3D(
  pcmt, 
  axes=c('MDS1', 'MDS2', 'MDS3'), 
  color="CaseString",
  title_extra="nMDS"
  )
```


## PCoA Plots {.tabset}
### Scree Plot
```{r}
pc = getPrinceComps(master_table, filtered_taxa, 'SampleName')
screeplot(pc)
```

### 2D
```{r}
for (var in variables)
{
  var_data = observational_variables[[var]]
  print(var_data)
  
  name = var_data$covariate_of_interest
  case = var_data$case
  control = var_data$control
  
  for (axes in axes_sets)
  {
    makePCOAPlot(
      master_table,
      method='bray',
      features=filtered_taxa,
      axes=axes,
      color=var,
      title_extra=paste(name, case, 'vs', control),
      elipses=F
    ) 
  }

}
# + aes(x=PC1, y=PC2, color=IGA, shape=CaseString)
```

### 3D
```{r}
pc = getPrinceComps(master_table, filtered_taxa, 'SampleName')
pcdf = data.frame(pc$x) %>% rownames_to_column('SampleName')
pcmt = inner_join(pcdf, master_table, by='SampleName')

plotOrd3D(
  princomp_master_table=pcmt,
  axes=c('PC1', 'PC2', 'PC3'),
  id_col="SampleName",
  color="CaseString",
  title_extra="Bray Distance PCoA"
)
```







