---
title: "MiSeq-${miseq_run_number}"
output:
  html_notebook:
    df_print: paged
    number_sections: yes
    theme: lumen
    toc: yes
    code_folding: hide
---

<style>
  html {
    font-size: 16pt;
  }
  
  body {
    font-size: 16pt;
  }
  
  h1 {
    font-size: 2.2rem;
  }
  
  h2 {
    font-size: 2rem;
  }
  
  h3 {
    font-size: 1.8rem;
    
  }
  
  h4 {
    font-size: 1.4rem;
  }
  
</style>


```{r}
source('${project_metadata_file}')
analysis_type = "${analysis_type}"
clustering_level = "${clustering_level}"
tables_dir = "${tables_dir}"
analysis_title = "${analysis_title}"
```


# MiSeq-${miseq_run_number} `r analysis_title` Calculated Data Plots {.tabset}

## Clustering Level
Data clustered at the `r toupper('${clustering_level}')` Level.

## Filtering
* **Relative Abundance Cutoff:** ${relative_abundance_cutoff}
* **Prevalence Cutoff:** ${prevalence_cutoff}
* **Min Count Cutoff:** ${min_count_cutoff}


## Setup

### Start Conda ENV
```{r}
startCondaEnv('base')
```

### Load Libraries
```{r}
library(rmarkdown)
library(knitr)

library(tidyverse)
library(magrittr)
library(ggplot2)
library(ggbeeswarm)
library(openxlsx)
library(DT)
library(pheatmap)
library(kableExtra)

### Custom libraries that can be loaded from GitHub
source('~/utilities/analysis-utilities/general_asv_data_analysis_utilities.R')
source('~/utilities/amd_templates/setup/amd_project_utilities.R')
```

### Set Knitr Options
```{r}

${knitr_options}
```

### Load Data
```{r}
filename = makeDataFileName(
  'master_table.xlsx', 
  tables_dir, 
  analysis_type,
  "${miseq_project_prefix}",
  "${clustering_level}"
  )

print(filename)

all_master_table = readWorkbook(filename, sheet='Master Table')
age_lt_90_master_table = all_master_table %>% filter(Age < 90)
amd_only_master_table = all_master_table %>% filter(CaseString == 'AMD')
filtered_taxa = readWorkbook(filename, sheet="${tools::toTitleCase(clustering_level)}") %>% pull("short_glommed_taxa")

source(file.path(
  "${project_root}", 
  "${metadata_dir}", 
  'observational_variables.R'
  ))

```

### Variables of interest
#### All Observational Variables
```{r}
# print(observational_variables)
```

#### AMD Only Observational Variables
```{r}
amd_only_variables = observational_variables

amd_only_excluded_variables = c(
  'Gender',
  'CaseString',
  'Age'
)

for (varname in amd_only_excluded_variables)
{
  print(sprintf('Excluding %s', varname))
  amd_only_variables[varname] = NULL
}

```

**Excluding: ** `r amd_only_excluded_variables`


#### Calculated Experimental Variables

```{r}
calculated_exp_vars = ${calculated_exp_vars}
```
**Calculated Experimental Variables: ** `r calculated_exp_vars`


### Set Metadata Factor Levels
```{r}
all_master_table = setFactorsLevels(
  all_master_table, 
  observational_variables
  )
age_lt_90_master_table = setFactorsLevels(
  age_lt_90_master_table, 
  observational_variables
  )
amd_only_master_table = setFactorsLevels(
  amd_only_master_table, 
  observational_variables
  )
```

## Calculated Data Plots
### All Subgroups
```{r}
makeBoxAndDotplot(
  master_table=all_master_table,
  indices=calculated_exp_vars,
  aesthetics=list(
    x_var='index', 
    color_var='index', 
    facet_var='index'
    ),
  title=paste(analysis_title, "-- All Subgroups")
)
```

### Age < 90
```{r}
makeBoxAndDotplot(
  master_table=age_lt_90_master_table,
  indices=calculated_exp_vars,
  aesthetics=list(
    x_var='index', 
    color_var='index', 
    facet_var='index'
    ),
  title=paste(analysis_title, "-- Age < 90")
)
```

### AMD Only

```{r}
makeBoxAndDotplot(
  master_table=amd_only_master_table,
  indices=calculated_exp_vars,
  aesthetics=list(
    x_var='index', 
    color_var='index', 
    facet_var='index'
    ),
  title=paste(analysis_title, "--AMD Only")
)
```
