---
title: "MiSeq-${miseq_run_number}"
output:
  html_notebook:
    df_print: paged
    number_sections: yes
    theme: lumen
    toc: yes
    code_folding: hide
---

<style>
  html {
    font-size: 16pt;
  }
  
  body {
    font-size: 16pt;
  }
  
  h1 {
    font-size: 2.2rem;
  }
  
  h2 {
    font-size: 2rem;
  }
  
  h3 {
    font-size: 1.8rem;
  }
  
  h4 {
    font-size: 1.4rem;
  }
  
</style>


```{r}
source('${project_metadata_file}')
analysis_type = "${analysis_type}"
clustering_level = "${clustering_level}"
tables_dir = file.path(analysis_dir, analysis_type, clustering_level, 'tables')
analysis_title = "${analysis_title}"
```

# MiSeq-${miseq_run_number} Pearson's $\Chi^2$ Test for Independence {.tabset}
## Clustering Level
Data clustered at the `r toupper('${clustering_level}')` Level.

## Filtering
* **Relative Abundance Cutoff:** ${relative_abundance_cutoff}
* **Prevalence Cutoff:** ${prevalence_cutoff}
* **Min Count Cutoff:** ${min_count_cutoff}


## Categorical Variables Included
```{r}
test_categorical_variables = ${test_categorical_variables}
num_test_vars = length(test_categorical_variables)
reference_categorical_variables = ${reference_categorical_variables}
num_ref_vars = length(reference_categorical_variables)
```

Using the following variables:

* **Reference Variables:** `r reference_categorical_variables`
* **Test Variables:** `r test_categorical_variables`

## Setup

### Start Conda ENV
```{r}
startCondaEnv('base')
```


### Load Libraries
```{r}
library(rmarkdown)
# library(knitr)

library(tidyverse)
library(magrittr)
library(DT)
library(openxlsx)
library(ggplot2)
library(ggbeeswarm)
library(pheatmap)
library(vcd)

### Custom libraries that can be loaded from GitHub
source('~/utilities/analysis-utilities/general_asv_data_analysis_utilities.R')
source('~/utilities/amd_templates/setup/amd_project_utilities.R')
```

### Set Knitr Options
```{r}
${knitr_options}

```

### Load Data
```{r}
filename = makeDataFileName(
  'master_table.xlsx',
  "${tables_dir}",
  "${analysis_type}",
  "${miseq_project_prefix}",
  "${clustering_level}"
)

all_master_table = readWorkbook(filename, sheet='Master Table')
age_lt_90_master_table = all_master_table %>% filter(Age < 90)
amd_only_master_table = all_master_table %>% filter(CaseString == 'AMD')
filtered_taxa = readWorkbook(filename, sheet="${tools::toTitleCase(clustering_level)}") %>% pull("short_glommed_taxa")

source(file.path(
  "${project_root}", 
  "${metadata_dir}", 
  'observational_variables.R'
  ))
```

### Set Metadata Factor Levels
```{r}
all_master_table = setFactorsLevels(
  all_master_table, 
  observational_variables
  )
age_lt_90_master_table = setFactorsLevels(
  age_lt_90_master_table, 
  observational_variables
  )
amd_only_master_table = setFactorsLevels(
  amd_only_master_table, 
  observational_variables
  )
```


## All Subgroups
### Run Pearson's $\Chi^2$ Test for Independence
```{r}
all_pval_table = matrix(ncol=num_ref_vars, nrow=num_test_vars)
colnames(all_pval_table) = reference_categorical_variables
rownames(all_pval_table) = test_categorical_variables

for (ref_var in reference_categorical_variables)
{
  for (test_var in test_categorical_variables)
  {
    print(sprintf('ref_var: %s, test_var: %s', ref_var, test_var))
    
    crosstab = 
      all_master_table %>% 
      select(!!ref_var, !!test_var) %>%
      table()
    
    print(crosstab)
    
    pvalue = chisq.test(crosstab) %>% .[['p.value']]
    print(pvalue)
    
    all_pval_table[test_var, ref_var] = pvalue
    
    mosaic(crosstab, shade=T)
  } 
}
```

### Heatmap
```{r}
pheatmap(
  mat=all_pval_table,
  cluster_cols = F,
  cluster_rows = F
)
```


### Write Stats Table
```{r}
filename = makeDataFileName(
  'all_chisq.xlsx', 
  "${tables_dir}",
  "${analysis_type}",
  "${miseq_project_prefix}",
  "${clustering_level}"
  )

print(filename)

write.xlsx(
  all_pval_table,
  file=filename,
  quote=F,
  col.names=T,
  row.names=T
)
```

