---
title: "${miseq_project_prefix}: DESeq2 Normalized ${analysis_type}"
output:
  html_notebook:
    df_print: paged
    number_sections: yes
    theme: lumen
    toc: yes
    code_folding: hide
---

<style>
  html {
    font-size: 16pt;
  }
  
  body {
    font-size: 16pt;
  }
  
  h1 {
    font-size: 2.2rem;
  }
  
  h2 {
    font-size: 2rem;
  }
  
  h3 {
    font-size: 1.8rem;
  }
  
  h4 {
    font-size: 1.4rem;
  }
  
</style>


```{r}
analysis_type = "${analysis_type}"
clustering_level = "${clustering_level}"
tables_dir = "${tables_dir}"
analysis_title = paste("DESeq2 Normalized", "${analysis_title}")
```

# ${miseq_project_prefix} ${analysis_title} Linear Regression Analysis {.tabset}
**Clustering**
Data clustered at the `r toupper('${clustering_level}')` Level.

**Subgroups**
```{r, results='asis', echo=FALSE}
for (sg_filter in c(${subgroup_filters}))
{
  cat(sg_filter, '\n\n')
}
```

**Filtering**
  * **Relative Abundance cutoff:** ${relative_abundance_cutoff}
* **Prevalence Cutoff:** ${prevalence_cutoff}
* **Min Count Cutoff:** ${min_count_cutoff}
```{r}
relative_abundance_cutoff = ${relative_abundance_cutoff}
prevalence_cutoff = ${prevalence_cutoff}
min_count_cutoff = ${min_count_cutoff}
```

## Setup
### Start Conda ENV
```{r}
startCondaEnv('deseq')
```

### Load Libraries
```{r}
library(rmarkdown)
library(tidyverse)
library(magrittr)
library(ggplot2)
library(ggbeeswarm)
library(openxlsx)
# library(DT)
library(pheatmap)
# library(kableExtra)
library(DESeq2)

### Custom libraries that can be loaded from GitHub
source('https://raw.githubusercontent.com/ohsu-microbiome/utilities/master/analysis-utilities/general_asv_data_analysis_utilities.R')

```


### Set Knitr Options
```{r}
${knitr_options}
```

### Load Data
```{r}
asv_table = 
  read.delim(
    '${asv_table_file}', 
    header=T, 
    sep='\t',
    stringsAsFactors=F
    )

taxonomy_table = 
  read.delim(
    '${taxonomy_table_file}', 
    header=T, 
    sep='\t',
    stringsAsFactors=F
    ) %>%
  mutate_all(as.character())

sample_data = read.xlsx('${sample_data_file}')

sample_data_columns = colnames(sample_data)
sample_names = sample_data$SampleName

source(file.path(
  "${project_root}", 
  "${metadata_dir}", 
  'observational_variables.R'
  ))

```


### Set Filtering
```{r}
relative_abundance_cutoff=${relative_abundance_cutoff}
prevalence_cutoff=${prevalence_cutoff}
min_count_cutoff=${min_count_cutoff}
```


### Extract Subgroups
```{r}
if (!is.null(c(${subgroup_filters})))
{
  subgroup_filter_string = paste0(${subgroup_filters}, collapse=" & ")
  print(sprintf("Subgroup Filter String: %s", subgroup_filter_string))
  subgroup_filter_expression = str2expression(subgroup_filter_string)
  sample_data = sample_data %>% filter(eval(subgroup_filter_expression))
  sample_names = sample_data$SampleName

  asv_table = asv_table %>% select('ASVs', sample_names)
}

```


### Set Metadata Factor Levels
```{r}
sample_data = setFactorsLevels(sample_data, observational_variables)
# sample_data %<>%
#   mutate(AREDS=
#     ifelse(is.na(AREDS), "MISSING", AREDS) %>%
#     factor(levels=c('N', 'Y', 'MISSING'))
#   )
```

### Variables of interest
#### Observational Variables
```{r}
case = ${case} %>% setdiff('')
key_covariates = ${key_covariates} %>% setdiff('')
linear_predictors = ${linear_predictors} %>% setdiff('')
```

* **Case:** ${case}
* **Key Covariates:** ${key_covariates}
* **linear_predictors:** ${linear_predictors}

### Create p-val names for contrasts
```{r}
all_contrast_names = makeContrastNames(
  observational_variables[linear_predictors], 
  sample_data
  )
```

### Initialize regression stats table

```{r}
regression_pvals = makeRegressionStatsContrastTemplate(all_contrast_names)

regression_effect_sizes = makeRegressionStatsContrastTemplate(all_contrast_names)
```


## Create DESeq2 Normalizede Master Table

### Set Normalization Method Method
Normalization Method: ${deseq_normalization_method}

```{r}
deseq_normalization_method = "${deseq_normalization_method}"

normalization_function = list(
    vst=DESeq2::varianceStabilizingTransformation,
    rlog=DESeq2::rlog
  )[["${deseq_normalization_method}"]]
```

### Get Normalized Counts
```{r}
taxa_counts = getFilteredTaxaCounts(
  asv_table=asv_table,
  taxonomy_table=taxonomy_table,
  sample_data=sample_data,
  cluster_by="Genus",
  filter_by="taxa",
  relative_abundance_cutoff=${relative_abundance_cutoff},
  prevalence_cutoff=${prevalence_cutoff},
  min_count_cutoff=${min_count_cutoff},
  clean_taxa=T,  ### remove NAs in lowest rank
  n_max_by_mean=F,
  id_col="SampleName", ### metadata column that containes the unique sample IDs
  add_glommed_names=T,
  normalize=F
)

sample_names = sample_data$SampleName

taxa_counts %<>%
  mutate_at(sample_names, list(function(x){x = x + 1}))

filtered_taxa = taxa_counts$short_glommed_taxa

dds = DESeqDataSetFromMatrix(
  countData=
    taxa_counts %>%
    select(sample_names, short_glommed_taxa) %>%
    column_to_rownames('short_glommed_taxa'),
  colData=sample_data,
  design= ~ CaseString
)

# r = rlog(dds, fitType='local')
# v = varianceStabilizingTransformation(dds, fitType='local')

normalized_counts =
  dds %>%
  normalization_function(., fitType="local") %>%
  assay() %>%
  t() %>%
  data.frame(stringsAsFactors=F) %>%
  rownames_to_column('SampleName') %>%
  inner_join(sample_data, by="SampleName") %>% 
  select(SampleName, !!linear_predictors, !!filtered_taxa)

```

### Plot Normalized Counts
```{r}
normalized_counts %>% 
  select(SampleName, !!filtered_taxa) %>%
  gather(key='taxa', value='abundance', -SampleName) %>%
  ggplot(aes(x=taxa, y=abundance)) +
  geom_boxplot()

normalized_counts %>% 
  select(SampleName, !!filtered_taxa) %>%
  gather(key='taxa', value='abundance', -SampleName) %>%
  ggplot(aes(x=taxa, y=abundance)) +
  geom_violin()
```



## Linear Regression {.tabset}

### All

#### Do Regression
```{r}
all_pvals = makeRegressionStatsContrastTemplate(all_contrast_names)

all_effect_sizes = makeRegressionStatsContrastTemplate(all_contrast_names)


exp_vars = filtered_taxa
predictors = observational_variables[linear_predictors]
master_table = normalized_counts
contrasts = all_contrast_names
index_name_template = ''

# regression_pvals = regression_pvals_template
# regression_effect_sizes = regression_pvals_template

for (var in exp_vars)
{
  index_name = paste0(index_name_template, var)
  print(index_name)
  
  fit = doMultipleRegression(
    var,
    predictors,
    response_vars=var,
    master_table,
    transformation="${regression_transform}",
    log_regularizer=${log_regularizer}
  )
  
  pvals = getRegressionPvals(
    fit,
    contrasts,
    index_name
  )
  
  all_pvals = 
    all_pvals %>% 
    add_row(!!!pvals)

  effect_sizes = getRegressionEffectSizes(
    fit,
    contrasts,
    index_name
  )

  all_effect_sizes = 
    all_effect_sizes %>% 
    add_row(!!!effect_sizes)

}


all_adjusted_pvals = 
  all_pvals %>%
  mutate_if(is.numeric, p.adjust)
```


#### Visualize
```{r}
pheatmap(all_pvals %>% column_to_rownames('Index'), cluster_cols=F, cluster_rows=F)

pheatmap(all_effect_sizes %>% column_to_rownames('Index'), cluster_cols=F, cluster_rows=F)
```


### Age < 90

#### Do Regression
```{r}
age_lt_90_pvals = makeRegressionStatsContrastTemplate(all_contrast_names)

age_lt_90_effect_sizes = makeRegressionStatsContrastTemplate(all_contrast_names)


exp_vars = filtered_taxa
predictors = observational_variables[linear_predictors]
master_table = normalized_counts %>% filter(Age < 90)
contrasts = all_contrast_names
index_name_template = ''

# regression_pvals = regression_pvals_template
# regression_effect_sizes = regression_pvals_template

for (var in exp_vars)
{
  index_name = paste0(index_name_template, var)
  print(index_name)
  
  fit = doMultipleRegression(
    var,
    predictors,
    response_vars=var,
    master_table,
    transformation="${regression_transform}",
    log_regularizer=${log_regularizer}
  )
  
  pvals = getRegressionPvals(
    fit,
    contrasts,
    index_name
  )
  
  age_lt_90_pvals = 
    age_lt_90_pvals %>% 
    add_row(!!!pvals)

  effect_sizes = getRegressionEffectSizes(
    fit,
    contrasts,
    index_name
  )

  age_lt_90_effect_sizes = 
    age_lt_90_effect_sizes %>% 
    add_row(!!!effect_sizes)

}


age_lt_90_adjusted_pvals = 
  age_lt_90_pvals %>%
  mutate_if(is.numeric, p.adjust)
```


#### Visualize
```{r}
pheatmap(age_lt_90_pvals %>% column_to_rownames('Index'), cluster_cols=F, cluster_rows=F)

pheatmap(age_lt_90_effect_sizes %>% column_to_rownames('Index'), cluster_cols=F, cluster_rows=F)
```


### AMD Only

#### Do Regression
```{r}
amd_only_pvals = makeRegressionStatsContrastTemplate(all_contrast_names)

amd_only_effect_sizes = makeRegressionStatsContrastTemplate(all_contrast_names)


exp_vars = filtered_taxa
predictors = observational_variables[linear_predictors %>% setdiff(case)]
master_table = normalized_counts %>% filter(CaseString == "AMD")
contrasts = all_contrast_names %>% unlist()
index_name_template = ''

for (var in exp_vars)
{
  index_name = paste0(index_name_template, var)
  print(index_name)
  
  fit = doMultipleRegression(
    var,
    predictors,
    response_vars=var,
    master_table,
    transformation="${regression_transform}",
    log_regularizer=${log_regularizer}
  )
  
  pvals = getRegressionPvals(
    fit,
    contrasts,
    index_name
  )
  
  amd_only_pvals = 
    amd_only_pvals %>% 
    add_row(!!!pvals)

  effect_sizes = getRegressionEffectSizes(
    fit,
    contrasts,
    index_name
  )

  amd_only_effect_sizes = 
    amd_only_effect_sizes %>%
    add_row(!!!effect_sizes)

}

amd_only_adjusted_pvals = 
  amd_only_pvals %>%
  mutate_if(is.numeric, p.adjust)
```

#### Visualize
```{r}
pheatmap(amd_only_pvals %>% column_to_rownames('Index'), cluster_cols=F, cluster_rows=F)

pheatmap(amd_only_effect_sizes %>% column_to_rownames('Index'), cluster_cols=F, cluster_rows=F)
```

### Write Regression Stats
```{r}
wb = createWorkbook()

addWorksheet(
  wb, 
  sheetName="All Unadjusted P-Values"
  )

openxlsx::writeData(
  wb, 
  sheet="All Unadjusted P-Values", 
  all_pvals
  )

addWorksheet(
  wb, 
  sheetName="All Adjusted P-Values"
  )

openxlsx::writeData(
  wb, 
  sheet="All Adjusted P-Values", 
  all_adjusted_pvals
  )

addWorksheet(
  wb, 
  sheetName="Age < 90 Unadjusted P-Values"
  )

openxlsx::writeData(
  wb, 
  sheet="Age < 90 Unadjusted P-Values", 
  age_lt_90_pvals
  )

addWorksheet(
  wb, 
  sheetName="Age < 90 Adjusted P-Values"
  )

openxlsx::writeData(
  wb, 
  sheet="All Adjusted P-Values", 
  age_lt_90_adjusted_pvals
  )

addWorksheet(
  wb, 
  sheetName="AMD Only Unadjusted P-Values"
  )

openxlsx::writeData(
  wb, 
  sheet="AMD Only Unadjusted P-Values", 
  amd_only_pvals
  )

addWorksheet(
  wb, 
  sheetName="AMD Only Adjusted P-Values"
  )

openxlsx::writeData(
  wb, 
  sheet="AMD Only Adjusted P-Values", 
  amd_only_adjusted_pvals
  )

addWorksheet(
  wb, 
  sheetName="All Effect Sizes"
  )

openxlsx::writeData(
  wb,
  sheet="All Effect Sizes", 
  all_effect_sizes
  )

addWorksheet(
  wb, 
  sheetName="Age < 90 Effect Sizes"
  )

openxlsx::writeData(
  wb,
  sheet="Age < 90 Effect Sizes", 
  age_lt_90_effect_sizes
  )

addWorksheet(
  wb, 
  sheetName="AMD Only Effect Sizes"
  )

openxlsx::writeData(
  wb,
  sheet="All Effect Sizes", 
  amd_only_effect_sizes
  )
# addWorksheet(regression_pvals, sheetName="Adjusted P-Values")
# openxlsx::writeData(regression_pvals, sheet="Adjusted P-Values", adjusted_regression_pvals)

filename = makeDataFileName(
  'deseq_normalized_linear_regression.xlsx',
  "${tables_dir}",
  "${analysis_type}",
  "${miseq_project_prefix}",
  "${clustering_level}"
)
saveWorkbook(wb, file=filename, overwrite=T)
```
