---
title: "${miseq_project_prefix}: ${analysis_type}"
output:
  html_notebook:
    df_print: paged
    number_sections: yes
    theme: lumen
    toc: yes
    code_folding: hide
---

<style>
  html {
    font-size: 16pt;
  }
  
  body {
    font-size: 16pt;
  }
  
  h1 {
    font-size: 2.2rem;
  }
  
  h2 {
    font-size: 2rem;
  }
  
  h3 {
    font-size: 1.8rem;
  }
  
  h4 {
    font-size: 1.4rem;
  }
  
  .dataTable {
    font-size: 9pt;
  }
  
</style>

```{r}
source('${project_metadata_file}')
analysis_type = "${analysis_type}"
clustering_level = "${clustering_level}"
filter_by = "${filter_by}"
tables_dir = "${tables_dir}"
analysis_title = tools::toTitleCase(gsub("_", " ", analysis_type))
cooks_cutoff = ${cooks_cutoff}
independent_filtering = ${independent_filtering}
```

# ${miseq_project_prefix} DESeq2 Analysis {.tabset}

**Clustering**
Data clustered at the `r toupper('${clustering_level}')` Level.

**Subgroups**
```{r, results='asis', echo=FALSE}
for (sg_filter in c(${subgroup_filters}))
{
  cat(sg_filter, '\n\n')
}
```

**Filtering**
Data filtered by `r toupper('${filter_by}')`

* **Relative Abundance cutoff:** ${relative_abundance_cutoff}
* **Prevalence Cutoff:** ${prevalence_cutoff}
* **Min Count Cutoff:** ${min_count_cutoff}

```{r}
relative_abundance_cutoff = ${relative_abundance_cutoff}
prevalence_cutoff = ${prevalence_cutoff}
min_count_cutoff = ${min_count_cutoff}
```

## Setup 

### [Close]

### Start Conda ENV
```{r}
startCondaEnv('deseq', lib='~/R35')
```

### Load Libraries
```{r}
library(rmarkdown)
library(tidyverse)
library(magrittr)
library(ggplot2)
library(openxlsx)

library(kableExtra)
library(DESeq2)

### Custom libraries that can be loaded from GitHub
source('${utilities_dir}/analysis-utilities/general_asv_data_analysis_utilities.R')
source('${utilities_dir}/amd_templates/setup/amd_project_utilities.R')
source('${utilities_dir}/analysis-utilities/deseq_utilities.R')

```

### Set Knitr Options
```{r}
${knitr_options}
```

### Load Data
```{r}
asv_table = 
  read.delim(
    '${asv_table_file}', 
    header=T, 
    sep='\t',
    stringsAsFactors=F
    )

taxonomy_table = 
  read.delim(
    '${taxonomy_table_file}', 
    header=T, 
    sep='\t',
    stringsAsFactors=F
    ) %>%
  mutate_all(as.character())

sample_data = read.xlsx('${sample_data_file}')

sample_data_columns = colnames(sample_data)
sample_names = sample_data$SampleName

sample_data %<>% filter(CaseString == 'AMD')

source(file.path(
  "${project_root}", 
  "${metadata_dir}", 
  'observational_variables.R'
  ))
```

### Set Metadata Factor Levels
```{r}
sample_data = setFactorsLevels(sample_data, observational_variables)
```

### Extract Subgroups
```{r}
if (!is.null(c(${subgroup_filters})))
{
  subgroup_filter_string = paste0(${subgroup_filters}, collapse=" & ")
  print(sprintf("Subgroup Filter String: %s", subgroup_filter_string))
  subgroup_filter_expression = str2expression(subgroup_filter_string)
  sample_data = sample_data %>% filter(eval(subgroup_filter_expression))
  sample_names = sample_data$SampleName

  asv_table = asv_table %>% select('ASVs', sample_names)
}

```

### Variables of interest
#### Observational Variables
```{r}
case = ${case} %>% setdiff('')
treatments = ${treatments} %>% setdiff('')
phenotypes = ${phenotypes} %>% setdiff('')
genotypes = ${genotypes} %>% setdiff('')
key_covariates = ${key_covariates} %>% setdiff('')

linear_predictors = ${linear_predictors} %>% setdiff('')
categorical_variables = ${categorical_variables} %>% setdiff('')
amd_only_variables = linear_predictors %>% setdiff(case)
````

* **Genotype variables:** `r genotypes`
* **Phenotype variables:** `r phenotypes`
* **Treatment variables:** `r treatments`
* **Other covariates:**  `r key_covariates`

### Set Filtering
```{r}
relative_abundance_cutoff=${relative_abundance_cutoff}
prevalence_cutoff=${prevalence_cutoff}
```

## Variables Independantly
### Create Workbook
```{r}
deseq_results_wb = createWorkbook()
```

### Run and add worksheets
```{r}

lowest_rank = 'Genus'

for (varname in amd_only_variables)
{
  
  var = observational_variables[[varname]]
  
  taxa_counts = getFilteredTaxaCounts(
    asv_table=asv_table,
    taxonomy_table=taxonomy_table,
    sample_data=sample_data,
    cluster_by=lowest_rank,
    filter_by="${filter_by}",
    relative_abundance_cutoff=relative_abundance_cutoff,
    prevalence_cutoff=prevalence_cutoff,
    min_count_cutoff=0,
    clean_taxa=T,  ### remove NAs in lowest rank
    n_max_by_mean=F,
    id_col="SampleName", ### metadata column that containes the unique sample IDs
    add_glommed_names=T,
    normalize=F
  )
  
  if (var$case != "" && var$control != "")
  {
    var_sample_data =
      sample_data %>% 
      filter(!!as.name(varname) %in% c(var$case, var$control))
  } else
  {
    var_sample_data = sample_data
  }
    
  results_df = runDeseqFromTables_dev(
    asv_table,
    taxonomy_table,
    var_sample_data,
    varname,
    lowest_rank=lowest_rank,
    filter_by="${filter_by}",
    relative_abundance_cutoff=relative_abundance_cutoff,
    prevalence_cutoff=prevalence_cutoff,
    include_covariates = 'All',
    cooks_cutoff=cooks_cutoff,
    independent_filtering=independent_filtering
    )
  
    match_string = paste0(varname, '.*\\.padj')
    padj_col = grep(match_string, colnames(results_df), value=T)
  
  results_df =
    results_df %>%
    select(Taxon, !!padj_col, starts_with(varname), everything()) %>%
    arrange(!!as.name(padj_col))
  
  addWorksheet(
    wb=deseq_results_wb, 
    sheetName=varname
    )
  
  openxlsx::writeData(
    wb=deseq_results_wb, 
    sheet=varname, 
    x=results_df
    )
  
  setColWidths(
    wb=deseq_results_wb, 
    sheet=varname, 
    cols=1, 
    widths = "auto"
  )
  
  freezePane(
    deseq_results_wb, 
    sheet=varname, 
    firstCol=T,
    firstRow=T
    )
}
```

### Save Workbook
```{r}
filename = makeDataFileName(
  'amd_only_individual_variables.xlsx',
  "${tables_dir}",
  "${analysis_type}",
  "${miseq_project_prefix}",
  "rank"
)

print(filename)

saveWorkbook(deseq_results_wb, file=filename, overwrite=T)
```

## Genotypes, Phenotypes, and AREDS with Controls
### Create Workbook
```{r}
deseq_results_wb = createWorkbook()
```

### Run and add worksheets
#### Key Covariates Alone
```{r}

varname = "None"
  

taxa_counts = getFilteredTaxaCounts(
  asv_table=asv_table,
  taxonomy_table=taxonomy_table,
  sample_data=sample_data,
  cluster_by=lowest_rank,
  filter_by="${filter_by}",
  relative_abundance_cutoff=relative_abundance_cutoff,
  prevalence_cutoff=prevalence_cutoff,
  min_count_cutoff=0,
  clean_taxa=T,  ### remove NAs in lowest rank
  n_max_by_mean=F,
  id_col="SampleName", ### metadata column that containes the unique sample IDs
  add_glommed_names=T,
  normalize=F
)
  
results_df = runDeseqFromTables_dev(
  asv_table,
  taxonomy_table,
  sample_data,
  variables=key_covariates,
  lowest_rank=lowest_rank,
  filter_by="${filter_by}",
  relative_abundance_cutoff=relative_abundance_cutoff,
  prevalence_cutoff=prevalence_cutoff,
  include_covariates = 'All',
  cooks_cutoff=cooks_cutoff,
  independent_filtering=independent_filtering
  )

addWorksheet(
  wb=deseq_results_wb, 
  sheetName=varname
  )

openxlsx::writeData(
  wb=deseq_results_wb, 
  sheet=varname, 
  x=results_df
  )
# 
# setColWidths(
#   wb=deseq_results_wb, 
#   sheet=varname, 
#   cols=1, 
#   widths = "auto"
# )
# 
# freezePane(
#   deseq_results_wb, 
#   sheet=varname, 
#   firstCol=T,
#   firstRow=T
#   )
```

#### Target Variables with covarariates
```{r}
lowest_rank = 'Genus'

for (varname in c(genotypes, phenotypes, treatments))
{
  
  vars = observational_variables[c(varname, key_covariates)]
  
  risk = vars[[varname]]$case
  non_risk = vars[[varname]]$control
  
  taxa_counts = getFilteredTaxaCounts(
    asv_table=asv_table,
    taxonomy_table=taxonomy_table,
    sample_data=sample_data,
    cluster_by=lowest_rank,
    filter_by="${filter_by}",
    relative_abundance_cutoff=relative_abundance_cutoff,
    prevalence_cutoff=prevalence_cutoff,
    min_count_cutoff=0,
    clean_taxa=T,  ### remove NAs in lowest rank
    n_max_by_mean=F,
    id_col="SampleName", ### metadata column that containes the unique sample IDs
    add_glommed_names=T,
    normalize=F
  )
    
  results_df = runDeseqFromTables_dev(
    asv_table,
    taxonomy_table,
    sample_data %>% filter(!!as.name(varname) %in% c(risk, non_risk)),
    vars %>% names(),
    lowest_rank=lowest_rank,
    filter_by="${filter_by}",
    relative_abundance_cutoff=relative_abundance_cutoff,
    prevalence_cutoff=prevalence_cutoff,
    include_covariates = 'All',
    cooks_cutoff=cooks_cutoff,
    independent_filtering=independent_filtering
    )
  
  match_string = paste0(varname, '.*\\.padj')
  padj_col = grep(match_string, colnames(results_df), value=T)
  
  results_df =
    results_df %>%
    select(Taxon, !!padj_col, starts_with(varname), everything()) %>%
    arrange(!!as.name(padj_col))
    
  addWorksheet(
    wb=deseq_results_wb, 
    sheetName=varname
    )
  
  openxlsx::writeData(
    wb=deseq_results_wb, 
    sheet=varname, 
    x=results_df
    )
  # 
  # setColWidths(
  #   wb=deseq_results_wb, 
  #   sheet=varname, 
  #   cols=1, 
  #   widths = "auto"
  # )
  # 
  # freezePane(
  #   deseq_results_wb, 
  #   sheet=varname, 
  #   firstCol=T,
  #   firstRow=T
  #   )

}
```

### Save Workbook
```{r}
filename = makeDataFileName(
  'amd_only_variables_with_controls.xlsx',
  "${tables_dir}",
  "${analysis_type}",
  "${miseq_project_prefix}",
  "rank"
)

print(filename)

saveWorkbook(deseq_results_wb, file=filename, overwrite=T)
```
