---
title: "MiSeq-${miseq_run_number}"
output:
  html_notebook:
    df_print: paged
    number_sections: yes
    theme: lumen
    toc: yes
    code_folding: hide
---

<style>
  html {
    font-size: 16pt;
  }
  
  body {
    font-size: 16pt;
  }
  
  h1 {
    font-size: 2.2rem;
  }
  
  h2 {
    font-size: 2rem;
  }
  
  h3 {
    font-size: 1.8rem;
  }
  
  h4 {
    font-size: 1.4rem;
  }
  
  .dataTable {
    font-size: 9pt;
  }
  
</style>

```{r}
source('${project_metadata_file}')
analysis_type = "${analysis_type}"
clustering_level = "${clustering_level}"
tables_dir = "${tables_dir}"
use_allbac = ${use_allbac}
analysis_title = tools::toTitleCase(gsub("_", " ", analysis_type))
```

# MiSeq-${miseq_run_number} DESeq2 Analysis {.tabset}

## Clustering
Data clustered at the `r toupper("${clustering_level}")` level.

## Filtering
* **Relative Abundance Cutoff:** ${relative_abundance_cutoff}
* **Prevalence Cutoff:** ${prevalence_cutoff}
* **Min Count Cutoff:** ${min_count_cutoff}


## Setup 

### [Close]

### Start Conda ENV
```{r}
startCondaEnv('deseq', lib='~/R35')
```

### Load Libraries
```{r}
library(rmarkdown)
library(tidyverse)
library(magrittr)
library(ggplot2)
library(openxlsx)
library(DT)
library(kableExtra)
library(DESeq2)

### Custom libraries that can be loaded from GitHub
source('~/utilities/analysis-utilities/general_asv_data_analysis_utilities.R')
source('~/utilities/analysis-utilities/deseq_utilities.R')
source('~/utilities/amd_templates/setup/amd_project_utilities.R')
```

### Set Knitr Options
```{r}
${knitr_options}
```

### Load Data
```{r}
asv_table = 
  read.delim(
    '${asv_table_file}', 
    header=T, 
    sep='\t',
    stringsAsFactors=F
    )

taxonomy_table = 
  read.delim(
    '${taxonomy_table_file}', 
    header=T, 
    sep='\t',
    stringsAsFactors=F
    ) %>%
  mutate_all(as.character())

sample_data = 
  read.delim(
    '${sample_data_file}', 
    header=T, 
    sep='\t',
    stringsAsFactors=F
    ) %>%
  filter(SampleID!='Undetermined')

if ("IGA" %in% colnames(sample_data) & use_allbac)
{
  sample_data = 
    sample_data %>%
    filter(IGA=='AllBac')

  sample_names = sample_data$SampleName
  
  asv_table = 
    asv_table %>%
    select('ASVs', ends_with('AllBac'))
}

sample_data_columns = colnames(sample_data)
sample_names = sample_data$SampleName

source(file.path(
  "${project_root}", 
  "${metadata_dir}", 
  'observational_variables.R'
  ))

```

## Covariates

### All and Age < 90 Observational variables
```{r}
# all_variable_names = setdiff(
#   observational_variables %>% names(),
#   c('ARMS2rs10490924', 'CFHrs1061170', 'CFHrs10737680', 'SKIV2rs429608', 'Pct_IgA_Pos')
# )
# 
# ### All samples from stool--no swabs
# 
# all_variable_names = c('CaseString', 'Age', 'Gender', 'AREDS')
# 
# all_variables = observational_variables[all_variable_names]
# print(all_variable_names)
# 
# age_lt_90_variables = all_variables

### Deal with NA values for AREDS
sample_data %<>%
  mutate(AREDS=
    ifelse(is.na(AREDS), "MISSING", AREDS) %>%
    factor(levels=c('N', 'Y', 'MISSING'))
  )

all_covaraites = ${all_covariates}
```

### Set Filtering
```{r}
relative_abundance_cutoff=${relative_abundance_cutoff}
prevalence_cutoff=${relative_abundance_cutoff}
```

## DESeq2 {.tabset}
```{r}
all_ranks = c('Phylum', 'Class', 'Order', 'Family', 'Genus')
```

### All Subgroups
#### Create Workbook
```{r}
deseq_results_wb = createWorkbook()
```

#### Run and add worksheets
```{r}
for (lowest_rank in all_ranks)
{
  results_df = runDeseqFromTables_dev(
    asv_table,
    taxonomy_table,
    sample_data,
    observational_variables[all_covaraites],
    lowest_rank=lowest_rank,
    relative_abundance_cutoff=relative_abundance_cutoff,
    prevalence_cutoff=prevalence_cutoff,
    include_covariates = 'CaseString'
    )
  
  addWorksheet(
    wb=deseq_results_wb, 
    sheetName=lowest_rank
    )
  
  openxlsx::writeData(
    wb=deseq_results_wb, 
    sheet=lowest_rank, 
    x=results_df
    )
}
```

#### Save Workbook
```{r}
filename = makeDataFileName(
  'all_deseq_results.xlsx',
  "${tables_dir}",
  "${analysis_type}",
  "${miseq_project_prefix}",
  "rank"
)

print(filename)

saveWorkbook(deseq_results_wb, file=filename, overwrite=T)
```


### Age < 90
```{r}
age_lt_90_covariates = ${age_lt_90_covariates}
```

#### Create Workbook
```{r}
deseq_results_wb = createWorkbook()
```

#### Run and add worksheets
```{r}
for (lowest_rank in all_ranks)
{
  results_df = runDeseqFromTables_dev(
    asv_table,
    taxonomy_table,
    sample_data %>% filter(Age < 90),
    observational_variables[age_lt_90_covariates],
    lowest_rank=lowest_rank,
    relative_abundance_cutoff=relative_abundance_cutoff,
    prevalence_cutoff=prevalence_cutoff,
    include_covariates = 'CaseString'
    )
  
  addWorksheet(
    wb=deseq_results_wb, 
    sheetName=lowest_rank
    )
  
  openxlsx::writeData(
    wb=deseq_results_wb, 
    sheet=lowest_rank, 
    x=results_df
    )
}
```

#### Save Workbook
```{r}
filename = makeDataFileName(
  'age_lt_90_deseq_results.xlsx',
  "${tables_dir}",
  "${analysis_type}",
  "${miseq_project_prefix}",
  "rank"
)

print(filename)

saveWorkbook(deseq_results_wb, file=filename, overwrite=T)
```
