---
title: "${miseq_project_prefix}: ${analysis_type}"
output:
  html_notebook:
    df_print: paged
    number_sections: yes
    theme: lumen
    toc: yes
    code_folding: hide
---

<style>
  html {
    font-size: 16pt;
  }
  
  body {
    font-size: 16pt;
  }
  
  h1 {
    font-size: 2.2rem;
  }
  
  h2 {
    font-size: 2rem;
  }
  
  h3 {
    font-size: 1.8rem;
  }
  
  h4 {
    font-size: 1.4rem;
  }
  
</style>

```{r}
source('${project_metadata_file}')
analysis_type = "${analysis_type}"
clustering_level = "${clustering_level}"
tables_dir = "${tables_dir}"
analysis_type_title = tools::toTitleCase(gsub("_", " ", analysis_type))
analysis_subtitle = "${analysis_subtitle}"
```

# ${miseq_project_prefix} `r analysis_type` Differential Abundance Plots{.tabset}
Regression coefficients as effect size.

**Clustering**
Data clustered at the `r toupper('${clustering_level}')` Level.

**Subgroups**
```{r, results='asis', echo=FALSE}
for (sg_filter in c(${subgroup_filters}))
{
  cat(sg_filter, '\n\n')
}
```

**Filtering**
* **Relative Abundance cutoff:** ${relative_abundance_cutoff}
* **Prevalence Cutoff:** ${prevalence_cutoff}
* **Min Count Cutoff:** ${min_count_cutoff}

```{r}
relative_abundance_cutoff = ${relative_abundance_cutoff}
prevalence_cutoff = ${prevalence_cutoff}
min_count_cutoff = ${min_count_cutoff}
```

## Setup {.tabset}

### [Close]

### Start Conda ENV
```{r}
startCondaEnv('deseq', lib="~/R35")
```

### Load Libraries
```{r}
library(rmarkdown)
library(tidyverse)
library(magrittr)
library(ggplot2)
library(ggbeeswarm)
library(openxlsx)
library(DT)
library(pheatmap)
library(kableExtra)
# library(vegan)
library(nlme)
library(tools)

### Custom libraries that can be loaded from GitHub
source('~/utilities/analysis-utilities/general_asv_data_analysis_utilities.R')
source('~/utilities/analysis-utilities/deseq_utilities.R')
source('~/utilities/amd_templates/setup/amd_project_utilities.R')
```

### Set Knitr Options
```{r}
${knitr_options}
```

### Load Data
```{r}
sample_data = read.xlsx('${sample_data_file}')

source(file.path(
  "${project_root}", 
  "${metadata_dir}", 
  'observational_variables.R'
  ))

```


### Variables of interest
#### Observational Variables
```{r}
case = ${case} %>% setdiff('')
treatments = ${treatments} %>% setdiff('')
phenotypes = ${phenotypes} %>% setdiff('')
genotypes = ${genotypes} %>% setdiff('')
key_covariates = ${key_covariates} %>% setdiff('')

linear_predictors = ${linear_predictors} %>% setdiff('')
categorical_variables = ${categorical_variables} %>% setdiff('')
amd_only_variables = linear_predictors %>% setdiff(case)
```


* **Genotype variables:** `r genotypes`
* **Phenotype variables:** `r phenotypes`
* **Treatment variables:** `r treatments`
* **Other covariates:**  `r key_covariates`

### Set Ranks
```{r}
all_ranks = c('Phylum', 'Class', 'Order', 'Family', 'Genus')
```

## Plots {.tabset}
### [ ] Close

### All Subgroups -- Case/Control
```{r, fig.width=10}
wb_filename = makeDataFileName(
  'all_subgroups_case_vs_control.xlsx',
  "${tables_dir}",
  "${analysis_type}",
  "${miseq_project_prefix}",
  "rank"
)

print(wb_filename)

for (rank in all_ranks)
{
  # rank = 'Genus'
  data = 
    readWorkbook(wb_filename, sheet=rank) 
  
  sig_data = data %>% filter(CaseString_AMD_vs_Control.padj <= 0.2)
  
  if (dim(sig_data)[1] < 1)
  {
    next
  }
  
  plotEffectSize2(
    data=sig_data,
    effect_size_col="CaseString_AMD_vs_Control.L2FC",
    effect_size_legend_title="Abundance in AMD",
    feature_col="Taxon",
    alpha_col='CaseString_AMD_vs_Control.padj',
    alpha_legend_title="Adjusted P-Values",
    plot_title=paste0(rank,": AMD/Control: Log2 Fold Change"),
    subtitle="${analysis_subtitle}"
  ) %>% print()
  
}
```

### All Subgroups -- With Covariates
```{r, fig.width=10}
wb_filename = makeDataFileName(
  'all_subgroups_with_covariates.xlsx',
  "${tables_dir}",
  "${analysis_type}",
  "${miseq_project_prefix}",
  "rank"
)

print(wb_filename)

for (rank in all_ranks)
{
  # rank = 'Genus'
  data = 
    readWorkbook(wb_filename, sheet=rank) 
  
  sig_data = data %>% filter(CaseString_AMD_vs_Control.padj <= 0.2)
  
  if (dim(sig_data)[1] < 1)
  {
    next
  }
  
  plotEffectSize2(
    data=sig_data,
    effect_size_col="CaseString_AMD_vs_Control.L2FC",
    effect_size_legend_title="Abundance in AMD",
    feature_col="Taxon",
    alpha_col='CaseString_AMD_vs_Control.padj',
    alpha_legend_title="Adjusted P-Values",
    plot_title=paste0(rank,": AMD/Control + Gender + Tissue_code: Log2 Fold Change"),
    subtitle="${analysis_subtitle}"
  ) %>% print()
  
}
```


### Age < 90 -- Case/Control
```{r, fig.width=10}
wb_filename = makeDataFileName(
  'age_lt_90_case_vs_control.xlsx',
  "${tables_dir}",
  "${analysis_type}",
  "${miseq_project_prefix}",
  "rank"
)

print(wb_filename)

for (rank in all_ranks)
{
  # rank = 'Genus'
  data = 
    readWorkbook(wb_filename, sheet=rank) 
  
  sig_data = data %>% filter(CaseString_AMD_vs_Control.padj <= 0.2)
  
  if (dim(sig_data)[1] < 1)
  {
    next
  }
  
  plotEffectSize2(
    data=sig_data,
    effect_size_col="CaseString_AMD_vs_Control.L2FC",
    effect_size_legend_title="Abundance in AMD",
    feature_col="Taxon",
    alpha_col='CaseString_AMD_vs_Control.padj',
    alpha_legend_title="Adjusted P-Values",
    plot_title=paste0(rank,": AMD/Control: Log2 Fold Change"),
    subtitle="${analysis_subtitle}"
  ) %>% print()
  
}
```

### Age < 90 -- With Covariates
```{r, fig.width=10}
wb_filename = makeDataFileName(
  'age_lt_90_with_covariates.xlsx',
  "${tables_dir}",
  "${analysis_type}",
  "${miseq_project_prefix}",
  "rank"
)

print(wb_filename)

for (rank in all_ranks)
{
  # rank = 'Genus'
  data = 
    readWorkbook(wb_filename, sheet=rank) 
  
  sig_data = data %>% filter(CaseString_AMD_vs_Control.padj <= 0.2)
  
  if (dim(sig_data)[1] < 1)
  {
    next
  }
  
  plotEffectSize2(
    data=sig_data,
    effect_size_col="CaseString_AMD_vs_Control.L2FC",
    effect_size_legend_title="Abundance in AMD",
    feature_col="Taxon",
    alpha_col='CaseString_AMD_vs_Control.padj',
    alpha_legend_title="Adjusted P-Values",
    plot_title=paste0(rank,": AMD/Control + Gender + Tissue_code: Log2 Fold Change"),
    subtitle="${analysis_subtitle}"
  ) %>% print()
  
}
```


### AMD Only -- individual variables
```{r, fig.width=10}
wb_filename = makeDataFileName(
  'amd_only_individual_variables.xlsx',
  "${tables_dir}",
  "${analysis_type}",
  "${miseq_project_prefix}",
  "rank"
)

print(wb_filename)

lowest_rank = 'Genus'

for (varname in amd_only_variables)
{
  # rank = 'Genus'
  data = 
    readWorkbook(wb_filename, sheet=varname) 
  
  match_string = paste0(varname, '.*\\.padj')
  padj_col = grep(match_string, colnames(data), value=T)
  
  match_string = paste0(varname, '.*\\.L2FC')
  effect_size_col = grep(match_string, colnames(data), value=T)
  
  sig_data = data %>% filter(!!as.name(padj_col) <= 0.2)
  
  if (dim(sig_data)[1] < 1)
  {
    next
  }
  
  plotEffectSize2(
    data=sig_data,
    effect_size_col=effect_size_col,
    effect_size_legend_title=paste0("Abundance in ", varname),
    feature_col="Taxon",
    alpha_col=padj_col,
    alpha_legend_title="Adjusted P-Values",
    plot_title=paste0(varname,": Log2 Fold Change"),
    subtitle="${analysis_subtitle}"
  ) %>% print()
  
}
```


### AMD Only -- Genotypes, Phenotypes, and AREDS with Controls
```{r, fig.width=10}
wb_filename = makeDataFileName(
  'amd_only_variables_with_controls.xlsx',
  "${tables_dir}",
  "${analysis_type}",
  "${miseq_project_prefix}",
  "rank"
)

print(wb_filename)

lowest_rank = 'Genus'

for (varname in c(genotypes, phenotypes, treatments))
{
  # rank = 'Genus'
  data = 
    readWorkbook(wb_filename, sheet=varname) 
  
  match_string = paste0(varname, '.*\\.padj')
  padj_col = grep(match_string, colnames(data), value=T)
  
  match_string = paste0(varname, '.*\\.L2FC')
  effect_size_col = grep(match_string, colnames(data), value=T)
  
  sig_data = data %>% filter(!!as.name(padj_col) <= 0.2)
  
  if (dim(sig_data)[1] < 1)
  {
    next
  }
  
  plotEffectSize2(
    data=sig_data,
    effect_size_col=effect_size_col,
    effect_size_legend_title=paste0("Abundance in ", varname),
    feature_col="Taxon",
    alpha_col=padj_col,
    alpha_legend_title="Adjusted P-Values",
    plot_title=paste0(varname, " + Gender + Tissue_code: Log2 Fold Change"),
    subtitle="${analysis_subtitle}"
  ) %>% print()
  
}
```


