---
title: "MiSeq-239: DESeq2 Publication Plots"
output:
  html_notebook:
    df_print: paged
    number_sections: yes
    theme: lumen
    toc: yes
    code_folding: hide
---

<style>
  html {
    font-size: 16pt;
  }
  
  body {
    font-size: 16pt;
  }
  
  h1 {
    font-size: 2.2rem;
  }
  
  h2 {
    font-size: 2rem;
  }
  
  h3 {
    font-size: 1.8rem;
  }
  
  h4 {
    font-size: 1.4rem;
  }
  
</style>

```{r}
source('${project_metadata_file}')
clustering_level = "${clustering_level}"
tables_dir = "${tables_dir}"
analysis_type = "${analysis_type}"
analysis_type_title = tools::toTitleCase(gsub("_", " ", analysis_type))
rank = "${clustering_level}"
```

# MiSeq-239 `r analysis_type` Differential Abundance Plots{.tabset}
Regression coefficients as effect size.

**Clustering**
Data clustered at the `r toupper('rank')` Level.

**Subgroups**
```{r, results='asis', echo=FALSE}
for (sg_filter in c(${subgroup_filters}))
{
  cat(sg_filter, '\n\n')
}
```

**Filtering**
* **Relative Abundance cutoff:** 0.002
* **Prevalence Cutoff:** 0.1
* **Min Count Cutoff:** 0

```{r}
relative_abundance_cutoff = 0.002
prevalence_cutoff = 0.1
min_count_cutoff = 0
```

## Setup {.tabset}

### [Close]

### Start Conda ENV
```{r}
startCondaEnv('deseq', lib="~/R35")
```

**Filtering**
* **Relative Abundance cutoff:** ${relative_abundance_cutoff}
* **Prevalence Cutoff:** ${prevalence_cutoff}
* **Min Count Cutoff:** ${min_count_cutoff}

```{r}
relative_abundance_cutoff = ${relative_abundance_cutoff}
prevalence_cutoff = ${prevalence_cutoff}
min_count_cutoff = ${min_count_cutoff}
```

## Setup {.tabset}

### [Close]

### Start Conda ENV
```{r}
startCondaEnv('deseq', lib="~/R35")
```

### Load Libraries
```{r}
# library(rmarkdown)
library(tidyverse)
library(magrittr)
library(ggplot2)
library(openxlsx)
# library(vegan)
library(tools)

### Custom libraries that can be loaded from GitHub
source('${utilities_dir}/analysis-utilities/general_asv_data_analysis_utilities.R')
source('${utilities_dir}/amd_templates/setup/amd_project_utilities.R')
source('${utilities_dir}/analysis-utilities/deseq_utilities.R')

```

### Set Knitr Options
```{r}
${knitr_options}
```

### Load Data
```{r}
sample_data = read.xlsx('${sample_data_file}')

source(file.path(
  "${project_root}", 
  "${metadata_dir}", 
  'observational_variables.R'
  ))

```
### Variables of interest

#### Observational Variables
```{r}
case = ${case} %>% setdiff('')
treatments = ${treatments} %>% setdiff('')
phenotypes = ${phenotypes} %>% setdiff('')
genotypes = ${genotypes} %>% setdiff('')
key_covariates = ${key_covariates} %>% setdiff('')

linear_predictors = ${linear_predictors} %>% setdiff('')
categorical_variables = ${categorical_variables} %>% setdiff('')
amd_only_variables = linear_predictors %>% setdiff(case)
```

* **Case:** `r case`
* **Key Covariates:** `r key_covariates`
* **Categorical Variables:**  `r categorical_variables`


## Plots {.tabset}
### [ ] Close

### getFigHeight
```{r}
getFigHeight = function(num_features)
{
  # fig_height = num_features*exp(0.004*num_features)
  fig_height = 3 + 0.3*num_features
  return(fig_height)
}
```


### All Subgroups -- Case/Control
```{r, fig.width=10}
wb_filename = makeDataFileName(
  'all_subgroups_case_vs_control.xlsx',
  "${tables_dir}",
  "${analysis_type}",
  "${miseq_project_prefix}",
  "${clustering_level}"
)

print(wb_filename)
exists(wb_filename)

data = 
  readWorkbook(wb_filename, sheet="Genus") 

data %>% 
  pull(CaseString_AMD_vs_Control.padj) %>%
  min() %>% 
  signif(., 3) %>%
  sprintf("Min Pval: %s", .) %>%
  print()

# sig_data = data %>% filter(CaseString_AMD_vs_Control.padj <= 0.2)
sig_data = 
  data %>% 
  arrange(CaseString_AMD_vs_Control.padj) %>% 
  filter(CaseString_AMD_vs_Control.padj < 0.1)

num_features = dim(sig_data)[1]
fig_height = getFigHeight(num_features)
print(fig_height)
```


```{r, fig.height=getFigHeight(num_features)}
if (dim(sig_data)[1] > 0)
{
  makePublicationEffectSizePlotBW(
    data=sig_data,
    var_data=observational_variables[['CaseString']],
    effect_size_col="CaseString_AMD_vs_Control.L2FC",
    feature_col="Taxon",
    feature_axis_label="Significant Taxa",
    alpha_col='CaseString_AMD_vs_Control.padj',
  )
}

```

### All Subgroups -- With Covariates
```{r, fig.width=10}
wb_filename = makeDataFileName(
  'all_subgroups_with_covariates.xlsx',
  "${tables_dir}",
  "${analysis_type}",
  "${miseq_project_prefix}",
  "${clustering_level}"
)

print(wb_filename)

data = 
  readWorkbook(wb_filename, sheet="Genus") 

data %>% 
  pull(CaseString_AMD_vs_Control.padj) %>%
  min() %>% 
  signif(., 3) %>%
  sprintf("Min Pval: %s", .) %>%
  print()

# sig_data = data %>% filter(CaseString_AMD_vs_Control.padj <= 0.2)
sig_data = 
  data %>% 
  arrange(CaseString_AMD_vs_Control.padj) %>% 
  filter(CaseString_AMD_vs_Control.padj < 0.1)

num_features = dim(sig_data)[1]
fig_height = getFigHeight(num_features)
print(fig_height)
```


```{r, fig.height=getFigHeight(num_features)}
if (dim(sig_data)[1] > 0)
{
  makePublicationEffectSizePlotBW(
    data=sig_data,
    var_data=observational_variables[['CaseString']],
    effect_size_col="CaseString_AMD_vs_Control.L2FC",
    feature_col="Taxon",
    feature_axis_label="Significant Taxa",
    alpha_col='CaseString_AMD_vs_Control.padj'
  )
}

  
```


### Age < 90 -- Case/Control
```{r, fig.width=10}
wb_filename = makeDataFileName(
  'age_lt_90_case_vs_control.xlsx',
  "${tables_dir}",
  "${analysis_type}",
  "${miseq_project_prefix}",
  "${clustering_level}"
)

print(wb_filename)

data = 
  readWorkbook(wb_filename, sheet="Genus") 

data %>% 
  pull(CaseString_AMD_vs_Control.padj) %>%
  min() %>% 
  signif(., 3) %>%
  sprintf("Min Pval: %s", .) %>%
  print()

# sig_data = data %>% filter(CaseString_AMD_vs_Control.padj <= 0.2)
sig_data = 
  data %>% 
  arrange(CaseString_AMD_vs_Control.padj) %>% 
  filter(CaseString_AMD_vs_Control.padj < 0.1)

num_features = dim(sig_data)[1]
fig_height = getFigHeight(num_features)
print(fig_height)
```


```{r, fig.height=getFigHeight(num_features)}
if (dim(sig_data)[1] > 0)
{
  makePublicationEffectSizePlotBW(
    data=sig_data,
    var_data=observational_variables[['CaseString']],
    effect_size_col="CaseString_AMD_vs_Control.L2FC",
    feature_col="Taxon",
    feature_axis_label="Significant Taxa",
    alpha_col='CaseString_AMD_vs_Control.padj'
  )
}

```

### Age < 90 -- With Covariates
```{r, fig.width=10}
wb_filename = makeDataFileName(
  'age_lt_90_with_covariates.xlsx',
  "${tables_dir}",
  "${analysis_type}",
  "${miseq_project_prefix}",
  "${clustering_level}"
)

print(wb_filename)

data = 
  readWorkbook(wb_filename, sheet="Genus") 

data %>% 
  pull(CaseString_AMD_vs_Control.padj) %>%
  min() %>% 
  signif(., 3) %>%
  sprintf("Min Pval: %s", .) %>%
  print()

# sig_data = data %>% filter(CaseString_AMD_vs_Control.padj <= 0.2)
sig_data = 
  data %>% 
  arrange(CaseString_AMD_vs_Control.padj) %>% 
  filter(CaseString_AMD_vs_Control.padj < 0.1)

num_features = dim(sig_data)[1]
fig_height = getFigHeight(num_features)
print(fig_height)
```


```{r, fig.height=getFigHeight(num_features)}
if(dim(sig_data)[1] > 0)
{ 
  makePublicationEffectSizePlotBW(
    data=sig_data,
    var_data=observational_variables[['CaseString']],
    effect_size_col="CaseString_AMD_vs_Control.L2FC",
    feature_col="Taxon",
    feature_axis_label="Significant Taxa",
    alpha_col='CaseString_AMD_vs_Control.padj'
  )
}

```


### AMD Only -- individual variables
```{r, fig.width=10}
wb_filename = makeDataFileName(
  'amd_only_individual_variables.xlsx',
  "${tables_dir}",
  "${analysis_type}",
  "${miseq_project_prefix}",
  "${clustering_level}"
)

print(wb_filename)
file.exists(wb_filename)

lowest_rank = 'Genus'
```

#### Gender
```{r}
varname = "Gender"
if (varname %in% getSheetNames(wb_filename))
{
  data = 
  readWorkbook(wb_filename, sheet=varname) 
  
  match_string = paste0(varname, '.*\\.padj')
  padj_col = grep(match_string, colnames(data), value=T)
  
  match_string = paste0(varname, '.*\\.L2FC')
  effect_size_col = grep(match_string, colnames(data), value=T)
  
  sig_data = data %>% filter(!!as.name(padj_col) <= 0.1)
  
  num_features = dim(sig_data)[1]
  print(num_features)
} else
{
  num_features = 0
}
```


```{r, fig.height=getFigHeight(num_features)}
if (num_features > 0)
{
  makePublicationEffectSizePlotBW(
    data=sig_data,
    var_data=observational_variables[[varname]],
    effect_size_col=effect_size_col,
    feature_col="Taxon",
    feature_axis_label="Significant Taxa",
    alpha_col=padj_col
  )
}
```

#### Tissue Code
```{r}
varname = "Tissue_code"
if (varname %in% getSheetNames(wb_filename))
{
  data = 
    readWorkbook(wb_filename, sheet=varname) 
  
  match_string = paste0(varname, '.*\\.padj')
  padj_col = grep(match_string, colnames(data), value=T)
  
  match_string = paste0(varname, '.*\\.L2FC')
  effect_size_col = grep(match_string, colnames(data), value=T)
  
  sig_data = data %>% filter(!!as.name(padj_col) <= 0.1)
  
  num_features = dim(sig_data)[1]
  print(num_features)
} else
{
  num_features=0
}
```


```{r, fig.height=getFigHeight(num_features)}
if (num_features > 0)
{
  makePublicationEffectSizePlotBW(
    data=sig_data,
    var_data=observational_variables[[varname]],
    effect_size_col=effect_size_col,
    feature_col="Taxon",
    feature_axis_label="Significant Taxa",
    alpha_col=padj_col
  )
}
```


#### GA_No_CNV_Either_Eye
```{r}
varname = "GA_No_CNV_Either_Eye"
if (varname %in% getSheetNames(wb_filename))
{
  data = 
  readWorkbook(wb_filename, sheet=varname) 
  
  match_string = paste0(varname, '.*\\.padj')
  padj_col = grep(match_string, colnames(data), value=T)
  
  match_string = paste0(varname, '.*\\.L2FC')
  effect_size_col = grep(match_string, colnames(data), value=T)
  
  sig_data = data %>% filter(!!as.name(padj_col) <= 0.1)
  
  num_features = dim(sig_data)[1]
  print(num_features)
} else
{
  num_features = 0
}
```


```{r, fig.height=getFigHeight(num_features)}
if (num_features > 0)
{
  makePublicationEffectSizePlotBW(
    data=sig_data,
    var_data=observational_variables[[varname]],
    effect_size_col=effect_size_col,
    feature_col="Taxon",
    feature_axis_label="Significant Taxa",
    alpha_col=padj_col
  )
}
```


#### CNV_Either_Eye
```{r}
varname = "CNV_Either_Eye"
if (varname %in% getSheetNames(wb_filename))
{
  data = 
  readWorkbook(wb_filename, sheet=varname) 
  
  match_string = paste0(varname, '.*\\.padj')
  padj_col = grep(match_string, colnames(data), value=T)
  
  match_string = paste0(varname, '.*\\.L2FC')
  effect_size_col = grep(match_string, colnames(data), value=T)
  
  sig_data = data %>% filter(!!as.name(padj_col) <= 0.1)
  
  num_features = dim(sig_data)[1]
  print(num_features)
} else
{
  num_features = 0
}
```


```{r, fig.height=getFigHeight(num_features)}
if (num_features > 0)
{
  makePublicationEffectSizePlotBW(
    data=sig_data,
    var_data=observational_variables[[varname]],
    effect_size_col=effect_size_col,
    feature_col="Taxon",
    feature_axis_label="Significant Taxa",
    alpha_col=padj_col
  )
}
```


#### CFHrs10737680
```{r}
varname = "CFHrs10737680"
if (varname %in% getSheetNames(wb_filename))
{
  data = 
  readWorkbook(wb_filename, sheet=varname) 
  
  match_string = paste0(varname, '.*\\.padj')
  padj_col = grep(match_string, colnames(data), value=T)
  
  match_string = paste0(varname, '.*\\.L2FC')
  effect_size_col = grep(match_string, colnames(data), value=T)
  
  sig_data = data %>% filter(!!as.name(padj_col) <= 0.1)
  
  num_features = dim(sig_data)[1]
  print(num_features)
} else
{
  num_features = 0
}
```


```{r, fig.height=getFigHeight(num_features)}
if (num_features > 0)
{
  makePublicationEffectSizePlotBW(
    data=sig_data,
    var_data=observational_variables[[varname]],
    effect_size_col=effect_size_col,
    feature_col="Taxon",
    feature_axis_label="Significant Taxa",
    alpha_col=padj_col,
    levels_in_label=T
  )
}

```

#### CFHrs1061170
```{r}
if (varname %in% getSheetNames(wb_filename))
{
  data = 
  readWorkbook(wb_filename, sheet=varname) 
  
  match_string = paste0(varname, '.*\\.padj')
  padj_col = grep(match_string, colnames(data), value=T)
  
  match_string = paste0(varname, '.*\\.L2FC')
  effect_size_col = grep(match_string, colnames(data), value=T)
  
  sig_data = data %>% filter(!!as.name(padj_col) <= 0.1)
  
  num_features = dim(sig_data)[1]
  print(num_features)
} else
{
  num_features = 0
}
```


```{r, fig.height=getFigHeight(num_features)}
if (num_features > 0)
{
  makePublicationEffectSizePlotBW(
    data=sig_data,
    var_data=observational_variables[[varname]],
    effect_size_col=effect_size_col,
    feature_col="Taxon",
    feature_axis_label="Significant Taxa",
    alpha_col=padj_col,
    levels_in_label=T
  )
}
```


#### ARMS2rs10490924
```{r}
varname = "ARMS2rs10490924"
if (varname %in% getSheetNames(wb_filename))
{
  data = 
  readWorkbook(wb_filename, sheet=varname) 
  
  match_string = paste0(varname, '.*\\.padj')
  padj_col = grep(match_string, colnames(data), value=T)
  
  match_string = paste0(varname, '.*\\.L2FC')
  effect_size_col = grep(match_string, colnames(data), value=T)
  
  sig_data = data %>% filter(!!as.name(padj_col) <= 0.1)
  
  num_features = dim(sig_data)[1]
  print(num_features)
} else
{
  num_features = 0
}
```


```{r, fig.height=getFigHeight(num_features)}
if (num_features > 0)
{
  makePublicationEffectSizePlotBW(
    data=sig_data,
    var_data=observational_variables[[varname]],
    effect_size_col=effect_size_col,
    feature_col="Taxon",
    feature_axis_label="Significant Taxa",
    alpha_col=padj_col,
    levels_in_label=T
  )
}
```


#### SKIV2Lrs429608
```{r}
varname = "SKIV2Lrs429608"
if (varname %in% getSheetNames(wb_filename))
{
  data = 
  readWorkbook(wb_filename, sheet=varname) 
  
  match_string = paste0(varname, '.*\\.padj')
  padj_col = grep(match_string, colnames(data), value=T)
  
  match_string = paste0(varname, '.*\\.L2FC')
  effect_size_col = grep(match_string, colnames(data), value=T)
  
  sig_data = data %>% filter(!!as.name(padj_col) <= 0.1)
  
  num_features = dim(sig_data)[1]
  print(num_features)
} else
{
  num_features = 0
}
```


```{r, fig.height=getFigHeight(num_features)}
if (num_features > 0)
{
  makePublicationEffectSizePlotBW(
    data=sig_data,
    var_data=observational_variables[[varname]],
    effect_size_col=effect_size_col,
    feature_col="Taxon",
    feature_axis_label="Significant Taxa",
    alpha_col=padj_col,
    levels_in_label=T
  )
}
```


#### AREDS
```{r}
varname = "AREDS"
if (varname %in% getSheetNames(wb_filename))
{
  data = 
  readWorkbook(wb_filename, sheet=varname) 
  
  match_string = paste0(varname, '.*\\.padj')
  padj_col = grep(match_string, colnames(data), value=T)
  
  match_string = paste0(varname, '.*\\.L2FC')
  effect_size_col = grep(match_string, colnames(data), value=T)
  
  sig_data = data %>% filter(!!as.name(padj_col) <= 0.1)
  
  num_features = dim(sig_data)[1]
  print(num_features)
} else
{
  num_features = 0
}
```


```{r, fig.height=getFigHeight(num_features)}
if (num_features > 0)
{
  makePublicationEffectSizePlotBW(
    data=sig_data,
    var_data=observational_variables[[varname]],
    effect_size_col=effect_size_col,
    feature_col="Taxon",
    feature_axis_label="Significant Taxa",
    alpha_col=padj_col
  )
}
```


### AMD Only -- Genotypes, Phenotypes, and AREDS with Controls
```{r, fig.width=10}
wb_filename = makeDataFileName(
  'amd_only_variables_with_controls.xlsx',
  "${tables_dir}",
  "${analysis_type}",
  "${miseq_project_prefix}",
  "${clustering_level}"
)

print(wb_filename)
file.exists(wb_filename)

lowest_rank = 'Genus'
```


#### Gender (+ tissue_code)
```{r}
varname = "Gender"

if (varname %in% getSheetNames(wb_filename))
{
  data = 
  readWorkbook(wb_filename, sheet=varname) 
  
  match_string = paste0(varname, '.*\\.padj')
  padj_col = grep(match_string, colnames(data), value=T)
  
  match_string = paste0(varname, '.*\\.L2FC')
  effect_size_col = grep(match_string, colnames(data), value=T)
  
  sig_data = data %>% filter(!!as.name(padj_col) <= 0.1)
  
  num_features = dim(sig_data)[1]
  print(num_features)
} else
{
  num_features = 0
}
```

```{r, fig.height=getFigHeight(num_features)}
if (num_features > 0)
{
  makePublicationEffectSizePlotBW(
    data=sig_data,
    var_data=observational_variables[[varname]],
    effect_size_col=effect_size_col,
    feature_col="Taxon",
    feature_axis_label="Significant Taxa",
    alpha_col=padj_col
  )
}
```


#### Tissue_code (+ Gender)
```{r}
varname = "Tissue_code"

if (varname %in% getSheetNames(wb_filename))
{
  data = 
  readWorkbook(wb_filename, sheet=varname) 
  
  match_string = paste0(varname, '.*\\.padj')
  padj_col = grep(match_string, colnames(data), value=T)
  
  match_string = paste0(varname, '.*\\.L2FC')
  effect_size_col = grep(match_string, colnames(data), value=T)
  
  sig_data = data %>% filter(!!as.name(padj_col) <= 0.1)
  
  num_features = dim(sig_data)[1]
  print(num_features)
} else
{
  num_features = 0
}
```


```{r, fig.height=getFigHeight(num_features)}
if (num_features > 0)
{
  makePublicationEffectSizePlotBW(
    data=sig_data,
    var_data=observational_variables[[varname]],
    effect_size_col=effect_size_col,
    feature_col="Taxon",
    feature_axis_label="Significant Taxa",
    alpha_col=padj_col
  )
}

```



#### GA_No_CNV_Either_Eye
```{r}
varname = "GA_No_CNV_Either_Eye"
if (varname %in% getSheetNames(wb_filename))
{
  data = 
  readWorkbook(wb_filename, sheet=varname) 
  
  match_string = paste0(varname, '.*\\.padj')
  padj_col = grep(match_string, colnames(data), value=T)
  
  match_string = paste0(varname, '.*\\.L2FC')
  effect_size_col = grep(match_string, colnames(data), value=T)
  
  sig_data = data %>% filter(!!as.name(padj_col) <= 0.1)
  
  num_features = dim(sig_data)[1]
  print(num_features)
} else
{
  num_features = 0
}
```


```{r, fig.height=getFigHeight(num_features)}
if (num_features > 0)
{
  makePublicationEffectSizePlotBW(
    data=sig_data,
    var_data=observational_variables[[varname]],
    effect_size_col=effect_size_col,
    feature_col="Taxon",
    feature_axis_label="Significant Taxa",
    alpha_col=padj_col
  )
}
```


#### CNV_Either_Eye
```{r}
varname = "CNV_Either_Eye"
if (varname %in% getSheetNames(wb_filename))
{
  data = 
  readWorkbook(wb_filename, sheet=varname) 
  
  match_string = paste0(varname, '.*\\.padj')
  padj_col = grep(match_string, colnames(data), value=T)
  
  match_string = paste0(varname, '.*\\.L2FC')
  effect_size_col = grep(match_string, colnames(data), value=T)
  
  sig_data = data %>% filter(!!as.name(padj_col) <= 0.1)
  
  num_features = dim(sig_data)[1]
  print(num_features)
} else
{
  num_features = 0
}
```


```{r, fig.height=getFigHeight(num_features)}
if (num_features > 0)
{
  makePublicationEffectSizePlotBW(
    data=sig_data,
    var_data=observational_variables[[varname]],
    effect_size_col=effect_size_col,
    feature_col="Taxon",
    feature_axis_label="Significant Taxa",
    alpha_col=padj_col
  )
}
```


#### CFHrs10737680
```{r}
varname = "CFHrs10737680"
if (varname %in% getSheetNames(wb_filename))
{
  data = 
  readWorkbook(wb_filename, sheet=varname) 
  
  match_string = paste0(varname, '.*\\.padj')
  padj_col = grep(match_string, colnames(data), value=T)
  
  match_string = paste0(varname, '.*\\.L2FC')
  effect_size_col = grep(match_string, colnames(data), value=T)
  
  sig_data = data %>% filter(!!as.name(padj_col) <= 0.1)
  
  num_features = dim(sig_data)[1]
  print(num_features)
} else
{
  num_features = 0
}
```


```{r, fig.height=getFigHeight(num_features)}
if (num_features > 0)
{
  makePublicationEffectSizePlotBW(
    data=sig_data,
    var_data=observational_variables[[varname]],
    effect_size_col=effect_size_col,
    feature_col="Taxon",
    feature_axis_label="Significant Taxa",
    alpha_col=padj_col,
    levels_in_label=T
  )
}
```

#### CFHrs1061170
```{r}
varname = "CFHrs1061170"
if (varname %in% getSheetNames(wb_filename))
{
  data = 
  readWorkbook(wb_filename, sheet=varname) 
  
  match_string = paste0(varname, '.*\\.padj')
  padj_col = grep(match_string, colnames(data), value=T)
  
  match_string = paste0(varname, '.*\\.L2FC')
  effect_size_col = grep(match_string, colnames(data), value=T)
  
  sig_data = data %>% filter(!!as.name(padj_col) <= 0.1)
  
  num_features = dim(sig_data)[1]
  print(num_features)
} else
{
  num_features = 0
}
```


```{r, fig.height=getFigHeight(num_features)}
if (num_features > 0)
{
  makePublicationEffectSizePlotBW(
    data=sig_data,
    var_data=observational_variables[[varname]],
    effect_size_col=effect_size_col,
    feature_col="Taxon",
    feature_axis_label="Significant Taxa",
    alpha_col=padj_col,
    levels_in_label=T
  )
}
```


#### ARMS2rs10490924
```{r}
varname = "ARMS2rs10490924"
if (varname %in% getSheetNames(wb_filename))
{
  data = 
  readWorkbook(wb_filename, sheet=varname) 
  
  match_string = paste0(varname, '.*\\.padj')
  padj_col = grep(match_string, colnames(data), value=T)
  
  match_string = paste0(varname, '.*\\.L2FC')
  effect_size_col = grep(match_string, colnames(data), value=T)
  
  sig_data = data %>% filter(!!as.name(padj_col) <= 0.1)
  
  num_features = dim(sig_data)[1]
  print(num_features)
} else
{
  num_features = 0
}
```


```{r, fig.height=getFigHeight(num_features)}
if (num_features > 0)
{
  makePublicationEffectSizePlotBW(
    data=sig_data,
    var_data=observational_variables[[varname]],
    effect_size_col=effect_size_col,
    feature_col="Taxon",
    feature_axis_label="Significant Taxa",
    alpha_col=padj_col,
    levels_in_label=T
  )
}
```


#### SKIV2Lrs429608
```{r}
varname = "SKIV2Lrs429608"
if (varname %in% getSheetNames(wb_filename))
{
  data = 
  readWorkbook(wb_filename, sheet=varname) 
  
  match_string = paste0(varname, '.*\\.padj')
  padj_col = grep(match_string, colnames(data), value=T)
  
  match_string = paste0(varname, '.*\\.L2FC')
  effect_size_col = grep(match_string, colnames(data), value=T)
  
  sig_data = data %>% filter(!!as.name(padj_col) <= 0.1)
  
  num_features = dim(sig_data)[1]
  print(num_features)
} else
{
  num_features = 0
}
```


```{r, fig.height=getFigHeight(num_features)}
if (num_features > 0)
{
  makePublicationEffectSizePlotBW(
    data=sig_data,
    var_data=observational_variables[[varname]],
    effect_size_col=effect_size_col,
    feature_col="Taxon",
    feature_axis_label="Significant Taxa",
    alpha_col=padj_col,
    levels_in_label=T
  )
}
```


#### AREDS
```{r}
varname = "AREDS"
if (varname %in% getSheetNames(wb_filename))
{
  data = 
  readWorkbook(wb_filename, sheet=varname) 
  
  match_string = paste0(varname, '.*\\.padj')
  padj_col = grep(match_string, colnames(data), value=T)
  
  match_string = paste0(varname, '.*\\.L2FC')
  effect_size_col = grep(match_string, colnames(data), value=T)
  
  sig_data = data %>% filter(!!as.name(padj_col) <= 0.1)
  
  num_features = dim(sig_data)[1]
  print(num_features)
} else
{
  num_features = 0
}
```


```{r, fig.height=getFigHeight(num_features)}
if (num_features > 0)
{
  makePublicationEffectSizePlotBW(
    data=sig_data,
    var_data=observational_variables[[varname]],
    effect_size_col=effect_size_col,
    feature_col="Taxon",
    feature_axis_label="Significant Taxa",
    alpha_col=padj_col
  )
}
```

