---
title: "MiSeq-${miseq_run_number}"
output:
  html_notebook:
    df_print: paged
    number_sections: yes
    theme: lumen
    toc: yes
    code_folding: hide
---

<style>
  html {
    font-size: 16pt;
  }
  
  body {
    font-size: 16pt;
  }
  
  h1 {
    font-size: 2.2rem;
  }
  
  h2 {
    font-size: 2rem;
  }
  
  h3 {
    font-size: 1.8rem;
  }
  
  h4 {
    font-size: 1.4rem;
  }
  
</style>


```{r}
source('${project_metadata_file}')
analysis_type = "${analysis_type}"
clustering_level = "${clustering_level}"
tables_dir = "${tables_dir}"
analysis_title = "${analysis_title}"
```

# MiSeq-${miseq_run_number} `r analysis_title` Logistic Regression Analysis {.tabset}
## Clustering Level
Data clustered at the `r toupper('${clustering_level}')` Level.

## Filtering
* **Relative Abundance Cutoff:** ${relative_abundance_cutoff}
* **Prevalence Cutoff:** ${prevalence_cutoff}
* **Min Count Cutoff:** ${min_count_cutoff}

## Setup
### Start Conda ENV
```{r}
startCondaEnv('regression')
```


### Load Libraries
```{r}
library(rmarkdown)
library(tidyverse)
library(magrittr)
library(ggplot2)
library(ggbeeswarm)
library(openxlsx)
library(DT)
library(pheatmap)
library(kableExtra)
library(vegan)
library(nlme)

### Custom libraries that can be loaded from GitHub
source('~/utilities/analysis-utilities/general_asv_data_analysis_utilities.R')
source('~/utilities/amd_templates/setup/amd_project_utilities.R')
```

### Set Knitr Options
```{r}
${knitr_options}

```

### Load Data
```{r}
filename = makeDataFileName(
  'master_table.xlsx',
  "${tables_dir}",
  "${analysis_type}",
  "${miseq_project_prefix}",
  "${clustering_level}"
)

print(filename)

feature_name = "${feature_name}"

all_master_table = readWorkbook(filename, sheet=paste(feature_name, 'Master Table'))
age_lt_90_master_table = all_master_table %>% filter(Age < 90)
amd_only_master_table = all_master_table %>% filter(CaseString == 'AMD')

feature_list = readWorkbook(filename, sheet=feature_name) %>% pull(feature_name)

source(file.path(
  "${project_root}", 
  "${metadata_dir}", 
  'observational_variables.R'
  ))
```

## Variables

### Set Metadata Factor Levels
```{r}
all_master_table = setFactorsLevels(
  all_master_table, 
  observational_variables
  )

age_lt_90_master_table = setFactorsLevels(
  age_lt_90_master_table, 
  observational_variables
  )
```

#### Observational Variables

```{r}
all_variables = observational_variables[${logistic_regression_predictors['All']}]
age_lt_90_variables = observational_variables[${logistic_regression_predictors['Age_lt_90']}]
```

* **All Subgroup Predictros:** ${logistic_regression_predictors['All']}
* **Age < 90 Predictors:** ${logistic_regression_predictors['Age_lt_90']}

### Raw Variables
```{r}
raw_exp_vars = ${raw_exp_vars}
print(raw_exp_vars)
```

### calculated_exp_vars
```{r}
calculated_exp_vars = ${calculated_exp_vars}
print(calculated_exp_vars)
```

### Filter Covariates
Determine which covariates have sufficient samples with each value or of perfect separation.
```{r}
# num_samples_threshold = 5
# 
# all_variables = filterVarsBySampleThreshold2(
#   master_table=all_master_table,
#   threshold=num_samples_threshold,
#   variable_list=all_variables
#   )
# 
# age_lt_90_variables = filterVarsBySampleThreshold2(
#   master_table=age_lt_90_master_table %>% filter(Age<90),
#   threshold=num_samples_threshold,
#   variable_list=age_lt_90_variables
#   )

### Remaining Covariates
# `r #all_variables %>% names()`
# `r #age_lt_90_variables %>% names()`
```

## All Subgroups

#### Calculate Stats
```{r}
covariates = all_variables

logistic_regression_stats = getLogisticRegressionStats(
  base_predictor_vars=covariates,
  additional_predictors=c(raw_exp_vars, calculated_exp_vars),
  logistic_response_var='CaseString',
  master_table=all_master_table,
  stats=c('pvalue', 'effect_size', 'std_error'),
  response_var_name='Index'
)

contrasts = makeContrastNames(covariates, all_master_table)

all_logistic_regression_stats = data.frame(Main=character())

for (contrast in c('Main', contrasts))
{
  print(contrast)
  
  pval_name = paste0(contrast, '.pval')
  odds_name = paste0(contrast, '.odds')
  print(pval_name)
  print(odds_name)
  all_logistic_regression_stats %<>% 
    mutate(
      !!pval_name := numeric(),
      !!odds_name := numeric()
    )
}


for(var in names(logistic_regression_stats))
{
  print(sprintf('var=%s', var))
  
  var_data = logistic_regression_stats[[var]]
  print('var_data')
  print(var_data)
  
  stats_list = list()
  stats_list[['Main']] = var
      
  pval_name = paste0('Main', '.pval')
  odds_name = paste0('Main', '.odds')
  
  stats_list[[pval_name]] = 
    var_data %>%
    filter(contrast == !!var) %>%
    pull(pvalue)
  stats_list[[odds_name]] = 
    var_data %>% 
    filter(contrast == !!var) %>%
    pull(effect_size) %>%
    exp()
  
  for (contrast in contrasts)
  {
    print(sprintf('contrast=%s', contrast))
    
    pval_name = paste0(contrast, '.pval')
    odds_name = paste0(contrast, '.odds')
    
    stats_list[[pval_name]] = 
      var_data %>%
      filter(contrast == !!contrast) %>%
      pull(pvalue)
    stats_list[[odds_name]] = 
      var_data %>% 
      filter(contrast == !!contrast) %>%
      pull(effect_size) %>%
      exp()
    
    print('stats_list')
    print(stats_list)
  }
    
  all_logistic_regression_stats %<>% add_row(!!!stats_list)
}

```

### Write Tables
```{r}
all_wb = createWorkbook()

addWorksheet(
  all_wb, 
  sheetName="Logistic Regression"
  )

openxlsx::writeData(
  all_wb, 
  sheet="Logistic Regression", 
  all_logistic_regression_stats
  )

addWorksheet(
  all_wb, 
  sheetName="Main Variables"
  )

openxlsx::writeData(
  all_wb,
  sheet="Main Variables",
  c(raw_exp_vars, calculated_exp_vars)
  )

filename = makeDataFileName(
  'all_logistic_regression.xlsx',
  "${tables_dir}",
  "${analysis_type}",
  "${miseq_project_prefix}",
  "${clustering_level}"
)

saveWorkbook(all_wb, file=filename, overwrite=T)
```

## Age < 90

#### Calculate Stats
```{r}

covariates = age_lt_90_variables

logistic_regression_stats = getLogisticRegressionStats(
  base_predictor_vars=covariates,
  additional_predictors=c(raw_exp_vars, calculated_exp_vars),
  logistic_response_var='CaseString',
  master_table=age_lt_90_master_table,
  stats=c('pvalue', 'effect_size', 'std_error'),
  response_var_name='Index'
)

contrasts = makeContrastNames(covariates, all_master_table)

age_lt_90_logistic_regression_stats = data.frame(Main=character())

for (contrast in c('Main', contrasts))
{
  print(contrast)
  
  pval_name = paste0(contrast, '.pval')
  odds_name = paste0(contrast, '.odds')
  print(pval_name)
  print(odds_name)
  age_lt_90_logistic_regression_stats %<>% 
    mutate(
      !!pval_name := numeric(),
      !!odds_name := numeric()
    )
}


for(var in names(logistic_regression_stats))
{
  print(sprintf('var=%s', var))
  
  var_data = logistic_regression_stats[[var]]
  print('var_data')
  print(var_data)
  
  stats_list = list()
  stats_list[['Main']] = var
      
  pval_name = paste0('Main', '.pval')
  odds_name = paste0('Main', '.odds')
  
  stats_list[[pval_name]] = 
    var_data %>%
    filter(contrast == !!var) %>%
    pull(pvalue)
  stats_list[[odds_name]] = 
    var_data %>% 
    filter(contrast == !!var) %>%
    pull(effect_size) %>%
    exp()
  
  for (contrast in contrasts)
  {
    print(sprintf('contrast=%s', contrast))
    
    pval_name = paste0(contrast, '.pval')
    odds_name = paste0(contrast, '.odds')
    
    stats_list[[pval_name]] = 
      var_data %>%
      filter(contrast == !!contrast) %>%
      pull(pvalue)
    stats_list[[odds_name]] = 
      var_data %>% 
      filter(contrast == !!contrast) %>%
      pull(effect_size) %>%
      exp()
    
    print('stats_list')
    print(stats_list)
  }
    
  age_lt_90_logistic_regression_stats %<>% add_row(!!!stats_list)
}

```


### Write Tables
```{r}
age_lt_90_wb = createWorkbook()

addWorksheet(
  age_lt_90_wb, 
  sheetName="Logistic Regression"
  )

openxlsx::writeData(
  age_lt_90_wb, 
  sheet="Logistic Regression", 
  age_lt_90_logistic_regression_stats
  )

addWorksheet(
  age_lt_90_wb, 
  sheetName="Main Variables"
  )

openxlsx::writeData(
  age_lt_90_wb,
  sheet="Main Variables",
  c(raw_exp_vars, calculated_exp_vars)
  )

filename = makeDataFileName(
  'age_lt_90_logistic_regression.xlsx',
  "${tables_dir}",
  "${analysis_type}",
  "${miseq_project_prefix}",
  "${clustering_level}"
)

saveWorkbook(age_lt_90_wb, file=filename, overwrite=T)
```

