---
title: "MiSeq=${miseq_project_prefix}: ${analysis_type}"
output:
  html_notebook:
    df_print: paged
    number_sections: yes
    theme: lumen
    toc: yes
    code_folding: hide
---

<style>
  html {
    font-size: 16pt;
  }
  
  body {
    font-size: 16pt;
  }
  
  h1 {
    font-size: 2.2rem;
  }
  
  h2 {
    font-size: 2rem;
  }
  
  h3 {
    font-size: 1.8rem;
    
  }
  
  h4 {
    font-size: 1.4rem;
  }
  
</style>


```{r}
source('${project_metadata_file}')
analysis_type = "${analysis_type}"
clustering_level = "${clustering_level}"
tables_dir = "${tables_dir}"
analysis_title = "${analysis_title}"
```

# ${miseq_project_prefix} `r analysis_title` Two Group Plots {.tabset}
Mann-Whitney Two Group Testing

**Clustering**
Data clustered at the `r toupper('${clustering_level}')` Level.

**Subgroups**
```{r, results='asis', echo=FALSE}
for (sg_filter in c(${subgroup_filters}))
{
  cat(sg_filter, '\n\n')
}
```

**Filtering**
* **Relative Abundance cutoff:** ${relative_abundance_cutoff}
* **Prevalence Cutoff:** ${prevalence_cutoff}
* **Min Count Cutoff:** ${min_count_cutoff}

```{r}
relative_abundance_cutoff = ${relative_abundance_cutoff}
prevalence_cutoff = ${prevalence_cutoff}
min_count_cutoff = ${min_count_cutoff}
```

## Setup

### Start Conda ENV
```{r}
startCondaEnv('igaseq')
```


### Load Libraries
```{r}
library(rmarkdown)
#library(knitr)

library(tidyverse)
library(magrittr)
library(DT)
library(openxlsx)
library(ggplot2)
library(ggbeeswarm)

### Custom libraries that can be loaded from GitHub
source('https://raw.githubusercontent.com/ohsu-microbiome/utilities/master/analysis-utilities/general_asv_data_analysis_utilities.R')
source('https://raw.githubusercontent.com/ohsu-microbiome/utilities/master/analysis-utilities/igaseq_utilities.R')

```

### Set Knitr Options
```{r}
${knitr_options}

```

### Load Data
```{r}
filename = makeDataFileName(
  'master_tables.xlsx', 
  "${tables_dir}",
  "${analysis_type}",
  "${miseq_project_prefix}",
  "${clustering_level}"
  )
print(sprintf("master file path: %s", filename))

short_glommed_taxa = readWorkbook(filename, "Taxa", colNames=F) %>% pull(1)
ici_score_master_table = readWorkbook(filename, sheet='Ici Score') 
log_ici_score_master_table = readWorkbook(filename, sheet='log Ici Score') 
iga_index_master_table = readWorkbook(filename, sheet='Iga Index')

source(file.path(
  "${project_root}", 
  "${metadata_dir}", 
  'observational_variables.R'
  ))
```

### Variables of interest
#### All Observational Variables
```{r}
# print(observational_variables)
```

### Set Metadata Factor Levels
```{r}
iga_index_master_table = setFactorsLevels(
  iga_index_master_table, 
  observational_variables
  )
ici_score_master_table = setFactorsLevels(
  ici_score_master_table, 
  observational_variables
  )
log_ici_score_master_table = setFactorsLevels(
  log_ici_score_master_table, 
  observational_variables
  )
```

## ICI Score 

### Setup data with pvals
```{r}
ici_score_data = ici_score_master_table %>%
  select(short_glommed_taxa, CaseString, SubjectID)


pvals = sapply(short_glommed_taxa, function(taxon)
{
  formula = paste(taxon, '~ CaseString') %>% as.formula()
  data = ici_score_data
  test = wilcox.test(formula, data)
  # test = t.test(formula, data)
  return(test$p.value)
}, USE.NAMES=T)


ici_score_with_pvals = 
  data.frame(
    short_glommed_taxa=names(pvals), 
    Pval=pvals, 
    Padj=p.adjust(pvals, method='BH'),
    stringsAsFactors=T) %>%
  inner_join(
    ici_score_data %>%
      column_to_rownames('SubjectID') %>%
      select(short_glommed_taxa) %>%
      t() %>%
      data.frame(stringsAsFactors=F) %>%
      rownames_to_column('short_glommed_taxa'),
    by='short_glommed_taxa'
  )

```

### Dotplots with pvals
```{r}
data_for_pval_plots = 
  ici_score_with_pvals %>%
  ### Filter by pval
  filter(Pval <= 0.2) %>%
  arrange(Pval)

for (rowname in data_for_pval_plots$short_glommed_taxa)
{
  # row = data_for_pval_plots[rowname,]
  row = data_for_pval_plots %>% filter(short_glommed_taxa == rowname)
  print(rowname)
  # print(row)
  
  makePvalPlot(
    data_with_metadata=ici_score_master_table,
    data_name='ICI Score', ### for title
    data_col=row$short_glommed_taxa, ### e.g. taxon
    variable_data=observational_variables$CaseString,
    p_value=list(type='Undjusted', value=row$Pval)
  )
}

```
### Index Barplot
```{r, fig.width=12, fig.height=10}
  makeIndexBarPlot(
    data_with_pvals=ici_score_with_pvals,
    sample_metadata=ici_score_master_table,
    variable_data=observational_variables$CaseString,
    data_name='ICI Score',
    pval_colname='Pval',
    pval_name='Undjusted P-Values'
    ) # + scale_fill_gradient2(low='#000000', mid='#222222', high='#FFFFFF')
```

## Log ICI Score 

### Setup data with pvals
```{r}
log_ici_score_data = log_ici_score_master_table %>%
  select(short_glommed_taxa, CaseString, SubjectID)


pvals = sapply(short_glommed_taxa, function(taxon)
{
  formula = paste(taxon, '~ CaseString') %>% as.formula()
  data = log_ici_score_data
  test = wilcox.test(formula, data)
  # test = t.test(formula, data)
  return(test$p.value)
}, USE.NAMES=T)


log_ici_score_with_pvals = 
  data.frame(
    short_glommed_taxa=names(pvals), 
    Pval=pvals, 
    Padj=p.adjust(pvals, method='BH'),
    stringsAsFactors=T) %>%
  inner_join(
    log_ici_score_data %>%
      column_to_rownames('SubjectID') %>%
      select(short_glommed_taxa) %>%
      t() %>%
      data.frame(stringsAsFactors=F) %>%
      rownames_to_column('short_glommed_taxa'),
    by='short_glommed_taxa'
  )

```

### Dotplots with pvals
```{r}
data_for_pval_plots = 
  log_ici_score_with_pvals %>%
  ### Filter by pval
  filter(Pval <= 0.2) %>%
  arrange(Pval)

for (rowname in data_for_pval_plots$short_glommed_taxa)
{
  # row = data_for_pval_plots[rowname,]
  row = data_for_pval_plots %>% filter(short_glommed_taxa == rowname)
  print(rowname)
  # print(row)
  
  makePvalPlot(
    data_with_metadata=log_ici_score_master_table,
    data_name='Log ICI Score', ### for title
    data_col=row$short_glommed_taxa, ### e.g. taxon
    variable_data=observational_variables$CaseString,
    p_value=list(type='Undjusted', value=row$Pval)
  )
}

```
### Index Barplot
```{r, fig.width=12, fig.height=10}
  makeIndexBarPlot(
    data_with_pvals=log_ici_score_with_pvals,
    sample_metadata=log_ici_score_master_table,
    variable_data=observational_variables$CaseString,
    data_name='Log ICI Score',
    pval_colname='Pval',
    pval_name='Undjusted P-Values'
    ) # + scale_fill_gradient2(low='#000000', mid='#222222', high='#FFFFFF')
```


## IgA Index

### Setup data with pvals
```{r}
iga_index_data = iga_index_master_table %>%
  select(short_glommed_taxa, CaseString, SubjectID)


pvals = sapply(short_glommed_taxa, function(taxon)
{
  formula = paste(taxon, '~ CaseString') %>% as.formula()
  data = ici_score_data
  test = wilcox.test(formula, data)
  # test = t.test(formula, data)
  return(test$p.value)
}, USE.NAMES=T)


iga_index_with_pvals = 
  data.frame(
    short_glommed_taxa=names(pvals), 
    Pval=pvals, 
    Padj=p.adjust(pvals, method='BH'),
    stringsAsFactors=T) %>%
  inner_join(
    iga_index_data %>%
      column_to_rownames('SubjectID') %>%
      select(short_glommed_taxa) %>%
      t() %>%
      data.frame(stringsAsFactors=F) %>%
      rownames_to_column('short_glommed_taxa'),
    by='short_glommed_taxa'
  )

```

### Dotplots with pvals
```{r}
data_for_pval_plots = 
  ici_score_with_pvals %>%
  ### Filter by pval
  filter(Pval <= 0.2) %>%
  arrange(Pval)

for (rowname in data_for_pval_plots$short_glommed_taxa)
{
  # row = data_for_pval_plots[rowname,]
  row = data_for_pval_plots %>% filter(short_glommed_taxa == rowname)
  print(rowname)
  # print(row)
  
  makePvalPlot(
    data_with_metadata=iga_index_master_table,
    data_name='IgA Index', ### for title
    data_col=row$short_glommed_taxa, ### e.g. taxon
    variable_data=observational_variables$CaseString,
    p_value=list(type='Undjusted', value=row$Pval)
  )
}

```

### Index Barplot
```{r, fig.width=12, fig.height=10}
  makeIndexBarPlot(
    data_with_pvals=iga_index_with_pvals,
    sample_metadata=iga_index_master_table,
    variable_data=observational_variables$CaseString,
    data_name='IgA Index',
    pval_colname='Pval',
    pval_name='Undjusted P-Values'
    ) # + scale_fill_gradient2(low='#000000', mid='#222222', high='#FFFFFF')
```


## Log ICI Score 

### Setup data with pvals
```{r}
log_ici_score_data = log_ici_score_master_table %>%
  select(short_glommed_taxa, CaseString, SubjectID)


pvals = sapply(short_glommed_taxa, function(taxon)
{
  formula = paste(taxon, '~ CaseString') %>% as.formula()
  data = log_ici_score_data
  test = wilcox.test(formula, data)
  # test = t.test(formula, data)
  return(test$p.value)
}, USE.NAMES=T)


log_ici_score_with_pvals = 
  data.frame(
    short_glommed_taxa=names(pvals), 
    Pval=pvals, 
    Padj=p.adjust(pvals, method='BH'),
    stringsAsFactors=T) %>%
  inner_join(
    log_ici_score_data %>%
      column_to_rownames('SubjectID') %>%
      select(short_glommed_taxa) %>%
      t() %>%
      data.frame(stringsAsFactors=F) %>%
      rownames_to_column('short_glommed_taxa'),
    by='short_glommed_taxa'
  )

```

### Dotplots with pvals
```{r}
data_for_pval_plots = 
  log_ici_score_with_pvals %>%
  ### Filter by pval
  filter(Pval <= 0.2) %>%
  arrange(Pval)

for (rowname in data_for_pval_plots$short_glommed_taxa)
{
  # row = data_for_pval_plots[rowname,]
  row = data_for_pval_plots %>% filter(short_glommed_taxa == rowname)
  print(rowname)
  # print(row)
  
  makePvalPlot(
    data_with_metadata=log_ici_score_master_table,
    data_name='ICI Score', ### for title
    data_col=row$short_glommed_taxa, ### e.g. taxon
    variable_data=observational_variables$CaseString,
    p_value=list(type='Undjusted', value=row$Pval)
  )
}

```
### Index Barplot
```{r, fig.width=12, fig.height=10}
  makeIndexBarPlot(
    data_with_pvals=log_ici_score_with_pvals,
    sample_metadata=log_ici_score_master_table,
    variable_data=observational_variables$CaseString,
    data_name='ICI Score',
    pval_colname='Pval',
    pval_name='Undjusted P-Values'
    ) # + scale_fill_gradient2(low='#000000', mid='#222222', high='#FFFFFF')
```

