---
title: "${miseq_project_prefix}: ${analysis_type}"
output:
  html_notebook:
  df_print: paged
number_sections: yes
theme: lumen
toc: yes
code_folding: hide
---

  <style>
  html {
    font-size: 16pt;
  }

body {
  font-size: 16pt;
}

h1 {
  font-size: 2.2rem;
}

h2 {
  font-size: 2rem;
}

h3 {
  font-size: 1.8rem;
}

h4 {
  font-size: 1.4rem;
}

</style>


```{r}
analysis_type = "${analysis_type}"
clustering_level = "${clustering_level}"
analysis_title = "IgA-Seq"
```

# MiSeq-0278 `r analysis_title` Two Group Tests {.tabset}

**Clustering**
Data clustered at the `r toupper('${clustering_level}')` Level.

**Subgroups**
```{r, results='asis', echo=FALSE}
for (sg_filter in c(${subgroup_filters}))
{
  cat(sg_filter, '\n\n')
}
```

**Filtering**
* **Relative Abundance cutoff:** ${relative_abundance_cutoff}
* **Prevalence Cutoff:** ${prevalence_cutoff}
* **Min Count Cutoff:** ${min_count_cutoff}

```{r}
relative_abundance_cutoff = ${relative_abundance_cutoff}
prevalence_cutoff = ${prevalence_cutoff}
min_count_cutoff = ${min_count_cutoff}
```

## Setup

### Start Conda ENV
```{r}
startCondaEnv('regression')
```

### Load Libraries
```{r}
library(rmarkdown)
#library(knitr)

library(tidyverse)
library(magrittr)
library(DT)
library(openxlsx)
library(ggplot2)
library(ggbeeswarm)

### Custom libraries that can be loaded from GitHub
source('~/utilities/analysis-utilities/general_asv_data_analysis_utilities.R')
source('~/utilities/amd_templates/setup/amd_project_utilities.R')
```

### Set Knitr Options
```{r}
${knitr_options}
```

### Load Data
```{r}
filename = makeDataFileName(
  'master_tables.xlsx',
  "${tables_dir}",
  "${analysis_type}",
  "${miseq_project_prefix}",
  "${clustering_level}"
)

iga_pos_master_table = readWorkbook(filename, sheet='Iga Pos')
iga_neg_master_table = readWorkbook(filename, sheet='Iga Neg')
iga_index_master_table = readWorkbook(filename, sheet='Iga Index')
ici_score_master_table = readWorkbook(filename, sheet='Ici Score')
log_ici_score_master_table = readWorkbook(filename, sheet='log Ici Score')
alt_index_master_table = readWorkbook(filename, sheet='Alt Index')
top_n_features = readWorkbook(filename, sheet='Taxa') %>%
  unlist() %>%
  unname()

filename = file.path(
  "${project_root}", 
  "${metadata_dir}", 
  'observational_variables.R'
  )
source(filename)
```

**Included Taxa: `print(top_n_features)`

## Variables of interest
#### Observational Variables
```{r}
case = ${case} %>% setdiff('')
key_covariates = ${key_covariates} %>% setdiff('')
linear_predictors = ${linear_predictors} %>% setdiff('')
logistic_predictors = ${logistic_predictors} %>% setdiff('')
```

* **Case:** ${case}
* **Key Covariates:** ${key_covariates}
* **linear_predictors:** ${linear_predictors}
* **logistic_predictors:** ${logistic_predictors}


### Response Variables

```{r}
raw_exp_vars = ${raw_exp_vars}
calculated_exp_vars = ${calculated_exp_vars}
all_experimental_variables = c(raw_exp_vars, calculated_exp_vars)
```


### Set Factor Levels
```{r}
iga_index_master_table = setFactorsLevels(
  iga_index_master_table, 
  observational_variables
  )
ici_score_master_table = setFactorsLevels(
  ici_score_master_table, 
  observational_variables
  )
log_ici_score_master_table = setFactorsLevels(
  log_ici_score_master_table, 
  observational_variables
  )
alt_index_master_table = setFactorsLevels(
  alt_index_master_table, 
  observational_variables
  )
# sample_metadata = setFactorsLevels(sample_metadata, categorical_variables)
```


### Determine which covariates have sufficient samples for each value
```{r}
# num_samples_threshold = 5
# 
# all_variables = filterVarsBySampleThreshold(
#   iga_index_master_table,
#   num_samples_threshold,
#   observational_variables[linear_predictors]
#   )
# 
# age_lt_90_variables = filterVarsBySampleThreshold(
#   iga_index_master_table,
#   num_samples_threshold,
#   observational_variables[linear_predictors]
#   )
# 
# amd_only_variables = filterVarsBySampleThreshold(
#   iga_index_master_table %>% select(-CaseString),
#   num_samples_threshold,
#   observational_variables[linear_predictors]
#   )
# 
# control_only_variables = filterVarsBySampleThreshold(
#   iga_index_master_table %>% filter(CaseString=='Control') %>% select(-CaseString),
#   num_samples_threshold,
#   observational_variables[linear_predictors]
#   )

```

**Raw Experimental Variables: ** `r raw_exp_vars`
**Calculated Experimental Variables: ** `r calculated_exp_vars`

## Regression analysis

### General Parameters
```{r}
stats = ${stats}
transformation = "${regression_transform}"
log_regularizer = ${log_regularizer}
adjustment_method = "${adjustment_method}"
response_var_name = "${response_var_name}"
```


### All

```{r}
all_iga_index_regression_stats = getMultipleRegressionStats(
  predictor_vars=observational_variables[linear_predictors],
  response_vars=top_n_features,
  master_table=iga_index_master_table,
  stats=stats,
  response_var_name=response_var_name,
  adjustment_method=adjustment_method,
  transformation=transformation,
  log_regularizer=log_regularizer
)

all_ici_score_regression_stats = getMultipleRegressionStats(
  predictor_vars=observational_variables[linear_predictors],
  response_vars=top_n_features,
  master_table=ici_score_master_table,
  stats=stats,
  response_var_name=response_var_name,
  adjustment_method=adjustment_method,
  transformation=transformation,
  log_regularizer=log_regularizer
)

all_log_ici_score_regression_stats = getMultipleRegressionStats(
  predictor_vars=observational_variables[linear_predictors],
  response_vars=top_n_features,
  master_table=log_ici_score_master_table,
  stats=stats,
  response_var_name=response_var_name,
  adjustment_method=adjustment_method,
  transformation=transformation,
  log_regularizer=log_regularizer
)

```


### Age < 90
```{r}
age_lt_90_iga_index_regression_stats = getMultipleRegressionStats(
  predictor_vars=observational_variables[linear_predictors],
  response_vars=top_n_features,
  master_table=iga_index_master_table %>% filter(Age < 90),
  stats=stats,
  response_var_name=response_var_name,
  adjustment_method=adjustment_method,
  transformation=transformation,
  log_regularizer=log_regularizer
)

age_lt_90_ici_score_regression_stats = getMultipleRegressionStats(
  predictor_vars=observational_variables[linear_predictors],
  response_vars=top_n_features,
  master_table=ici_score_master_table %>% filter(Age < 90),
  stats=stats,
  response_var_name=response_var_name,
  adjustment_method=adjustment_method,
  transformation=transformation,
  log_regularizer=log_regularizer
)

age_lt_90_log_ici_score_regression_stats = getMultipleRegressionStats(
  predictor_vars=observational_variables[linear_predictors],
  response_vars=top_n_features,
  master_table=log_ici_score_master_table %>% filter(Age < 90),
  stats=stats,
  response_var_name=response_var_name,
  adjustment_method=adjustment_method,
  transformation=transformation,
  log_regularizer=log_regularizer
)

```

### AMD Only
```{r}
amd_only_linear_predictors = linear_predictors %>% setdiff(c("CaseString","AREDS"))

amd_only_iga_index_regression_stats = getMultipleRegressionStats(
  predictor_vars=observational_variables[amd_only_linear_predictors],
  response_vars=top_n_features,
  master_table=iga_index_master_table %>% filter(CaseString=='AMD'),
  stats=stats,
  response_var_name=response_var_name,
  adjustment_method=adjustment_method,
  transformation=transformation,
  log_regularizer=log_regularizer
)

amd_only_ici_score_regression_stats = getMultipleRegressionStats(
  predictor_vars=observational_variables[amd_only_linear_predictors],
  response_vars=top_n_features,
  master_table=ici_score_master_table %>% filter(CaseString=='AMD'),
  stats=stats,
  response_var_name=response_var_name,
  adjustment_method=adjustment_method,
  transformation=transformation,
  log_regularizer=log_regularizer
)

amd_only_log_ici_score_regression_stats = getMultipleRegressionStats(
  predictor_vars=observational_variables[amd_only_linear_predictors],
  response_vars=top_n_features,
  master_table=log_ici_score_master_table %>% filter(CaseString=='AMD'),
  stats=stats,
  response_var_name=response_var_name,
  adjustment_method=adjustment_method,
  transformation=transformation,
  log_regularizer=log_regularizer
)

```

## Logistic Regression
### Logistic Predictor Vars
```{r}
# num_samples_threshold = 5
# 
# all_variables = filterVarsBySampleThreshold2(
#   all_master_table,
#   num_samples_threshold,
#   observational_variables[logistic_predictors]
#   )
# 
# vars = setdiff(all_variables %>% names(), c('CaseString', 'AREDS'))
```

```
### Run Regression

all_logistic_regression_stats = getLogisticRegressionStats(
  base_predictor_vars=observational_variables[vars],
  additional_predictors=c(raw_exp_vars, calculated_exp_vars),
  logistic_response_var='CaseString',
  master_table=all_master_table,
  stats=c('pvalue', 'effect_size', 'std_error'),
  response_var_name='Index'
)

contrasts = makeContrastNames(observational_variables[vars], all_master_table)

logistic_regression_stats = data.frame(ResponseVar=character())
for (contrast in contrasts)
{
  logistic_regression_stats %<>% mutate(!!contrast := numeric())
}
for (var in c(raw_exp_vars, calculated_exp_vars))
{
  logistic_regression_stats %<>% mutate(!!var := numeric())
}

logistic_regression_pvalues = logistic_regression_stats
logistic_regression_odds_ratios = logistic_regression_stats

for(var in names(all_logistic_regression_stats))
{
  
  temp = 
    all_logistic_regression_stats[[var]] %>%
    select(contrast, pvalue, effect_size) %>%
    column_to_rownames('contrast') %>%
    t() %>%
    data.frame() %>%
    rownames_to_column('stat') %>%
    mutate(ResponseVar='CaseString')
  
  pvalues = temp %>% filter(stat=='pvalue') %>% select(-stat)
  odds_ratios = temp %>% filter(stat=='effect_size') %>% select(-stat)
    
  logistic_regression_pvalues %<>% add_row(!!!pvalues)
  logistic_regression_odds_ratios %<>% add_row(!!!odds_ratios)
}

```

## Write Regression Stats
```{r}

regression_stats_table_names = c(
  "all_iga_index", 
  "age_lt_90_iga_index",
  "amd_only_iga_index",
  "all_ici_score",
  "age_lt_90_ici_score",
  "amd_only_ici_score",
  "all_log_ici_score",
  "age_lt_90_log_ici_score",
  "amd_only_log_ici_score"
  )


contrast_names = makeContrastNames(
  observational_variables,
  iga_index_master_table
)

for (tab_name in regression_stats_table_names)
{
  contrast_regression_stats = createWorkbook()
  
  print(tab_name)
  stats_table_name = paste0(tab_name, "_regression_stats")
  print(stats_table_name)
  stats_table = get(stats_table_name)
  for (contrast in contrast_names)
  {
    print(contrast)
    addWorksheet(contrast_regression_stats, sheetName=contrast)
    openxlsx::writeData(
    wb=contrast_regression_stats, 
    sheet=contrast, 
    x=stats_table[[contrast]]
    )
  }
  
  filename = makeDataFileName(
    paste0(stats_table_name, '_regression_stats_tables.xlsx'),
    "${tables_dir}",
    "${analysis_type}",
    "${miseq_project_prefix}",
    "${clustering_level}"
  )
  print(filename)
  
  saveWorkbook(contrast_regression_stats, file=filename, overwrite=T)
  
}

```

```
addWorksheet(
  regression_pvalues, 
  sheetName="Logistic P-Values"
  )

openxlsx::writeData(
  regression_pvalues,
  sheet="Logistic P-Values", 
  logistic_regression_pvalues
  )

addWorksheet(
  regression_pvalues, 
  sheetName="Logistic Odds Ratios"
  )

openxlsx::writeData(
  regression_pvalues,
  sheet="Logistic Odds Ratios", 
  logistic_regression_odds_ratios
  )
```

