---
title: "${miseq_project_prefix}: ${analysis_type}"
output:
  html_notebook:
    df_print: paged
    number_sections: yes
    theme: lumen
    toc: yes
    code_folding: hide
---

<style>
  html {
    font-size: 16pt;
  }
  
  body {
    font-size: 16pt;
  }
  
  h1 {
    font-size: 2.2rem;
  }
  
  h2 {
    font-size: 2rem;
  }
  
  h3 {
    font-size: 1.8rem;
  }
  
  h4 {
    font-size: 1.4rem;
  }
  
</style>

```{r}
source('${project_metadata_file}')
analysis_type = "${analysis_type}"
clustering_level = "${clustering_level}"
tables_dir = "${tables_dir}"
```

# ${miseq_project_prefix} ${analysis_title} Linear Regression Plots{.tabset}
Regression coefficients as effect size.

**Clustering**
Data clustered at the `r toupper('${clustering_level}')` Level.

**Subgroups**
  ```{r, results='asis', echo=FALSE}
for (sg_filter in c(${subgroup_filters}))
{
  cat(sg_filter, '\n\n')
}
```

**Filtering**
* **Relative Abundance cutoff:** ${relative_abundance_cutoff}
* **Prevalence Cutoff:** ${prevalence_cutoff}
* **Min Count Cutoff:** ${min_count_cutoff}

```{r}
relative_abundance_cutoff = ${relative_abundance_cutoff}
prevalence_cutoff = ${prevalence_cutoff}
min_count_cutoff = ${min_count_cutoff}
```

## Setup {.tabset}

### [Close]

### Start Conda ENV
```{r}
startCondaEnv('regression', lib="~/R35")
```


### Load Libraries
```{r}
library(rmarkdown)
library(tidyverse)
library(magrittr)
library(ggplot2)
library(ggbeeswarm)
library(openxlsx)
library(DT)
library(pheatmap)
library(kableExtra)
# library(vegan)
# library(nlme)

### Custom libraries that can be loaded from GitHub
source('https://raw.githubusercontent.com/ohsu-microbiome/utilities/master/analysis-utilities/general_asv_data_analysis_utilities.R')

```

### Set Knitr Options
```{r}
${knitr_options}
```

### Load Data
```{r}
filename = makeDataFileName(
  'master_table.xlsx',
  tables_dir,
  analysis_type,
  "${miseq_project_prefix}",
  "${clustering_level}"
)

print(filename)

all_master_table = readWorkbook(filename, sheet='Master Table')
age_lt_90_master_table = all_master_table %>% filter(Age < 90)
amd_only_master_table = all_master_table %>% filter(CaseString == 'AMD')
filtered_taxa = readWorkbook(filename, sheet="${tools::toTitleCase(clustering_level)}") %>% pull("short_glommed_taxa")

sample_data = read.xlsx('${sample_data_file}')

source(file.path(
  "${project_root}", 
  "${metadata_dir}", 
  'observational_variables.R'
  ))

filename = makeDataFileName(
  'linear_regression.xlsx',
  tables_dir,
  analysis_type,
  "${miseq_project_prefix}",
  "${clustering_level}"
)

pvalues = readWorkbook(filename, sheet='Unadjusted P-Values')
effect_sizes = readWorkbook(filename, sheet='Effect Sizes')

```

## Plots {.tabset}
### [ ] Close
### Gather Stats
```{r}
pvalues = 
  pvalues %>%
  remove_rownames() %>%
  column_to_rownames('Index') %>%
  t() %>% data.frame() %>%
  rownames_to_column('Predictor')

effect_sizes = 
  effect_sizes  %>%
  remove_rownames() %>%
  column_to_rownames('Index') %>%
  t() %>% data.frame() %>%
  rownames_to_column('Predictor')
```

### Plot Effect Size
```{r}
response_vars = colnames(pvalues) %>% setdiff('Predictor')

for (var in response_vars)
{
  plotEffectSizes(
    pvals,
    effect_sizes,
    response_var=var,
    effect_size_template='(log) Regression Cofficient',
    category_label='Observational Variables',
    subtitle="${analysis_subtitle}"
  ) 
}
```


