---
title: "${miseq_project_prefix}: ${analysis_type}"
output:
  html_notebook:
    df_print: paged
    number_sections: yes
    theme: lumen
    toc: yes
    code_folding: hide
---

<style>
  html {
    font-size: 16pt;
  }
  
  body {
    font-size: 16pt;
  }
  
  h1 {
    font-size: 2.2rem;
  }
  
  h2 {
    font-size: 2rem;
  }
  
  h3 {
    font-size: 1.8rem;
  }
  
  h4 {
    font-size: 1.4rem;
  }
  
</style>


```{r}
analysis_type = "${analysis_type}"
clustering_level = "${clustering_level}"
tables_dir = "${tables_dir}"
analysis_title = "${analysis_title}"
```

# ${miseq_project_prefix} ${analysis_title} Linear Regression Analysis {.tabset}
**Clustering**
Data clustered at the `r toupper('${clustering_level}')` Level.

**Subgroups**
```{r, results='asis', echo=FALSE}
for (sg_filter in c(${subgroup_filters}))
{
  cat(sg_filter, '\n\n')
}
```

**Filtering**
  * **Relative Abundance cutoff:** ${relative_abundance_cutoff}
* **Prevalence Cutoff:** ${prevalence_cutoff}
* **Min Count Cutoff:** ${min_count_cutoff}
```{r}
relative_abundance_cutoff = ${relative_abundance_cutoff}
prevalence_cutoff = ${prevalence_cutoff}
min_count_cutoff = ${min_count_cutoff}
```

## Setup
### Start Conda ENV
```{r}
startCondaEnv('regression')
```

### Load Libraries
```{r}
library(rmarkdown)
library(tidyverse)
library(magrittr)
library(ggplot2)
library(ggbeeswarm)
library(openxlsx)
library(DT)
library(pheatmap)
library(kableExtra)
library(vegan)
library(nlme)

### Custom libraries that can be loaded from GitHub
source('~/utilities/analysis-utilities/general_asv_data_analysis_utilities.R')
source('~/utilities/amd_templates/setup/amd_project_utilities.R')
```


### Set Knitr Options
```{r}
${knitr_options}
```

### Load Data
```{r}
filename = makeDataFileName(
  'master_table.xlsx',
  "${tables_dir}",
  "${analysis_type}",
  "${miseq_project_prefix}",
  "${clustering_level}"
)

print(filename)

all_master_table = readWorkbook(filename, sheet='Master Table')
age_lt_90_master_table = all_master_table %>% filter(Age < 90)
amd_only_master_table = all_master_table %>% filter(CaseString == 'AMD')
filtered_taxa = readWorkbook(filename, sheet="${tools::toTitleCase(clustering_level)}") %>% pull("short_glommed_taxa")


source(file.path(
  "${project_root}", 
  "${metadata_dir}", 
  'observational_variables.R'
  ))
```


### Set Metadata Factor Levels
```{r}
all_master_table = setFactorsLevels(all_master_table, observational_variables)
age_lt_90_master_table = setFactorsLevels(age_lt_90_master_table, observational_variables)
amd_only_master_table = setFactorsLevels(amd_only_master_table, observational_variables)
```

### Variables of interest
#### Observational Variables
```{r}
case = ${case} %>% setdiff('')
key_covariates = ${key_covariates} %>% setdiff('')
linear_predictors = ${linear_predictors} %>% setdiff('')
```

* **Case:** ${case}
* **Key Covariates:** ${key_covariates}
* **linear_predictors:** ${linear_predictors}

#### Calculated Experimental Variables
```{r}
raw_exp_vars = ${raw_exp_vars}
calculated_exp_vars = ${calculated_exp_vars}
```
**Calculated Experimental Variables: ** `r calculated_exp_vars`

### Create p-val names for contrasts
```{r}
all_contrast_names = makeContrastNames(
  observational_variables[linear_predictors], 
  all_master_table
  )
```

### Initialize regression stats table

```{r}
regression_pvals = makeRegressionStatsContrastTemplate(all_contrast_names)

regression_effect_sizes = makeRegressionStatsContrastTemplate(all_contrast_names)
```

## Linear Regression {.tabset}

### All

```{r}
exp_vars = c(raw_exp_vars, calculated_exp_vars)
predictors = observational_variables[linear_predictors]
master_table = all_master_table
contrasts = all_contrast_names
index_name_template = 'All_'

# regression_pvals = regression_pvals_template
# regression_effect_sizes = regression_pvals_template

for (var in exp_vars)
{
  index_name = paste0(index_name_template, var)
  print(index_name)
  
  fit = doMultipleRegression(
    var,
    predictors,
    response_vars=var,
    master_table,
    transformation="${regression_transform}",
    log_regularizer=${log_regularizer}
  )
  
  pvals = getRegressionPvals(
    fit,
    contrasts,
    index_name
  )
  
 regression_pvals = regression_pvals %>% add_row(!!!pvals)

  effect_sizes = getRegressionEffectSizes(
    fit,
    contrasts,
    index_name
  )

  regression_effect_sizes = regression_effect_sizes %>% add_row(!!!effect_sizes)

}
```


#### Visualize
```{r}
pheatmap(regression_pvals %>% column_to_rownames('Index'), cluster_cols=F, cluster_rows=F)

pheatmap(regression_effect_sizes %>% column_to_rownames('Index'), cluster_cols=F, cluster_rows=F)
```


### Age < 90

```{r}
exp_vars = c(raw_exp_vars, calculated_exp_vars)
predictors = observational_variables[linear_predictors]
master_table = age_lt_90_master_table
contrasts = all_contrast_names
index_name_template = 'Age_lt_90_'

# regression_pvals = regression_pvals_template
# regression_effect_sizes = regression_pvals_template

for (var in exp_vars)
{
  index_name = paste0(index_name_template, var)
  print(index_name)
  
  fit = doMultipleRegression(
    var,
    predictors,
    response_vars=var,
    master_table,
    transformation="${regression_transform}",
    log_regularizer=${log_regularizer}
  )
  
  pvals = getRegressionPvals(
    fit,
    contrasts,
    index_name
  )
  
 regression_pvals = regression_pvals %>% add_row(!!!pvals)

  effect_sizes = getRegressionEffectSizes(
    fit,
    contrasts,
    index_name
  )

  regression_effect_sizes = regression_effect_sizes %>% add_row(!!!effect_sizes)

}
```


#### Visualize
```{r}
pheatmap(regression_pvals %>% column_to_rownames('Index'), cluster_cols=F, cluster_rows=F)

pheatmap(regression_effect_sizes %>% column_to_rownames('Index'), cluster_cols=F, cluster_rows=F)
```


### AMD Only

```{r}
exp_vars = c(raw_exp_vars, calculated_exp_vars)
predictors = observational_variables[linear_predictors %>% setdiff(case)]
master_table = amd_only_master_table
contrasts = all_contrast_names %>% unlist()
index_name_template = 'AMD_Only_'

# regression_pvals = regression_pvals_template
# regression_effect_sizes = regression_pvals_template

for (var in exp_vars)
{
  index_name = paste0(index_name_template, var)
  print(index_name)
  
  fit = doMultipleRegression(
    var,
    predictors,
    response_vars=var,
    master_table,
    transformation="${regression_transform}",
    log_regularizer=${log_regularizer}
  )
  
  pvals = getRegressionPvals(
    fit,
    contrasts,
    index_name
  )
  
 regression_pvals = regression_pvals %>% add_row(!!!pvals)

  effect_sizes = getRegressionEffectSizes(
    fit,
    contrasts,
    index_name
  )

  regression_effect_sizes = regression_effect_sizes %>% add_row(!!!effect_sizes)

}

regression_adjusted_pvals = 
  regression_pvals %>%
  mutate_if(is.numeric, p.adjust)
```

#### Visualize
```{r}
pheatmap(regression_pvals %>% column_to_rownames('Index'), cluster_cols=F, cluster_rows=F)

pheatmap(regression_effect_sizes %>% column_to_rownames('Index'), cluster_cols=F, cluster_rows=F)
```

### Write Regression Stats
```{r}
wb = createWorkbook()

addWorksheet(
  wb, 
  sheetName="Unadjusted P-Values"
  )

openxlsx::writeData(
  wb, 
  sheet="Unadjusted P-Values", 
  regression_pvals
  )

addWorksheet(
  wb, 
  sheetName="Adjusted P-Values"
  )

openxlsx::writeData(
  wb, 
  sheet="Adjusted P-Values", 
  regression_adjusted_pvals
  )

addWorksheet(
  wb, 
  sheetName="Effect Sizes"
  )

openxlsx::writeData(
  wb,
  sheet="Effect Sizes", 
  regression_effect_sizes
  )

# addWorksheet(regression_pvals, sheetName="Adjusted P-Values")
# openxlsx::writeData(regression_pvals, sheet="Adjusted P-Values", adjusted_regression_pvals)

filename = makeDataFileName(
  'linear_regression.xlsx',
  "${tables_dir}",
  "${analysis_type}",
  "${miseq_project_prefix}",
  "${clustering_level}"
)
saveWorkbook(wb, file=filename, overwrite=T)
```
