---
title: "MiSeq-${miseq_run_number}"
output:
  html_notebook:
    df_print: paged
    number_sections: yes
    theme: lumen
    toc: yes
    code_folding: hide
---

<style>
  html {
    font-size: 16pt;
  }
  
  body {
    font-size: 16pt;
  }
  
  h1 {
    font-size: 2.2rem;
  }
  
  h2 {
    font-size: 2rem;
  }
  
  h3 {
    font-size: 1.8rem;
  }
  
  h4 {
    font-size: 1.4rem;
  }
  
</style>


```{r}
source('${project_metadata_file}')
analysis_type = "${analysis_type}"
clustering_level = "${clustering_level}"
tables_dir = "${tables_dir}"
analysis_title = "${analysis_title}"
```

# MiSeq-${miseq_run_number} `r analysis_title` Logistic Regression Analysis {.tabset}
## Clustering Level
Data clustered at the `r toupper('${clustering_level}')` Level.

## Filtering
* **Relative Abundance Cutoff:** ${relative_abundance_cutoff}
* **Prevalence Cutoff:** ${prevalence_cutoff}
* **Min Count Cutoff:** ${min_count_cutoff}

## Setup
### Start Conda ENV
```{r}
startCondaEnv('regression')
```


### Load Libraries
```{r}
library(rmarkdown)
library(tidyverse)
library(magrittr)
library(ggplot2)
library(ggbeeswarm)
library(openxlsx)
library(DT)
library(pheatmap)
library(kableExtra)
library(vegan)
library(nlme)

### Custom libraries that can be loaded from GitHub
source('~/utilities/analysis-utilities/general_asv_data_analysis_utilities.R')
source('~/utilities/analysis-utilities/alpha_div_utilities.R')
source('~/utilities/amd_templates/setup/amd_project_utilities.R')
```

### Set Knitr Options
```{r}
${knitr_options}

```

### Load Data
```{r}
filename = makeDataFileName(
  'master_table.xlsx',
  "${tables_dir}",
  "${analysis_type}",
  "${miseq_project_prefix}",
  "${clustering_level}"
)

print(filename)

all_master_table = readWorkbook(filename, sheet='Master Table')
age_lt_90_master_table = all_master_table %>% filter(Age < 90)
amd_only_master_table = all_master_table %>% filter(CaseString == 'AMD')
filtered_taxa = readWorkbook(filename, sheet="${tools::toTitleCase(clustering_level)}") %>% pull("short_glommed_taxa")

source(file.path(
  "${project_root}", 
  "${metadata_dir}", 
  'observational_variables.R'
  ))
```

## Variables

### Set Metadata Factor Levels
```{r}
all_master_table = setFactorsLevels(
  all_master_table, 
  observational_variables
  )
```

### Observational Variables

```{r}
observational_variable_names = ${logistic_regression_predictors}
observational_variables = observational_variables[observational_variable_names]
```

Using `r observational_variable_names`

### Raw Variables
```{r}
raw_exp_vars = ${raw_exp_vars}
print(raw_exp_vars)
```

### calculated_exp_vars
```{r}
calculated_exp_vars = ${calculated_exp_vars}
print(calculated_exp_vars)
```

### Filter Covariates
Determine which covariates have sufficient samples with each value or of perfect separation.
```{r}
num_samples_threshold = 5

all_variables = filterVarsBySampleThreshold2(
  master_table=all_master_table,
  threshold=num_samples_threshold,
  variable_list=observational_variables
  )

age_lt_90_variables = filterVarsBySampleThreshold2(
  master_table=age_lt_90_master_table %>% filter(Age<90),
  threshold=num_samples_threshold,
  variable_list=observational_variables
  )
```

### Remaining Covariates
`r all_variables %>% names()`
`r age_lt_90_variables %>% names()`

## Logistic Regression
### All Subgroups

```{r}
varnames = all_variables %>% names()

all_logistic_regression_stats = getLogisticRegressionStats(
  base_predictor_vars=observational_variables[varnames],
  additional_predictors=c(raw_exp_vars, calculated_exp_vars),
  logistic_response_var='CaseString',
  master_table=all_master_table,
  stats=c('pvalue', 'effect_size', 'std_error'),
  response_var_name='Index'
)

contrasts = makeContrastNames(all_variables, all_master_table)

logistic_regression_stats = data.frame(ResponseVar=character())
for (contrast in contrasts)
{
  logistic_regression_stats %<>% mutate(!!contrast := numeric())
}
for (var in c(raw_exp_vars, calculated_exp_vars))
{
  logistic_regression_stats %<>% mutate(!!var := numeric())
}

all_logistic_regression_pvalues = logistic_regression_stats
all_logistic_regression_log_odds_ratios = logistic_regression_stats

for(var in names(all_logistic_regression_stats))
{
  temp = 
    all_logistic_regression_stats[[var]] %>%
    select(contrast, pvalue, effect_size) %>%
    column_to_rownames('contrast') %>%
    t() %>%
    data.frame() %>%
    rownames_to_column('stat') %>%
    mutate(ResponseVar='CaseString')
  
  pvalues = temp %>% filter(stat=='pvalue') %>% select(-stat)
  odds_ratios = temp %>% filter(stat=='effect_size') %>% select(-stat)
    
  all_logistic_regression_pvalues %<>% add_row(!!!pvalues)
  all_logistic_regression_log_odds_ratios %<>% add_row(!!!odds_ratios)
}

```

### Write Tables
```{r}
all_wb = createWorkbook()

addWorksheet(
  all_wb, 
  sheetName="Unadjusted P-Values"
  )

openxlsx::writeData(
  all_wb, 
  sheet="Unadjusted P-Values", 
  all_logistic_regression_pvalues
  )

addWorksheet(
  all_wb, 
  sheetName="Log Odds Ratios"
  )

openxlsx::writeData(
  all_wb,
  sheet="Log Odds Ratios", 
  all_logistic_regression_log_odds_ratios
  )

# addWorksheet(regression_pvals, sheetName="Adjusted P-Values")
# openxlsx::writeData(regression_pvals, sheet="Adjusted P-Values", adjusted_regression_pvals)

filename = makeDataFileName(
  'all_logistic_regression.xlsx',
  "${tables_dir}",
  "${analysis_type}",
  "${miseq_project_prefix}",
  "${clustering_level}"
)

saveWorkbook(all_wb, file=filename, overwrite=T)
```

### Age < 90

```{r}
age_lt_90_logistic_regression_stats = getLogisticRegressionStats(
  base_predictor_vars=observational_variables[vars],
  additional_predictors=c(raw_exp_vars, calculated_exp_vars),
  logistic_response_var='CaseString',
  master_table=age_lt_90_master_table,
  stats=c('pvalue', 'effect_size', 'std_error'),
  response_var_name='Index'
)

contrasts = makeContrastNames(age_lt_90_variables, all_master_table)

logistic_regression_stats = data.frame(ResponseVar=character())
for (contrast in contrasts)
{
  logistic_regression_stats %<>% mutate(!!contrast := numeric())
}
for (var in c(raw_exp_vars, calculated_exp_vars))
{
  logistic_regression_stats %<>% mutate(!!var := numeric())
}

age_lt_90_logistic_regression_pvalues = logistic_regression_stats
age_lt_90_logistic_regression_log_odds_ratios = logistic_regression_stats

for(var in names(all_logistic_regression_stats))
{
  temp = 
    age_lt_90_logistic_regression_stats[[var]] %>%
    select(contrast, pvalue, effect_size) %>%
    column_to_rownames('contrast') %>%
    t() %>%
    data.frame() %>%
    rownames_to_column('stat') %>%
    mutate(ResponseVar='CaseString')
  
  pvalues = temp %>% filter(stat=='pvalue') %>% select(-stat)
  odds_ratios = temp %>% filter(stat=='effect_size') %>% select(-stat)
    
  age_lt_90_logistic_regression_pvalues %<>% add_row(!!!pvalues)
  age_lt_90_logistic_regression_log_odds_ratios %<>% add_row(!!!odds_ratios)
}

```


## Write Tables
```{r}
age_lt_90_wb = createWorkbook()

addWorksheet(
  age_lt_90_wb, 
  sheetName="Unadjusted P-Values"
  )

openxlsx::writeData(
  age_lt_90_wb, 
  sheet="Unadjusted P-Values", 
  age_lt_90_logistic_regression_pvalues
  )

addWorksheet(
  age_lt_90_wb, 
  sheetName="Log Odds Ratios"
  )

openxlsx::writeData(
  age_lt_90_wb,
  sheet="Log Odds Ratios", 
  age_lt_90_logistic_regression_log_odds_ratios
  )

# addWorksheet(regression_pvals, sheetName="Adjusted P-Values")
# openxlsx::writeData(regression_pvals, sheet="Adjusted P-Values", adjusted_regression_pvals)

filename = makeDataFileName(
  'age_lt_90_logistic_regression.xlsx',
  "${tables_dir}",
  "${analysis_type}",
  "${miseq_project_prefix}",
  "${clustering_level}"
)

saveWorkbook(age_lt_90_wb, file=filename, overwrite=T)
```
