---
title: "${miseq_project_prefix}: ${analysis_type}"
output:
  html_notebook:
    df_print: paged
    number_sections: yes
    theme: lumen
    toc: yes
    code_folding: hide
---

<style>
  html {
    font-size: 16pt;
  }
  
  body {
    font-size: 16pt;
  }
  
  h1 {
    font-size: 2.2rem;
  }
  
  h2 {
    font-size: 2rem;
  }
  
  h3 {
    font-size: 1.8rem;
  }
  
  h4 {
    font-size: 1.4rem;
  }
  
</style>

```{r}
source('${project_metadata_file}')
analysis_type = "${analysis_type}"
clustering_level = "${clustering_level}"
filter_by = "${filter_by}"
tables_dir = "${tables_dir}"
```

# ${miseq_project_prefix} Master Table Prep{.tabset}
**Clustering**
  Data clustered at the `r toupper('${clustering_level}')` Level.

**Subgroups**
```{r, results='asis', echo=FALSE}
for (sg_filter in c(${subgroup_filters}))
{
  cat(sg_filter, '\n\n')
}
```

**Filtering**
Data filtered by `r toupper('${filter_by}')`

* **Relative Abundance cutoff:** ${relative_abundance_cutoff}
* **Prevalence Cutoff:** ${prevalence_cutoff}
* **Min Count Cutoff:** ${min_count_cutoff}
```{r}
relative_abundance_cutoff = ${relative_abundance_cutoff}
prevalence_cutoff = ${prevalence_cutoff}
min_count_cutoff = ${min_count_cutoff}
```

## Setup

### [Close]

### Start Conda ENV
```{r}
startCondaEnv('base')
```


### Load Libraries
```{r}
library(rmarkdown)
#library(knitr)
library(tidyverse)
library(magrittr)
library(ggplot2)
library(openxlsx)

library(pheatmap)
library(kableExtra)

### Custom libraries that can be loaded from GitHub
source('${utilities_dir}/analysis-utilities/general_asv_data_analysis_utilities.R')
source('${utilities_dir}/amd_templates/setup/amd_project_utilities.R')

```


### Set Knitr Options
```{r}

${knitr_options}
```

### Load Data
```{r}
asv_table = 
  read.delim(
    '${asv_table_file}', 
    header=T, 
    sep='\t',
    stringsAsFactors=F
    )

taxonomy_table = 
  read.delim(
    '${taxonomy_table_file}', 
    header=T, 
    sep='\t',
    stringsAsFactors=F
    ) %>%
  mutate_all(as.character())

sample_data = read.xlsx('${sample_data_file}')

sample_data_columns = colnames(sample_data)
sample_names = sample_data$SampleName

```


### Extract Subgroups
```{r}
if (!is.null(c(${subgroup_filters})))
{
  subgroup_filter_string = paste0(${subgroup_filters}, collapse=" & ")
  print(sprintf("Subgroup Filter String: %s", subgroup_filter_string))
  subgroup_filter_expression = str2expression(subgroup_filter_string)
  sample_data = sample_data %>% filter(eval(subgroup_filter_expression))
  sample_names = sample_data$SampleName

  asv_table = asv_table %>% select('ASVs', sample_names)
}

```

### Set Filtering
```{r}
relative_abundance_cutoff = ${relative_abundance_cutoff}
prevalence_cutoff = ${prevalence_cutoff}
```
**relative abundance cutoff:** `r relative_abundance_cutoff`
**prevalence cutoff:** `r prevalence_cutoff`

### Get Filtered Taxa Abundance
```{r}

all_ranks = c('Phylum', 'Class', 'Order', 'Family', 'Genus')


for (cluster_by in rev(c(all_ranks, 'ASV')))
{
  if (cluster_by == 'ASV')
  {
    ranks_to_keep = c(all_ranks, 'ASVs')
  } else
  {
    ranks_to_keep = all_ranks[1:match(cluster_by, all_ranks)]
  }
  
  wb = createWorkbook()
  
  raw_abundance = getFilteredTaxaCounts(
    asv_table,
    taxonomy_table,
    sample_data, 
    cluster_by=cluster_by,
    filter_by="${filter_by}",
    relative_abundance_cutoff=relative_abundance_cutoff,
    prevalence_cutoff=prevalence_cutoff,
    id_col="SampleName",
    normalize=F
    )
  
  addWorksheet(wb, sheetName="Raw Abundance")
  openxlsx::writeData(wb, sheet="Raw Abundance", raw_abundance)
  
  raw_master_table = getMasterTable(
    raw_abundance,
    sample_data,
    key_column='SampleName',
    taxa_colnames='short_glommed_taxa'
    )
  
  addWorksheet(wb, sheetName="Raw Master Table")
  openxlsx::writeData(wb, sheet="Raw Master Table", raw_master_table)
  
    
  raw_taxa = 
    raw_abundance %>% 
    select(ranks_to_keep, glommed_taxa, short_glommed_taxa)
  
  addWorksheet(wb, sheetName="Raw Taxa")
  openxlsx::writeData(wb, sheet="Raw Taxa", raw_taxa)
  
  normalized_abundance = getFilteredTaxaCounts(
    asv_table,
    taxonomy_table,
    sample_data, 
    cluster_by=cluster_by,
    filter_by="${filter_by}",
    relative_abundance_cutoff=relative_abundance_cutoff,
    prevalence_cutoff=prevalence_cutoff,
    min_count_cutoff=0,
    id_col="SampleName",
    normalize=T
    )
  
  addWorksheet(wb, sheetName="Normalized Abundance")
  openxlsx::writeData(wb, sheet="Normalized Abundance", normalized_abundance)

  normalized_master_table = getMasterTable(
    normalized_abundance,
    sample_data,
    key_column='SampleName',
    taxa_colnames='short_glommed_taxa'
    )
  
  addWorksheet(wb, sheetName="Normalized Master Table")
  openxlsx::writeData(wb, sheet="Normalized Master Table", normalized_master_table)
    
    
  normalized_taxa = 
    normalized_abundance %>% 
    select(ranks_to_keep, glommed_taxa, short_glommed_taxa)
  
  addWorksheet(wb, sheetName="Normalized Taxa")
  openxlsx::writeData(wb, sheet="Normalized Taxa", raw_taxa)
  
  
  filename = makeDataFileName(
    'master_tables.xlsx',
    "${tables_dir}",
    NULL,
    "${miseq_project_prefix}",
    tolower(cluster_by)
    )
  
  print(filename)
  
  saveWorkbook(wb, file=filename, overwrite=T)
}


```

