---
title: "${miseq_project_prefix}: ${analysis_type}"
output:
  html_notebook:
    df_print: paged
    number_sections: yes
    theme: lumen
    toc: yes
    code_folding: hide
---

<style>
  html {
    font-size: 16pt;
  }
  
  body {
    font-size: 16pt;
  }
  
  h1 {
    font-size: 2.2rem;
  }
  
  h2 {
    font-size: 2rem;
  }
  
  h3 {
    font-size: 1.8rem;
  }
  
  h4 {
    font-size: 1.4rem;
  }
  
</style>

```{r}
source('${project_metadata_file}')
analysis_type = "${analysis_type}"
clustering_level = "${clustering_level}"
tables_dir = "${tables_dir}"
analysis_title = tools::toTitleCase(gsub("_", " ", analysis_type))
```

# ${miseq_project_prefix} Logistic Regression AMD Only {.tabset}

**Clustering**
Data clustered at the `r toupper('${clustering_level}')` Level.

**Subgroups**
```{r, results='asis', echo=FALSE}
for (sg_filter in c(${subgroup_filters}))
{
  cat(sg_filter, '\n\n')
}
```

**Filtering**
* **Relative Abundance cutoff:** ${relative_abundance_cutoff}
* **Prevalence Cutoff:** ${prevalence_cutoff}
* **Min Count Cutoff:** ${min_count_cutoff}

```{r}
relative_abundance_cutoff = ${relative_abundance_cutoff}
prevalence_cutoff = ${prevalence_cutoff}
min_count_cutoff = ${min_count_cutoff}
```

## Setup

### Start Conda ENV
```{r}
startCondaEnv('regression', lib='~/R35')
```
### Load Libraries
```{r}
library(rmarkdown)
library(tidyverse)
library(magrittr)
library(ggplot2)
library(ggbeeswarm)
library(openxlsx)
library(DT)
library(pheatmap)
library(kableExtra)
library(vegan)
library(nlme)

### Custom libraries that can be loaded from GitHub
source('https://raw.githubusercontent.com/ohsu-microbiome/utilities/master/analysis-utilities/general_asv_data_analysis_utilities.R')

```

### Set Knitr Options
```{r}
${knitr_options}
```


### Load Data
```{r}
filename = makeDataFileName(
  'master_table.xlsx',
  "${tables_dir}",
  "${analysis_type}",
  "${miseq_project_prefix}",
  "${clustering_level}"
)

print(filename)

master_table = readWorkbook(filename, sheet='Master Table')
amd_only_master_table = master_table %>% filter(CaseString == 'AMD')

filtered_taxa = readWorkbook(
  filename, 
  sheet="${tools::toTitleCase(clustering_level)}"
  ) %>% 
  pull("short_glommed_taxa")

source(file.path(
  "${project_root}", 
  "${metadata_dir}", 
  'observational_variables.R'
  ))
```

### Extract Subgroups
**Subgroup Filters Applied in Prep: ** ${subgroup_filters}

### Set Factor Levels
```{r}
amd_only_master_table = setFactorsLevels(
  amd_only_master_table,
  observational_variables
  )
```

### Variables of interest
#### Observational Variables
```{r}
case = ${case} %>% setdiff('')
treatments = ${treatments} %>% setdiff('')
phenotypes = ${phenotypes} %>% setdiff('')
genotypes = ${genotypes} %>% setdiff('')
key_covariates = ${key_covariates} %>% setdiff('')

logistic_covariates = ${logistic_covariates}

```

* **Case:** `r case`
* **Genotype variables:** `r genotypes`
* **Phenotype variables:** `r phenotypes`
* **Treatment variables:** `r treatments`
* **Key covariates:**  `r key_covariates`

#### Raw Variables
```{r}
raw_exp_vars = ${raw_exp_vars}
print(raw_exp_vars)
```

#### calculated_exp_vars
```{r}
calculated_exp_vars = ${calculated_exp_vars}
print(calculated_exp_vars)
```


#### Response Variables
```{r}
target_variables = 
  c(raw_exp_vars, calculated_exp_vars) %>%
  setdiff("")

response_variables = c(phenotypes) %>% setdiff("")
```


## AMD Only
### Create Workbook
```{r}
wb = createWorkbook()

addWorksheet(
  wb, 
  sheetName="Response Variables"
  )
openxlsx::writeData(
  wb,
  sheet="Response Variables",
  response_variables
  )

addWorksheet(
  wb, 
  sheetName="Target Variables"
  )
openxlsx::writeData(
  wb,
  sheet="Target Variables",
  target_variables
  )

```

### Regressions Results Template
```{r}
covariates = observational_variables[logistic_covariates]
contrasts = makeContrastNames(covariates, amd_only_master_table)
logistic_regression_results_template = data.frame(
  Target_Variable=character()
  )

for (contrast in c("Target_Variable", contrasts))
{
  print(contrast)
  
  pval_name = paste0(contrast, '.pval')
  odds_name = paste0(contrast, '.odds')
  print(pval_name)
  print(odds_name)
  logistic_regression_results_template %<>% 
    mutate(
      !!pval_name := numeric(),
      !!odds_name := numeric()
    )
}
```

### Logistic Regression
```{r}
for(response_var in response_variables)
{
  
  addWorksheet(wb, sheetName=response_var)
  
  logistic_regression_results = logistic_regression_results_template
  
  
  ### Target Variables Only
  logistic_regression_stats = getLogisticRegressionStats(
    base_predictor_vars=c(),
    additional_predictors=target_variables,
    logistic_response_var=observational_variables[[response_var]],
    master_table=amd_only_master_table,
    stats=c('pvalue', 'effect_size', 'std_error'),
    response_var_name='Index'
  )
  
  for(var in names(logistic_regression_stats))
  {
    print(sprintf('var=%s', var))
    
    var_data = logistic_regression_stats[[var]]
    print('var_data')
    glimpse(var_data)
    
    stats_list = list()
    stats_list[['Target_Variable']] = var
        
    pval_name = paste0('Target_Variable', '.pval')
    odds_name = paste0('Target_Variable', '.odds')
  
    stats_list[[pval_name]] =
      var_data %>%
      filter(contrast == !!var) %>%
      pull(pvalue)
    stats_list[[odds_name]] =
      var_data %>%
      filter(contrast == !!var) %>%
      pull(effect_size) %>%
      exp()
  
    print('stats_list')
    print(stats_list)
  
    setdiff(names(stats_list), colnames(logistic_regression_results))
      
    logistic_regression_results %<>% add_row(!!!stats_list)
  }
  
  
  
  ### Key Covariates Only
  logistic_regression_stats = getLogisticRegressionStats(
    base_predictor_vars=covariates,
    additional_predictors=c(''),
    logistic_response_var=observational_variables[[response_var]],
    master_table=amd_only_master_table,
    stats=c('pvalue', 'effect_size', 'std_error'),
    response_var_name='Index'
  )
  
  logistic_regression_stats = list(
    None=logistic_regression_stats %>%
      .[[2]] %>%
      data.frame() %>%
      filter(contrast != '')
  )
  
  contrasts = makeContrastNames(covariates, amd_only_master_table)
  
  ### Fill stats_list with results for each special variable
  for(var in c("None"))
  {
    print(sprintf('var=%s', var))
    
    var_data = logistic_regression_stats[[var]]
    print('var_data')
    glimpse(var_data)
    
    stats_list = list()
    
    for (contrast in contrasts)
    {
      print(sprintf('contrast=%s', contrast))
      
      pval_name = paste0(contrast, '.pval')
      odds_name = paste0(contrast, '.odds')
      
      print(sprintf('pval_name=%s  odds_name=%s', pval_name, odds_name))
      
      stats_list[[pval_name]] = 
        var_data %>%
        filter(contrast == !!contrast) %>%
        pull(pvalue)
      stats_list[[odds_name]] = 
        var_data %>% 
        filter(contrast == !!contrast) %>%
        pull(effect_size) %>%
        exp()
      
      print('stats_list')
      print(stats_list)
    }
     
      
    stats_list = c(
      list(
        Target_Variable=NA,
        Target_Variable.pval=NA,
        Target_Variable.odds=NA
      ),
      stats_list
    )
    
    logistic_regression_results %<>% add_row(!!!stats_list)
  }


  ### Target Variables
  logistic_regression_stats = getLogisticRegressionStats(
    base_predictor_vars=covariates,
    additional_predictors=target_variables,
    logistic_response_var=observational_variables[[response_var]],
    master_table=amd_only_master_table,
    stats=c('pvalue', 'effect_size', 'std_error'),
    response_var_name='Index'
  )
  
  for(target_var in names(logistic_regression_stats))
  {
    print(sprintf('target_var=%s', target_var))
    
    var_data = logistic_regression_stats[[target_var]]
    print('var_data')
    glimpse(var_data)
    
    stats_list = list()
    stats_list[['Target_Variable']] = target_var
        
    pval_name = paste0('Target_Variable', '.pval')
    odds_name = paste0('Target_Variable', '.odds')
  
    stats_list[[pval_name]] =
      var_data %>%
      filter(contrast == !!target_var) %>%
      pull(pvalue)
    stats_list[[odds_name]] =
      var_data %>%
      filter(contrast == !!target_var) %>%
      pull(effect_size) %>%
      exp()
  
    print('stats_list')
    print(stats_list)
  
    for (contrast in contrasts)
    {
      print(sprintf('contrast=%s', contrast))
      
      pval_name = paste0(contrast, '.pval')
      odds_name = paste0(contrast, '.odds')
      
      stats_list[[pval_name]] = 
        var_data %>%
        filter(contrast == !!contrast) %>%
        pull(pvalue)
      stats_list[[odds_name]] = 
        var_data %>% 
        filter(contrast == !!contrast) %>%
        pull(effect_size) %>%
        exp()
      
      print('stats_list')
      print(stats_list)
      
    } ### end contrast, meaning covariate levels
    
    setdiff(names(stats_list), colnames(logistic_regression_results))
      
    logistic_regression_results %<>% add_row(!!!stats_list)
    
  } ### end response_vars, e.g. phenotype
    
  ### Write the worksheet for given response_var
  openxlsx::writeData(
    wb,
    sheet=response_var,
    logistic_regression_results
    )
  
} ### end target_vars, e.g. phyla
  

```


### Save Workbook
```{r}
filename = makeDataFileName(
  'logistic_regression_amd_only.xlsx',
  "${tables_dir}",
  "${analysis_type}",
  "${miseq_project_prefix}",
  "${clustering_level}"
)

saveWorkbook(wb, file=filename, overwrite=T)
```

