---
title: "MiSeq-${miseq_run_number}"
output:
  html_notebook:
    df_print: paged
    number_sections: yes
    theme: lumen
    toc: yes
    code_folding: hide
---

<style>
  html {
    font-size: 16pt;
  }
  
  body {
    font-size: 16pt;
  }
  
  h1 {
    font-size: 2.2rem;
  }
  
  h2 {
    font-size: 2rem;
  }
  
  h3 {
    font-size: 1.8rem;
  }
  
  h4 {
    font-size: 1.4rem;
  }
  
</style>


```{r}
source('${project_metadata_file}')
analysis_type = "${analysis_type}"
clustering_level = "${clustering_level}"
tables_dir = "${tables_dir}"
analysis_title = "${analysis_title}"
```

# MiSeq-${miseq_run_number} `r analysis_title` Logistic Regression Analysis {.tabset}
## Clustering Level
Data clustered at the `r toupper('${clustering_level}')` Level.

## Setup
### Start Conda ENV
```{r}
startCondaEnv('alphadiv')
```


### Load Libraries
```{r}
library(rmarkdown)
library(tidyverse)
library(magrittr)
library(ggplot2)
library(ggbeeswarm)
library(openxlsx)
library(DT)
library(pheatmap)
library(kableExtra)
library(vegan)
library(nlme)

### Custom libraries that can be loaded from GitHub
source('~/utilities/analysis-utilities/general_asv_data_analysis_utilities.R')
source('~/utilities/analysis-utilities/alpha_div_utilities.R')
source('~/utilities/amd_templates/setup/amd_project_utilities.R')
```

### Set Knitr Options
```{r}
${knitr_options}

```

### Load Data
```{r}
filename = makeDataFileName(
  'master_table.xlsx',
  "${tables_dir}",
  "${analysis_type}",
  "${miseq_project_prefix}",
  "${clustering_level}"
)

master_table = readWorkbook(filename, sheet='Master Table')
filtered_taxa = readWorkbook(filename, sheet="Filtered Taxa", rowNames=F) %>% pull(1)

source(file.path(
  "${project_root}", 
  "${metadata_dir}", 
  'observational_variables.R'
  ))
```

## Variables
### Observational Variable Names
```{r}
observational_variable_names = observational_variables %>% names() %>% unname()
print(observational_variable_names)
```

### Set Metadata Factor Levels
```{r}
master_table = setFactorsLevels(
  master_table, 
  observational_variable_names
  )
```

### Explicitly remove variables that should not be used

#### Remove AREDS
AREDS cannot be viewed as a predictor for Case.
```{r}
observational_variables$AREDS = NULL
```

#### Remove CaseString
Case cannot be viewed as a predictor for Case
```{r}
observational_variables$CaseString = NULL
```

### Filter Covariates
Determine which covariates have sufficient samples with each value or of perfect separation.
```{r}
num_samples_threshold = 5

all_variables = filterVarsBySampleThreshold2(
  master_table=master_table,
  threshold=num_samples_threshold,
  variable_list=observational_variables
  )

age_lt_90_variables = filterVarsBySampleThreshold2(
  master_table=all_master_table %>% filter(Age<90),
  threshold=num_samples_threshold,
  variable_list=observational_variables
  )
```

### Remaining Covariates
`r all_variables`
`r age_lt_90_variables`


## Logistic Regression {.tabset}
### [ ] Close

### Run Regression Tests

```{r}
vars = all_variables %>% names()

all_logistic_regression_stats = getLogisticRegressionStats(
  base_predictor_vars=observational_variables[vars],
  additional_predictors=c(filtered_taxa, 'pct_prevotella'),
  logistic_response_var='CaseString',
  master_table=prev_master_table,
  stats=c('pvalue', 'effect_size', 'std_error'),
  response_var_name='Taxa'
)

taxa_stats = data.frame(
  Taxa=character(), 
  pvalue=numeric(), 
  effect_size=numeric(),
  std_error=numeric()
  )

for(taxon in names(all_logistic_regression_stats))
{
  stats = 
    all_logistic_regression_stats[[taxon]] %>% 
    filter(contrast==taxon) %>% 
    # select(-std_error) %>%
    rename(Taxa=contrast) %>%
    as.list()
  
  # print(stats)
  
  taxa_stats %<>% add_row(!!!stats)
}

taxa_stats = taxa_stats %>% arrange(pvalue)
```

### Results
```{r}
kable(taxa_stats, class='cell-border stripe compact') %>%
  kable_styling(
    bootstrap_options = "striped", 
    full_width = F, 
    font_size=10
    )
```

### Write Results
```{r}
filename = makeDataFileName(
  'regression_stats.xlsx', 
  "${tables_dir}",
  "${analysis_type}",
  "${miseq_project_prefix}",
  "${clustering_level}"
  )

write.xlsx(
  taxa_stats,
  file=filename
)
```



```
all_prevotella_taxa = c("Bacteroidetes_Alloprevotella", "Bacteroidetes_Paraprevotella", "Bacteroidetes_Prevotella", "Bacteroidetes_Prevotella_2", "Bacteroidetes_Prevotella_6", "Bacteroidetes_Prevotella_7", "Bacteroidetes_Prevotella_9", "Bacteroidetes_Prevotellaceae_Ga6A1_group", "Bacteroidetes_Prevotellaceae_NK3B31_group", "Bacteroidetes_Prevotellaceae_UCG_dash_001", "Bacteroidetes_Prevotellaceae_UCG_dash_003", "Bacteroidetes_Prevotellaceae_UCG_dash_004")

prevotella_taxa_in_filtered_taxa = intersect(all_prevotella_taxa, filtered_taxa)
```

```
prev_master_table =
  all_master_table %>%
  mutate(pct_prevotella = rowSums(select(., prevotella_taxa_in_filtered_taxa)))
```
