---
title: "MiSeq-${miseq_run_number}"
output:
  html_notebook:
    df_print: paged
    number_sections: yes
    theme: lumen
    toc: yes
    code_folding: hide
---

<style>
  html {
    font-size: 16pt;
  }
  
  body {
    font-size: 16pt;
  }
  
  h1 {
    font-size: 2.2rem;
  }
  
  h2 {
    font-size: 2rem;
  }
  
  h3 {
    font-size: 1.8rem;
  }
  
  h4 {
    font-size: 1.4rem;
  }
  
</style>


```{r}
source('${project_metadata_file}')
analysis_type = "${analysis_type}"
clustering_level = "${clustering_level}"
tables_dir = "${tables_dir}"
analysis_title = "${analysis_title}"
```

# MiSeq-${miseq_run_number} ${analysis_title} Linear Regression Analysis {.tabset}
## Clustering Level
Data clustered at the `r toupper('${clustering_level}')` Level.

## Filtering
* **Relative Abundance Cutoff:** ${relative_abundance_cutoff}
* **Prevalence Cutoff:** ${prevalence_cutoff}
* **Min Count Cutoff:** ${min_count_cutoff}

## Setup
### Start Conda ENV
```{r}
startCondaEnv('regression')
```

### Load Libraries
```{r}
library(rmarkdown)
library(tidyverse)
library(magrittr)
library(ggplot2)
library(ggbeeswarm)
library(openxlsx)
library(DT)
library(pheatmap)
library(kableExtra)
library(vegan)
library(nlme)

### Custom libraries that can be loaded from GitHub
source('~/utilities/analysis-utilities/general_asv_data_analysis_utilities.R')
source('~/utilities/analysis-utilities/alpha_div_utilities.R')
source('~/utilities/amd_templates/setup/amd_project_utilities.R')
```


### Set Knitr Options
```{r}
${knitr_options}
```

### Load Data
```{r}
filename = makeDataFileName(
  'master_table.xlsx',
  "${tables_dir}",
  "${analysis_type}",
  "${miseq_project_prefix}",
  "${clustering_level}"
)

print(filename)

all_master_table = readWorkbook(filename, sheet='Master Table')
age_lt_90_master_table = all_master_table %>% filter(Age < 90)
amd_only_master_table = all_master_table %>% filter(CaseString == 'AMD')
filtered_taxa = readWorkbook(filename, sheet="${tools::toTitleCase(clustering_level)}") %>% pull("short_glommed_taxa")


source(file.path(
  "${project_root}", 
  "${metadata_dir}", 
  'observational_variables.R'
  ))
```

### Variables of interest
#### All Observational Variables
```{r}
observational_variable_names = observational_variables %>% names()
print(observational_variables %>% names())
```
#### Selected Variables
```{r}

```

#### Calculated Experimental Variables

```{r}
raw_exp_vars = ${raw_exp_vars}
calculated_exp_vars = ${calculated_exp_vars}
```
**Calculated Experimental Variables: ** `r calculated_exp_vars`

### Set Metadata Factor Levels
```{r}
all_master_table = setFactorsLevels(all_master_table, observational_variables)
age_lt_90_master_table = setFactorsLevels(age_lt_90_master_table, observational_variables)
amd_only_master_table = setFactorsLevels(amd_only_master_table, observational_variables)
```


### Determine which covariates have sufficient samples with each value
```{r}
num_samples_threshold = 5

all_variables = filterVarsBySampleThreshold(
  all_master_table,
  num_samples_threshold,
  observational_variables
  )

age_lt_90_variables = filterVarsBySampleThreshold(
  age_lt_90_master_table,
  num_samples_threshold,
  observational_variables
  )

amd_only_variables = filterVarsBySampleThreshold(
  amd_only_master_table %>% select(-CaseString),
  num_samples_threshold,
  observational_variables
  )

control_only_variables = filterVarsBySampleThreshold(
  all_master_table %>% filter(CaseString=='Control') %>% select(-CaseString),
  num_samples_threshold,
  observational_variables
  )

# all_variables = observational_variables
# age_lt_90_variables = all_variables

```


### Create p-val names for contrasts

```{r}
all_contrast_names = makeContrastNames(
  all_variables, 
  all_master_table
  )

age_lt_90_contrast_names = makeContrastNames(
  age_lt_90_variables, 
  all_master_table
  )

amd_only_contrast_names = makeContrastNames(
  amd_only_variables, 
  all_master_table
  )

control_only_contrast_names = makeContrastNames(
  control_only_variables, 
  all_master_table
  )
```

### Initialize regression stats table

```{r}
regression_pvals = makeRegressionStatsContrastTemplate(all_contrast_names)

regression_effect_sizes = makeRegressionStatsContrastTemplate(all_contrast_names)
```

## Linear Regression {.tabset}

### All

```{r}
exp_vars = c(raw_exp_vars, calculated_exp_vars)
predictors = all_variables
master_table = all_master_table
contrasts = all_contrast_names
index_name_template = 'All_'

# regression_pvals = regression_pvals_template
# regression_effect_sizes = regression_pvals_template

for (var in exp_vars)
{
  index_name = paste0(index_name_template, var)
  print(index_name)
  
  fit = doMultipleRankRegression(
    var,
    predictors,
    master_table
  )
  
  # print(vif(fit))
  
  pvals = getRegressionPvals(
    fit,
    contrasts,
    index_name
  )
  
  regression_pvals = regression_pvals %>% add_row(!!!pvals)

  effect_sizes = getRegressionEffectSizes(
    fit,
    contrasts,
    index_name
  )

  regression_effect_sizes = regression_effect_sizes %>% add_row(!!!effect_sizes)

}
```

### Age < 90

```{r}
exp_vars = c(raw_exp_vars, calculated_exp_vars)
predictors = age_lt_90_variables
master_table = age_lt_90_master_table
contrasts = age_lt_90_contrast_names %>% unlist()
index_name_template = 'Age_lt_90_'

# regression_pvals = regression_pvals_template
# regression_effect_sizes = regression_pvals_template

for (var in exp_vars)
{
  index_name = paste0(index_name_template, var)
  print(index_name)
  
  fit = doMultipleRankRegression(
    var,
    predictors,
    master_table,
    transform="${regression_transform}",
    log_regularizer=${}
  )
  
  pvals = getRegressionPvals(
    fit,
    contrasts,
    index_name
  )
  
  regression_pvals = regression_pvals %>% add_row(!!!pvals)

  effect_sizes = getRegressionEffectSizes(
    fit,
    contrasts,
    index_name
  )

  regression_effect_sizes = regression_effect_sizes %>% add_row(!!!effect_sizes)

}
```


### AMD Only

```{r}
exp_vars = c(raw_exp_vars, calculated_exp_vars)
predictors = amd_only_variables
master_table = amd_only_master_table
contrasts = amd_only_contrast_names %>% unlist()
index_name_template = 'AMD_Only_'

# regression_pvals = regression_pvals_template
# regression_effect_sizes = regression_pvals_template

for (var in exp_vars)
{
  index_name = paste0(index_name_template, var)
  print(index_name)
  
  fit = doMultipleRankRegression(
    var,
    predictors,
    master_table
  )
  
  pvals = getRegressionPvals(
    fit,
    contrasts,
    index_name
  )
  
  regression_pvals = regression_pvals %>% add_row(!!!pvals)

  effect_sizes = getRegressionEffectSizes(
    fit,
    contrasts,
    index_name
  )

  regression_effect_sizes = regression_effect_sizes %>% add_row(!!!effect_sizes)

}
```
## Visualize
```{r}
pheatmap(regression_pvals %>% column_to_rownames('Index'), cluster_cols=F, cluster_rows=F)

pheatmap(regression_effect_sizes %>% column_to_rownames('Index'), cluster_cols=F, cluster_rows=F)
```


## Write Tables
```{r}
regression_pvalues = createWorkbook()

addWorksheet(
  regression_pvalues, 
  sheetName="Unadjusted P-Values"
  )

openxlsx::writeData(
  regression_pvalues, 
  sheet="Unadjusted P-Values", 
  regression_pvals
  )

addWorksheet(
  regression_pvalues, 
  sheetName="Effect Sizes"
  )

openxlsx::writeData(
  regression_pvalues,
  sheet="Effect Sizes", 
  regression_effect_sizes
  )

# addWorksheet(regression_pvals, sheetName="Adjusted P-Values")
# openxlsx::writeData(regression_pvals, sheet="Adjusted P-Values", adjusted_regression_pvals)

filename = makeDataFileName(
  'linear_regression.xlsx',
  "${tables_dir}",
  "${analysis_type}",
  "${miseq_project_prefix}",
  "${clustering_level}"
)

saveWorkbook(regression_pvalues, file=filename, overwrite=T)
```


```
categorical_data = 
  sample_data %>%
  select(
    Gender,
    CaseString,
    ARMS2rs10490924, 
    CFHrs1061170, 
    CFHrs10737680, 
    SKIV2Lrs429608, 
    CNV_Either_Eye, 
    GA_No_CNV_Either_Eye, 
    AREDS
  ) 

# %>%
#   mutate_all(list(~ifelse(is.na(.), 0, 1))) 
```


```
num_vars = length(categorical_variable_names)

categorical_chisq_pvals = matrix(rep(0, num_vars*num_vars), ncol=num_vars, nrow=num_vars)
colnames(categorical_chisq_pvals) = categorical_variable_names
rownames(categorical_chisq_pvals) = categorical_variable_names

for (var in categorical_variable_names)
{
  for(other_var in categorical_variable_names)
  {
    print(sprintf('var: %s, other_var: %s', var, other_var))
    
    if (var != other_var)
    {
      test =
        sample_data %>%
        select(var, other_var) %>%
        table() %>%
        chisq.test()

      categorical_chisq_pvals[var, other_var] = round(test$p.value,3)
    }

  }
}
  
```




```
## Logistic Regression
### All Subgroups
#### Logistic Predictor Vars

num_samples_threshold = 5

all_variables = filterVarsBySampleThreshold2(
  all_master_table,
  num_samples_threshold,
  observational_variables
  )

vars = setdiff(all_variables %>% names(), c('CaseString', 'AREDS'))
```

```
#### Run Regression

all_logistic_regression_stats = getLogisticRegressionStats(
  base_predictor_vars=observational_variables[vars],
  additional_predictors=c(raw_exp_vars, calculated_exp_vars),
  logistic_response_var='CaseString',
  master_table=all_master_table,
  stats=c('pvalue', 'effect_size', 'std_error'),
  response_var_name='Index'
)

contrasts = makeContrastNames(observational_variables[vars], all_master_table)

logistic_regression_stats = data.frame(ResponseVar=character())
for (contrast in contrasts)
{
  logistic_regression_stats %<>% mutate(!!contrast := numeric())
}
for (var in c(raw_exp_vars, calculated_exp_vars))
{
  logistic_regression_stats %<>% mutate(!!var := numeric())
}

logistic_regression_pvalues = logistic_regression_stats
logistic_regression_odds_ratios = logistic_regression_stats

for(var in names(all_logistic_regression_stats))
{
  
  temp = 
    all_logistic_regression_stats[[var]] %>%
    select(contrast, pvalue, effect_size) %>%
    column_to_rownames('contrast') %>%
    t() %>%
    data.frame() %>%
    rownames_to_column('stat') %>%
    mutate(ResponseVar='CaseString')
  
  pvalues = temp %>% filter(stat=='pvalue') %>% select(-stat)
  odds_ratios = temp %>% filter(stat=='effect_size') %>% select(-stat)
    
  logistic_regression_pvalues %<>% add_row(!!!pvalues)
  logistic_regression_odds_ratios %<>% add_row(!!!odds_ratios)
}

```

