---
title: "MiSeq-${miseq_run_number}"
output:
  html_notebook:
    df_print: paged
    number_sections: yes
    theme: lumen
    toc: yes
    code_folding: hide
---

<style>
  html {
    font-size: 16pt;
  }
  
  body {
    font-size: 16pt;
  }
  
  h1 {
    font-size: 2.2rem;
  }
  
  h2 {
    font-size: 2rem;
  }
  
  h3 {
    font-size: 1.8rem;
  }
  
  h4 {
    font-size: 1.4rem;
  }
  
</style>


```{r}
source('${project_metadata_file}')
analysis_type = "${analysis_type}"
clustering_level = "${clustering_level}"
tables_dir = "${tables_dir}"
analysis_title = "${analysis_title}"
```

# MiSeq-${miseq_run_number} ${analysis_title} Linear Regression Analysis {.tabset}
## Clustering Level
Data clustered at the `r toupper('${clustering_level}')` Level.

## Filtering
* **Relative Abundance Cutoff:** ${relative_abundance_cutoff}
* **Prevalence Cutoff:** ${prevalence_cutoff}
* **Min Count Cutoff:** ${min_count_cutoff}

## Setup
### Start Conda ENV
```{r}
startCondaEnv('regression')
```

### Load Libraries
```{r}
library(rmarkdown)
library(tidyverse)
library(magrittr)
library(ggplot2)
library(ggbeeswarm)
library(openxlsx)
library(DT)
library(pheatmap)
library(kableExtra)
library(vegan)
library(nlme)

### Custom libraries that can be loaded from GitHub
source('~/utilities/analysis-utilities/general_asv_data_analysis_utilities.R')
source('~/utilities/analysis-utilities/alpha_div_utilities.R')
source('~/utilities/amd_templates/setup/amd_project_utilities.R')
```


### Set Knitr Options
```{r}
${knitr_options}
```

### Load Data
```{r}
filename = makeDataFileName(
  'master_table.xlsx',
  "${tables_dir}",
  "${analysis_type}",
  "${miseq_project_prefix}",
  "${clustering_level}"
)

print(filename)

all_master_table = readWorkbook(filename, sheet='Master Table')
age_lt_90_master_table = all_master_table %>% filter(Age < 90)
amd_only_master_table = all_master_table %>% filter(CaseString == 'AMD')
filtered_taxa = readWorkbook(filename, sheet="${tools::toTitleCase(clustering_level)}") %>% pull("short_glommed_taxa")


source(file.path(
  "${project_root}", 
  "${metadata_dir}", 
  'observational_variables.R'
  ))
```

### Variables of interest
#### All Observational Variables
```{r}
observational_variable_names = observational_variables %>% names()
print(observational_variables %>% names())
```
#### Selected Variables
```{r}

```

#### Calculated Experimental Variables

```{r}
raw_exp_vars = ${raw_exp_vars}
calculated_exp_vars = ${calculated_exp_vars}
```
**Calculated Experimental Variables: ** `r calculated_exp_vars`

### Set Metadata Factor Levels
```{r}
all_master_table = setFactorsLevels(all_master_table, observational_variables)
age_lt_90_master_table = setFactorsLevels(age_lt_90_master_table, observational_variables)
amd_only_master_table = setFactorsLevels(amd_only_master_table, observational_variables)
```


### Determine which covariates have sufficient samples with each value
```{r}
num_samples_threshold = 5

all_variables = filterVarsBySampleThreshold(
  all_master_table,
  num_samples_threshold,
  observational_variables
  )

age_lt_90_variables = filterVarsBySampleThreshold(
  age_lt_90_master_table,
  num_samples_threshold,
  observational_variables
  )

amd_only_variables = filterVarsBySampleThreshold(
  amd_only_master_table %>% select(-CaseString),
  num_samples_threshold,
  observational_variables
  )

control_only_variables = filterVarsBySampleThreshold(
  all_master_table %>% filter(CaseString=='Control') %>% select(-CaseString),
  num_samples_threshold,
  observational_variables
  )

# all_variables = observational_variables
# age_lt_90_variables = all_variables

```


### Create p-val names for contrasts

```{r}
all_contrast_names = makeContrastNames(
  all_variables, 
  all_master_table
  )

age_lt_90_contrast_names = makeContrastNames(
  age_lt_90_variables, 
  all_master_table
  )

amd_only_contrast_names = makeContrastNames(
  amd_only_variables, 
  all_master_table
  )

control_only_contrast_names = makeContrastNames(
  control_only_variables, 
  all_master_table
  )
```

## Response Variables

```{r}
raw_exp_vars = ${raw_exp_vars}
calculated_exp_vars = ${calculated_exp_vars}
all_experimental_variables = c(raw_exp_vars, calculated_exp_vars)
```
**Raw Experimental Variables: ** `r raw_exp_vars`
**Calculated Experimental Variables: ** `r calculated_exp_vars`

## Regression analysis

### All

```{r}
all_linear_regression_stats = getMultipleRegressionStats(
  predictor_vars=all_variables,
  response_vars=all_experimental_variables,
  master_table=all_master_table,
  stats=${regression_stats},
  response_var_name=response_var_name,
  adjustment_method=adjustment_method,
  transformation='${taxa_transform}',
  log_regularizer=${relative_abundance_cutoff}
)

```


### Age < 90
```{r}
age_lt_90_linear_regression_stats = getMultipleRegressionStats(
  predictor_vars=age_lt_90_variables,
  response_vars=all_experimental_variables,
  master_table=all_master_table %>% filter(Age < 90),
  stats=stats,
  response_var_name=response_var_name,
  adjustment_method=adjustment_method,
  transformation='${taxa_transform}',
  log_regularizer=${relative_abundance_cutoff}
)


```

### AMD Only
```{r}
amd_only_regression_stats = getMultipleRegressionStats(
  predictor_vars=amd_only_variables,
  response_vars=all_experimental_variables,
  master_table=all_master_table %>% filter(CaseString=='AMD'),
  stats=stats,
  response_var_name=response_var_name,
  adjustment_method=adjustment_method,
  transformation='${taxa_transform}',
  log_regularizer=${relative_abundance_cutoff}
)
```

## Write Regression Stats
```{r}

wb = createWorkbook()

regression_stats_table_names = c(
  "all_regression_stats",
  "age_lt_90_regression_stats",
  "amd_only_regression_stats"
  )

contrast_names = makeContrastNames(
  observational_variables,
  all_master_table
)


for (tab_name in regression_stats_table_names)
{
  contrast_regression_stats = createWorkbook()
  
  print(tab_name)
  stats_table_name = paste0(tab_name, "_regression_stats")
  print(stats_table_name)
  stats_table = get(stats_table_name)
  
  filename = makeDataFileName(
    paste0(tab_name, '_linear_regression.xlsx'),
    "${tables_dir}",
    "${analysis_type}",
    "${miseq_project_prefix}",
    "${clustering_level}"
  )
  print(filename)
  
  saveWorkbook(contrast_regression_stats, file=filename, overwrite=T)
  
}
```