---
title: "MiSeq-${miseq_run_number}"
output:
  html_notebook:
    df_print: paged
    number_sections: yes
    theme: lumen
    toc: yes
    code_folding: hide
---

<style>
  html {
    font-size: 16pt;
  }
  
  body {
    font-size: 16pt;
  }
  
  h1 {
    font-size: 2.2rem;
  }
  
  h2 {
    font-size: 2rem;
  }
  
  h3 {
    font-size: 1.8rem;
  }
  
  h4 {
    font-size: 1.4rem;
  }
  
  .dataTable {
    font-size: 9pt;
  }
  
</style>

```{r}
source('${project_metadata_file}')
analysis_type = "${analysis_type}"
clustering_level = "${clustering_level}"
tables_dir = "${tables_dir}"
use_allbac = ${use_allbac}
analysis_title = tools::toTitleCase(gsub("_", " ", analysis_type))
```

# MiSeq-${miseq_run_number} DESeq2 Analysis {.tabset}

## Clustering
Data clustered at the `r toupper("${clustering_level}")` level.

## Filtering
* **Relative Abundance Cutoff:** ${relative_abundance_cutoff}
* **Prevalence Cutoff:** ${prevalence_cutoff}
* **Min Count Cutoff:** ${min_count_cutoff}


## Setup 

### [Close]

### Start Conda ENV
```{r}
startCondaEnv('deseq', lib='~/R35')
```

### Load Libraries
```{r}
library(rmarkdown)
library(tidyverse)
library(magrittr)
library(ggplot2)
library(openxlsx)
library(DT)
library(kableExtra)
library(DESeq2)

### Custom libraries that can be loaded from GitHub
source('~/utilities/analysis-utilities/general_asv_data_analysis_utilities.R')
source('~/utilities/analysis-utilities/deseq_utilities.R')
source('~/utilities/amd_templates/setup/amd_project_utilities.R')
```

### Set Knitr Options
```{r}
${knitr_options}
```

### Load Data
```{r}
pathway_counts_table_file = file.path(
  "${tables_dir}",
  "picrust_out",
  "pathways_out",
  "miseq-${miseq_run_number}_path_abun_unstrat.xlsx"
)

pathway_counts_table = read.xlsx(
  pathway_counts_table_file,
  colNames=T
  )

sample_data = read.xlsx(
  '${sample_data_file}',
  colNames=T
  )

if ("IGA" %in% colnames(sample_data) & use_allbac)
{
  sample_data = 
    sample_data %>%
    filter(IGA=='AllBac')

  sample_names = sample_data$SampleName
  
  asv_table = 
    asv_table %>%
    select('ASVs', ends_with('AllBac'))
}

sample_data_columns = colnames(sample_data)
sample_names = sample_data$SampleName

source(file.path(
  "${project_root}", 
  "${metadata_dir}", 
  'observational_variables.R'
  ))

```

### Integerize pathway counts
```{r}
pathway_counts_table = 
  pathway_counts_table %>%
  remove_rownames() %>%
  mutate_at(vars(sample_names), ~floor(.+0.00000001))
```


### Set Metadata Factor Levels
```{r}
sample_data = setFactorsLevels(sample_data, observational_variables)
# sample_data %<>%
#   mutate(AREDS=
#     ifelse(is.na(AREDS), "MISSING", AREDS) %>%
#     factor(levels=c('N', 'Y', 'MISSING'))
#   )
```

### Variables of interest
#### Observational Variables
```{r}
all_covariates = observational_variables[${linear_regression_predictors['All']}]
age_lt_90_covariates = observational_variables[${linear_regression_predictors['Age_lt_90']}]
genotypes = ${genotypes}
phenotypes = ${phenotypes}
```

#### Remove Age
```{r}
all_covariates$Age = NULL
age_lt_90_covariates$Age = NULL
```

* **All Variables:** `r all_covariates %>% names()`
* **Age < 90 Variables:** `r age_lt_90_covariates %>% names()`

## All Subgroups Case/Control Only
### Create Workbook
```{r}
deseq_results_wb = createWorkbook()
```

#### Run and add worksheets
```{r}
results_df = runDeseqFromPathwayCounts(
  pathway_counts=pathway_counts_table,
  sample_data,
  variables="CaseString",
  include_covariates = 'All'
  )

padj_col = "CaseString_AMD_vs_Control.padj"

results_df =
  results_df %>%
  select(pathway, !!padj_col, everything()) %>%
  arrange(!!as.name(padj_col))

addWorksheet(
  wb=deseq_results_wb, 
  sheetName=""
  )

openxlsx::writeData(
  wb=deseq_results_wb, 
  sheet="", 
  x=results_df
  )
  
setColWidths(
  wb=deseq_results_wb, 
  sheet="", 
  cols=1, 
  widths = "auto"
)

freezePane(
  deseq_results_wb, 
  sheet="", 
  firstCol=T,
  firstRow=T
)
```

#### Save Workbook
```{r}
filename = makeDataFileName(
  'all_subgroups_case_vs_control.xlsx',
  "${tables_dir}",
  "${analysis_type}",
  "${miseq_project_prefix}",
  "rank"
)

print(filename)

saveWorkbook(deseq_results_wb, file=filename, overwrite=T)
```

## All Subgroups -- All Variables
### Create Workbook
```{r}
deseq_results_wb = createWorkbook()
```

### Run and add worksheets
```{r}
results_df = runDeseqFromPathwayCounts(
  pathway_counts=pathway_counts_table,
  sample_data,
  variables=all_covariates %>% names(),
  include_covariates = 'All'
  )
  
padj_col = "CaseString_AMD_vs_Control.padj"
  
results_df =
  results_df %>%
  select(pathway, !!padj_col, starts_with("CaseString"), everything()) %>%
  arrange(!!as.name(padj_col))

addWorksheet(
  wb=deseq_results_wb, 
  sheetName=""
  )

openxlsx::writeData(
  wb=deseq_results_wb, 
  sheet="", 
  x=results_df
  )
  
setColWidths(
  wb=deseq_results_wb, 
  sheet="", 
  cols=1, 
  widths = "auto"
)

freezePane(
  deseq_results_wb, 
  sheet="", 
  firstCol=T,
  firstRow=T
)
```

### Save Workbook
```{r}
filename = makeDataFileName(
  'all_subgroups_with_covariates.xlsx',
  "${tables_dir}",
  "${analysis_type}",
  "${miseq_project_prefix}",
  "rank"
)

print(filename)

saveWorkbook(deseq_results_wb, file=filename, overwrite=T)
```


## Age < 90 Case/Control Only
### Create Workbook
```{r}
deseq_results_wb = createWorkbook()
```

#### Run and add worksheets
```{r}
results_df = runDeseqFromCounts(
  counts_table=pathway_counts_table,
  sample_data %>% filter(Age < 90),
  variables="CaseString",
  include_covariates = 'All'
  )

padj_col = "CaseString_AMD_vs_Control.padj"

results_df =
  results_df %>%
  select(pathway, !!padj_col, everything()) %>%
  arrange(!!as.name(padj_col))

addWorksheet(
  wb=deseq_results_wb, 
  sheetName=""
  )

openxlsx::writeData(
  wb=deseq_results_wb, 
  sheet="", 
  x=results_df
  )
  
setColWidths(
  wb=deseq_results_wb, 
  sheet="", 
  cols=1, 
  widths = "auto"
)

freezePane(
  deseq_results_wb, 
  sheet="", 
  firstCol=T,
  firstRow=T
)
```

#### Save Workbook
```{r}
filename = makeDataFileName(
  'all_subgroups_case_vs_control.xlsx',
  "${tables_dir}",
  "${analysis_type}",
  "${miseq_project_prefix}",
  "rank"
)

print(filename)

saveWorkbook(deseq_results_wb, file=filename, overwrite=T)
```

## Age < 90 -- All Variables
### Create Workbook
```{r}
deseq_results_wb = createWorkbook()
```

### Run and add worksheets
```{r}
results_df = runDeseqFromCounts(
  counts_table=pathway_counts_table,
  sample_data %>% filter(Age < 90),
  variables=age_lt_90_covariates %>% names(),
  include_covariates = 'All'
  )
  
padj_col = "CaseString_AMD_vs_Control.padj"
  
results_df =
  results_df %>%
  select(pathway, !!padj_col, starts_with("CaseString"), everything()) %>%
  arrange(!!as.name(padj_col))

addWorksheet(
  wb=deseq_results_wb, 
  sheetName=""
  )

openxlsx::writeData(
  wb=deseq_results_wb, 
  sheet="", 
  x=results_df
  )
  
setColWidths(
  wb=deseq_results_wb, 
  sheet="", 
  cols=1, 
  widths = "auto"
)

freezePane(
  deseq_results_wb, 
  sheet="", 
  firstCol=T,
  firstRow=T
)
```

### Save Workbook
```{r}
filename = makeDataFileName(
  'all_subgroups_with_covariates.xlsx',
  "${tables_dir}",
  "${analysis_type}",
  "${miseq_project_prefix}",
  "rank"
)

print(filename)

saveWorkbook(deseq_results_wb, file=filename, overwrite=T)
```

