---
title: "MiSeq-${miseq_run_number}"
output:
  html_notebook:
    df_print: paged
    number_sections: yes
    theme: lumen
    toc: yes
    code_folding: hide
---

<style>
  html {
    font-size: 16pt;
  }
  
  body {
    font-size: 16pt;
  }
  
  h1 {
    font-size: 2.2rem;
  }
  
  h2 {
    font-size: 2rem;
  }
  
  h3 {
    font-size: 1.8rem;
  }
  
  h4 {
    font-size: 1.4rem;
  }
  
</style>

```{r}
source('${project_metadata_file}')
analysis_type = "${analysis_type}"
clustering_level = "${clustering_level}"
tables_dir = "${tables_dir}"
use_allbac = ${use_allbac}
```

# MiSeq-${miseq_run_number} PICRUSt2 Data Prep{.tabset}

## Filtering

```{r}
relative_abundance_cutoff = ${relative_abundance_cutoff}
prevalence_cutoff = ${prevalence_cutoff}
min_count_cutoff = ${min_count_cutoff}
```
**relative abundance cutoff:** ${relative_abundance_cutoff}
**prevalence cutoff:** ${prevalence_cutoff}
**min count cutoff:** ${min_count_cutoff}

## Setup 

### [Close]

### Start Conda ENV
```{r}
startCondaEnv('picrust2')
```

### Load Libraries
```{r}
library(rmarkdown)
library(tidyverse)
library(magrittr)
library(ggplot2)
library(ggbeeswarm)
library(openxlsx)
library(DT)
library(pheatmap)
library(kableExtra)
library(stringr)

### Custom libraries that can be loaded from GitHub
source('~/utilities/analysis-utilities/general_asv_data_analysis_utilities.R')
source('~/utilities/amd_templates/setup/amd_project_utilities.R')
```

### Set Knitr Options
```{r}
${knitr_options}
```

### Load Data
```{r}
asv_table = 
  read.delim(
    '${asv_table_file}', 
    header=T, 
    sep='\t',
    stringsAsFactors=F
    )

sample_data = read.xlsx('${sample_data_file}')

if ("IGA" %in% colnames(sample_data) & use_allbac)
{
  sample_data = 
    sample_data %>%
    filter(IGA=='AllBac')

  sample_names = sample_data$SampleName
  
  asv_table = 
    asv_table %>%
    select('ASVs', ends_with('AllBac'))
}

sample_data_columns = colnames(sample_data)
sample_names = sample_data$SampleName

```

## Filter ASV Table
```{r}
filtered_asv_table = 
  asv_table %>%
  applyMinCountFilter(min_count_cutoff) %>%
  applyRelativeAbundanceFilter(relative_abundance_cutoff) %>%
  applyPrevalenceFilter(prevalence_cutoff) %>%
  select('ASVs', sample_names)
```

## Write ASV Count Table
```{r}
asv_counts_table = makeDataFileName(
  'asv_counts.tsv',
  "${tables_dir}",
  "${analysis_type}",
  "${miseq_project_prefix}",
  "${clustering_level}"
)

write_delim(
  filtered_asv_table,
  path=asv_counts_table,
  delim='\t',
  col_names=T
)
```

## Create ASV Fasta File
### Write ASVs to file
```{r}
asv_list_file = makeDataFileName(
  'asv_list',
  "${tables_dir}",
  "${analysis_type}",
  "${miseq_project_prefix}",
  "${clustering_level}"
)

write_delim(
  filtered_asv_table %>% select('ASVs'), 
  path=asv_list_file,
  col_names=F
  )
```

### Insert Sequence IDs
```{r}
asv_fna_filename = makeDataFileName(
  'asvs.fna',
  "${tables_dir}",
  "${analysis_type}",
  "${miseq_project_prefix}",
  "${clustering_level}"
)

cmd = paste0(
  "awk '{print \">\" $0; print $0}' ",
  asv_list_file,
  " > ",
  asv_fna_filename
  )
print(cmd)
cat(cmd)

system(cmd)
```

