---
title: "MiSeq-${miseq_run_number}"
output:
  html_notebook:
    df_print: paged
    number_sections: yes
    theme: lumen
    toc: yes
    code_folding: hide
---

<style>
  html {
    font-size: 16pt;
  }
  
  body {
    font-size: 16pt;
  }
  
  h1 {
    font-size: 2.2rem;
  }
  
  h2 {
    font-size: 2rem;
  }
  
  h3 {
    font-size: 1.8rem;
  }
  
  h4 {
    font-size: 1.4rem;
  }
  
</style>

```{r}
source('${project_metadata_file}')
analysis_type = "${analysis_type}"
clustering_level = "${clustering_level}"
tables_dir = "${tables_dir}"
use_allbac = ${use_allbac}
```

# MiSeq-${miseq_run_number} PICRUSt2 Analysis {.tabset}

## Setup 

### [Close]

### Start Conda ENV
```{r}
startCondaEnv('picrust2')
```

### Load Libraries
```{r}
library(rmarkdown)
library(tidyverse)
library(magrittr)
library(ggplot2)
library(ggbeeswarm)
library(openxlsx)
library(DT)
library(pheatmap)
library(kableExtra)
library(glue)

### Custom libraries that can be loaded from GitHub
source('~/utilities/analysis-utilities/general_asv_data_analysis_utilities.R')
source('~/utilities/amd_templates/setup/amd_project_utilities.R')
```


### Set Knitr Options
```{r}
${knitr_options}
```

### Load Sample Data
```{r}
sample_data = read.xlsx('${sample_data_file}')

if ("IGA" %in% colnames(sample_data) & use_allbac)
{
  sample_data = 
    sample_data %>%
    filter(IGA=='AllBac')

  sample_names = sample_data$SampleName
  
  asv_table = 
    asv_table %>%
    select('ASVs', ends_with('AllBac'))
}

sample_data_columns = colnames(sample_data)
sample_names = sample_data$SampleName

```


## Run PICRUSt2

### Construct Filenames
```{r}
asv_fna_filename = makeDataFileName(
  'asvs.fna',
  "${tables_dir}",
  "${analysis_type}",
  "${miseq_project_prefix}",
  "${clustering_level}"
)

print(sprintf('asv_fna_filename: %s', asv_fna_filename))

asv_counts_table_filename = makeDataFileName(
  'asv_counts.tsv',
  "${tables_dir}",
  "${analysis_type}",
  "${miseq_project_prefix}",
  "${clustering_level}"
)

print(sprintf('asv_counts_table_filename: %s', asv_counts_table_filename))

```

### Delete existing output directory
```{r}
system('rm -rf ${tables_dir}/picrust_out')
```


### Create Run String
```
${picrust_run_string_template}
```

```{r}
num_procs = ${num_procs}
tables_dir = "${tables_dir}"
src_dir = sub('tables', 'src', tables_dir)

run_string_template = "${picrust_run_string_template}"
run_string = glue(run_string_template)

print("Run String:")
cat(run_string)

run_picrust_file = file.path(src_dir, 'run_picrust.sh')
write(run_string, run_picrust_file)
```

### Run PICRUSt2
```{r}
print("sending run string to system")
system_call = paste0('bash -i ', run_picrust_file)
print(system_call)
system(system_call, intern = FALSE, wait = TRUE, show.output.on.console = TRUE)
```

## Convert Files to XSLX
### Unzip all gzipped files
```{r}
cmd = 'find ${tables_dir} -type f -name "*.gz" -exec gunzip {} +'
print(cmd)

system(cmd)
```

### Get list of TSV Files
```{r}
cmd = 'find ${tables_dir} -name "*.tsv"'
print(cmd)

tsv_files = system(cmd, intern=T)

print("tsv files")
print(tsv_files)
```

### Load TSV files and Save as XLSX
```{r}
for (filename in tsv_files)
{
  
  data = read.delim(
    filename,
    sep='\t',
    stringsAsFactors=F
  )

  base_name = basename(filename)
  dir_name = dirname(filename)
  
  new_filename = file.path(
    dir_name,
    paste0('miseq-${miseq_run_number}_', gsub('tsv', 'xlsx', base_name))
  )
  
  print(new_filename)

  write.xlsx(
    data,
    file=new_filename,
    col.names=T,
    row.names=F
  )

}
```

## Create Master Table
### Load KO Abundances
```{r}
ko_counts_table_file = file.path(
  "${tables_dir}",
  "picrust_out",
  "KO_metagenome_out",
  "miseq-239_pred_metagenome_unstrat.xlsx"
)

ko_counts_table = read.xlsx(
  ko_counts_table_file,
  colNames=T
  )

### Tidy Table
ko_counts_table = 
  ko_counts_table %>%
  remove_rownames() %>%
  # mutate_at(vars(sample_names), ~floor(.+0.00000001)) %>%
  rename(KO=function.)

ec_counts_table_file = file.path(
  "${tables_dir}",
  "picrust_out",
  "EC_metagenome_out",
  "miseq-239_pred_metagenome_unstrat.xlsx"
)

ec_counts_table = read.xlsx(
  ec_counts_table_file,
  colNames=T
  )

### Tidy Table
ec_counts_table = 
  ec_counts_table %>%
  remove_rownames() %>%
  # mutate_at(vars(sample_names), ~floor(.+0.00000001)) %>%
  rename(EC=function.)
```

### Get Master Tables
```{r}
ko_master_table = getMasterTable(
  taxa_counts=ko_counts_table,
  metadata=sample_data,
  key_column='SampleName',
  taxa_colnames='KO'
)

ec_master_table = getMasterTable(
  taxa_counts=ec_counts_table,
  metadata=sample_data,
  key_column='SampleName',
  taxa_colnames='EC'
)
```

### Write Master Tables
```{r}
wb = createWorkbook()

addWorksheet(wb, sheetName="KO Master Table")
openxlsx::writeData(wb, sheet="KO Master Table", ko_master_table)
addWorksheet(wb, sheetName="EC Master Table")
openxlsx::writeData(wb, sheet="EC Master Table", ec_master_table)

just_ko = 
  ko_counts_table %>% 
  select(KO)
addWorksheet(wb, sheetName="KO")
openxlsx::writeData(wb, sheet="KO", just_ko)

just_ec = 
  ec_counts_table %>% 
  select(EC)
addWorksheet(wb, sheetName="EC")
openxlsx::writeData(wb, sheet="EC", just_ec)

filename = makeDataFileName(
  'master_table.xlsx',
  "${tables_dir}",
  "${analysis_type}",
  "${miseq_project_prefix}",
  "${clustering_level}"
)
print(filename)

saveWorkbook(wb, file=filename, overwrite=T)
```










