---
title: "${miseq_project_prefix}: ${analysis_type}"
output:
  html_notebook:
    df_print: paged
    number_sections: yes
    theme: lumen
    toc: yes
    code_folding: hide
---

<style>
  html {
    font-size: 16pt;
  }
  
  body {
    font-size: 16pt;
  }
  
  h1 {
    font-size: 2.2rem;
  }
  
  h2 {
    font-size: 2rem;
  }
  
  h3 {
    font-size: 1.8rem;
  }
  
  h4 {
    font-size: 1.4rem;
  }
  
  .dataTable {
    font-size: 9pt;
  }
  
</style>

```{r}
source('${project_metadata_file}')
analysis_type = "${analysis_type}"
clustering_level = "${clustering_level}"
tables_dir = "${tables_dir}"
analysis_title = tools::toTitleCase(gsub("_", " ", analysis_type))
cooks_cutoff = ${cooks_cutoff}
independent_filtering = ${independent_filtering}
```

# ${miseq_project_prefix} DESeq2 Analysis {.tabset}
**Clustering**
Data clustered at the `r toupper('${clustering_level}')` Level.

**Subgroups**
```{r, results='asis', echo=FALSE}
for (sg_filter in c(${subgroup_filters}))
{
  cat(sg_filter, '\n\n')
}
```

**Filtering**
* **Relative Abundance cutoff:** ${relative_abundance_cutoff}
* **Prevalence Cutoff:** ${prevalence_cutoff}
* **Min Count Cutoff:** ${min_count_cutoff}

```{r}
relative_abundance_cutoff = ${relative_abundance_cutoff}
prevalence_cutoff = ${prevalence_cutoff}
min_count_cutoff = ${min_count_cutoff}
```

## Setup 

### [Close]

### Start Conda ENV
```{r}
startCondaEnv('piphillin', lib='~/R35')
```

### Load Libraries
```{r}
# library(rmarkdown)
library(tidyverse)
library(magrittr)
library(ggplot2)
library(openxlsx)

library(kableExtra)
library(DESeq2)
library(KEGGREST)

### Custom libraries that can be loaded from GitHub
source('${utilities_dir}/analysis-utilities/general_asv_data_analysis_utilities.R')
source('${utilities_dir}/amd_templates/setup/amd_project_utilities.R')
source('${utilities_dir}/analysis-utilities/deseq_utilities.R')

source('${utilities_dir}/analysis-utilities/pathway_utilities.R')
```

### Set Knitr Options
```{r}
${knitr_options}
```

### Variables of interest
#### Observational Variables
```{r}
case = ${case} %>% setdiff('')
treatments = ${treatments} %>% setdiff('')
phenotypes = ${phenotypes} %>% setdiff('')
genotypes = ${genotypes} %>% setdiff('')
key_covariates = ${key_covariates} %>% setdiff('')

linear_predictors = ${linear_predictors} %>% setdiff('')
categorical_variables = ${categorical_variables} %>% setdiff('')
amd_only_variables = linear_predictors %>% setdiff(case)
```

* **Case:** `r case`
* **Key Covariates:** `r key_covariates`
* **Categorical Variables:**  `r categorical_variables`


### Load Data
```{r}
sample_data = read.xlsx(
  '${sample_data_file}',
  colNames=T
  )

sample_data_columns = colnames(sample_data)
sample_names = sample_data$SampleName

kegg_pathway_counts_file = file.path(
  "${tables_dir}",
  "kegg_pathway_counts.xlsx"
)

kegg_pathway_counts = read.xlsx(
  kegg_pathway_counts_file,
  colNames=T
  ) %>% 
  remove_rownames() %>%
  mutate_at(vars(-PathwayName), ~as.integer(.))

```

### Load Master Table
```{r}
master_table_filename = makeDataFileName(
  'master_table.xlsx',
  "${tables_dir}",
  "${analysis_type}",
  "${miseq_project_prefix}",
  "${clustering_level}"
)
print(master_table_filename)
pathway_name_map = read.xlsx(
  master_table_filename, 
  sheet='Pathway Names'
  )

source(file.path(
  "${project_root}", 
  "${metadata_dir}", 
  'observational_variables.R'
  ))

```

### Extract Subgroups
```{r}
if (!is.null(c(${subgroup_filters})))
{
  subgroup_filter_string = paste0(${subgroup_filters}, collapse=" & ")
  print(sprintf("Subgroup Filter String: %s", subgroup_filter_string))
  subgroup_filter_expression = str2expression(subgroup_filter_string)
  sample_data = sample_data %>% filter(eval(subgroup_filter_expression))
  sample_names = sample_data$SampleName

  kegg_pathway_counts = 
    kegg_pathway_counts %>% 
    select(PathwayName, sample_names)
}

```


### Set Metadata Factor Levels
```{r}
sample_data = setFactorsLevels(sample_data, observational_variables)
# sample_data %<>%
#   mutate(AREDS=
#     ifelse(is.na(AREDS), "MISSING", AREDS) %>%
#     factor(levels=c('N', 'Y', 'MISSING'))
#   )
```

## All Subgroups Case/Control Only
### Create Workbook
```{r}
deseq_results_wb = createWorkbook()
```

#### Run and add worksheets
```{r}
results_df = runDeseqFromPathwayCounts(
  pathway_counts=kegg_pathway_counts,
  sample_data,
  variables="CaseString",
  include_covariates = 'All',
  pathway_colname='PathwayName',
  cooks_cutoff=cooks_cutoff,
  independent_filtering=independent_filtering
  )

padj_col = "CaseString_AMD_vs_Control.padj"

results_df =
  results_df %>%
  arrange(!!as.name(padj_col)) %>%
  left_join(kegg_pathway_counts %>% select(PathwayName), by='PathwayName') %>%
  select(PathwayName, !!padj_col, everything())

addWorksheet(
  wb=deseq_results_wb, 
  sheetName="Sheet"
  )

openxlsx::writeData(
  wb=deseq_results_wb, 
  sheet="Sheet", 
  x=results_df
  )
```

#### Save Workbook
```{r}
filename = makeDataFileName(
  'all_subgroups_case_vs_control.xlsx',
  "${tables_dir}",
  "${analysis_type}",
  "${miseq_project_prefix}",
  "${clustering_level}"
)

print(filename)

saveWorkbook(deseq_results_wb, file=filename, overwrite=T)
```

## All Subgroups -- All Variables
### Create Workbook
```{r}
deseq_results_wb = createWorkbook()
```

### Run and add worksheets
```{r}
sample_data_no_na = 
  sample_data %>% 
  filter(
    paste0('!is.na(', categorical_variables, ')', collapse=' & ') %>%
    str2expression() %>%
    eval()
  )

results_df = runDeseqFromPathwayCounts(
  pathway_counts=kegg_pathway_counts,
  sample_data_no_na,
  variables=c(case, key_covariates),
  include_covariates = 'All',
  pathway_colname='PathwayName',
  cooks_cutoff=cooks_cutoff,
  independent_filtering=independent_filtering
  )
  
padj_col = "CaseString_AMD_vs_Control.padj"
  
results_df =
  results_df %>%
  arrange(!!as.name(padj_col)) %>%
  left_join(kegg_pathway_counts %>% select(PathwayName), by='PathwayName') %>%
  select(PathwayName, !!padj_col, everything())

addWorksheet(
  wb=deseq_results_wb, 
  sheetName="Sheet"
  )

openxlsx::writeData(
  wb=deseq_results_wb, 
  sheet="Sheet", 
  x=results_df
  )
```

### Save Workbook
```{r}
filename = makeDataFileName(
  'all_subgroups_with_covariates.xlsx',
  "${tables_dir}",
  "${analysis_type}",
  "${miseq_project_prefix}",
  "${clustering_level}"
)

print(filename)

saveWorkbook(deseq_results_wb, file=filename, overwrite=T)
```


## Age < 90 Case/Control Only
### Create Workbook
```{r}
deseq_results_wb = createWorkbook()
```

#### Run and add worksheets
```{r}
results_df = runDeseqFromPathwayCounts(
  pathway_counts=kegg_pathway_counts,
  sample_data %>% filter(Age < 90),
  variables="CaseString",
  include_covariates = 'All',
  pathway_colname='PathwayName',
  cooks_cutoff=cooks_cutoff,
  independent_filtering=independent_filtering
  )

padj_col = "CaseString_AMD_vs_Control.padj"

results_df =
  results_df %>%
  arrange(!!as.name(padj_col)) %>%
  left_join(kegg_pathway_counts %>% select(PathwayName), by='PathwayName') %>%
  select(PathwayName, !!padj_col, everything())


addWorksheet(
  wb=deseq_results_wb, 
  sheetName="Sheet"
  )

openxlsx::writeData(
  wb=deseq_results_wb, 
  sheet="Sheet", 
  x=results_df
  )
```

#### Save Workbook
```{r}
filename = makeDataFileName(
  'age_lt_90_case_vs_control.xlsx',
  "${tables_dir}",
  "${analysis_type}",
  "${miseq_project_prefix}",
  "${clustering_level}"
)

print(filename)

saveWorkbook(deseq_results_wb, file=filename, overwrite=T)
```

## Age < 90 -- All Variables
### Create Workbook
```{r}
deseq_results_wb = createWorkbook()
```

### Run and add worksheets
```{r}
sample_data_no_na = 
  sample_data %>% 
  filter(
    paste0('!is.na(', categorical_variables, ')', collapse=' & ') %>%
    str2expression() %>%
    eval()
  )

results_df = runDeseqFromPathwayCounts(
  pathway_counts=kegg_pathway_counts,
  sample_data_no_na %>% filter(Age < 90),
  variables=c(case, key_covariates),
  include_covariates = 'All',
  pathway_colname='PathwayName',
  cooks_cutoff=cooks_cutoff,
  independent_filtering=independent_filtering
  )
  
padj_col = "CaseString_AMD_vs_Control.padj"
  
results_df =
  results_df %>%
  arrange(!!as.name(padj_col)) %>%
    left_join(kegg_pathway_counts %>% select(PathwayName), by='PathwayName') %>%
  select(PathwayName, !!padj_col, everything())


addWorksheet(
  wb=deseq_results_wb, 
  sheetName="Sheet"
  )

openxlsx::writeData(
  wb=deseq_results_wb, 
  sheet="Sheet", 
  x=results_df
  )

```

### Save Workbook
```{r}
filename = makeDataFileName(
  'age_lt_90_with_covariates.xlsx',
  "${tables_dir}",
  "${analysis_type}",
  "${miseq_project_prefix}",
  "${clustering_level}"
)
print(filename)

saveWorkbook(deseq_results_wb, file=filename, overwrite=T)
```
















