---
title: "${miseq_project_prefix}: ${analysis_type}"
output:
  html_notebook:
    df_print: paged
    number_sections: yes
    theme: lumen
    toc: yes
    code_folding: hide
---

<style>
  html {
    font-size: 16pt;
  }
  
  body {
    font-size: 16pt;
  }
  
  h1 {
    font-size: 2.2rem;
  }
  
  h2 {
    font-size: 2rem;
  }
  
  h3 {
    font-size: 1.8rem;
  }
  
  h4 {
    font-size: 1.4rem;
  }
  
  .dataTable {
    font-size: 9pt;
  }
  
</style>

```{r}
source('${project_metadata_file}')
analysis_type = "${analysis_type}"
clustering_level = "${clustering_level}"
tables_dir = "${tables_dir}"
analysis_title = tools::toTitleCase(gsub("_", " ", analysis_type))
cooks_cutoff = ${cooks_cutoff}
independent_filtering = ${independent_filtering}
```

# ${miseq_project_prefix} DESeq2 Analysis {.tabset}
**Clustering**
Data clustered at the `r toupper('${clustering_level}')` Level.

**Subgroups**
```{r, results='asis', echo=FALSE}
for (sg_filter in c(${subgroup_filters}))
{
  cat(sg_filter, '\n\n')
}
```


## Setup 

### [Close]

### Start Conda ENV
```{r}
startCondaEnv('piphillin', lib='~/R35')
```

### Load Libraries
```{r}
# library(rmarkdown)
library(tidyverse)
library(magrittr)
library(openxlsx)
library(KEGGREST)

### Custom libraries that can be loaded from GitHub
source('~/utilities/analysis-utilities/general_asv_data_analysis_utilities.R')
source('~/utilities/amd_templates/setup/amd_project_utilities.R')
source('~/utilities/analysis-utilities/pathway_utilities.R')
```

### Set Knitr Options
```{r}
${knitr_options}
```


### Load Sample Data
```{r}
sample_data = read.xlsx(
  '${sample_data_file}',
  colNames=T
  )

sample_data_columns = colnames(sample_data)
sample_names = sample_data$SampleName
```


### Extract Subgroups
```{r}
if (!is.null(c(${subgroup_filters})))
{
  subgroup_filter_string = paste0(${subgroup_filters}, collapse=" & ")
  print(sprintf("Subgroup Filter String: %s", subgroup_filter_string))
  subgroup_filter_expression = str2expression(subgroup_filter_string)
  sample_data = sample_data %>% filter(eval(subgroup_filter_expression))
  sample_names = sample_data$SampleName
}
```


### Load Raw Pathway Abundance
```{r}
piphillin_results_dir = "${piphillin_results_dir}"
print(piphillin_results_dir)

ko_abundance_table_file = file.path(
  "${piphillin_results_dir}",
  "ko_pathway_abund_table_unnorm.txt"
  )

ko_abundance_table = read.delim(
  ko_abundance_table_file,
  header=T,
  sep='\t'
  ) %>% remove_rownames()


ko_abundance_table %<>%
  select(Pathway, sample_names)
```

### Filter pathway abundance
**Pathway relative abundance cutoff:** ${pathway_relative_abundance_cutoff}
**Pathway prevalence filter:** ${pathway_prevalence_cutoff}
```{r}
ko_abundance_table = 
  ko_abundance_table %>%
  applyRelativeAbundanceFilter(${pathway_relative_abundance_cutoff}) %>%
  applyPrevalenceFilter(${pathway_prevalence_cutoff})
```


### Extract Subgroups
```{r}
if (!is.null(c(${subgroup_filters})))
{
  subgroup_filter_string = paste0(${subgroup_filters}, collapse=" & ")
  print(sprintf("Subgroup Filter String: %s", subgroup_filter_string))
  subgroup_filter_expression = str2expression(subgroup_filter_string)
  sample_data = sample_data %>% filter(eval(subgroup_filter_expression))
  sample_names = sample_data$SampleName

  ko_abundance_table = 
    ko_abundance_table %>% 
    select('Pathway', sample_names)
}
```

### Get KEGG Pathway Names
```{r}
kegg_pathway_abundance = 
  ko_abundance_table %>%
  mutate(PathwayName = getKEGGPathways(Pathway))

pathway_name_map = 
  kegg_pathway_abundance %>%
  select(Pathway, PathwayName)
```

### Integerize pathway counts and fix pathway names
```{r}
### Make PathwayNames unique and suitable for column names
kegg_pathway_counts =
  kegg_pathway_abundance %>%
  mutate(PathwayName =
           PathwayName %>%
           str_replace_all(' ', '_') %>%
           str_replace_all('/', '_slash_') %>%
           str_replace_all('-', '_dash_') %>%
           str_replace_all('\\(', '_paren_') %>%
           str_replace_all('\\)', '_unparen_') %>%
           str_replace_all('\\.', '_dot_') %>%
           make.names()
    ) %>%
  select(-Pathway) %>%
  group_by(PathwayName) %>%
  summarise_all(sum) %>%
  mutate_at(vars(sample_names), ~floor(.+1))

# kegg_pathway_counts = 
#   ko_abundance_table %>%
#   group_by(Pathway) %>%
#   summarise_all(sum) %>%
#   mutate_at(vars(sample_names), ~floor(.+1))
```

### Save as XLSX
```{r}
kegg_pathway_counts_xlsx_filename = file.path(
  "${analysis_dir}", 
  "tables",
  "kegg_pathway_counts.xlsx"
  )

write.xlsx(kegg_pathway_counts, file=kegg_pathway_counts_xlsx_filename)
```

### Make Master Table
```{r}
piphillin_master_table = getMasterTable(
  taxa_counts=kegg_pathway_counts,
  metadata=sample_data,
  key_column='SampleName',
  taxa_colnames='PathwayName'
)
```

## Write Master Table
```{r}
wb = createWorkbook()

addWorksheet(wb, sheetName="Master Table")
openxlsx::writeData(
  wb, 
  sheet="Master Table", 
  piphillin_master_table
)

addWorksheet(wb, sheetName="Pathway Names")

openxlsx::writeData(
  wb, 
  sheet="Pathway Names", 
  pathway_name_map
)

filename = makeDataFileName(
  'master_table.xlsx',
  "${tables_dir}",
  "${analysis_type}",
  "${miseq_project_prefix}",
  "${clustering_level}"
)
print(filename)

saveWorkbook(wb, file=filename, overwrite=T)
```


### Graph
```{r}
piphillin_master_table %>% 
  select(SampleName, pathway_name_map$PathwayName) %>% 
  gather(key='PathwayName', value='abundance', -SampleName) %>% 
  ggplot(aes(x=SampleName, y=log(abundance))) + 
  geom_boxplot() 
```


```{r}
log_abundance = kegg_pathway_counts %>% mutate_at(sample_names, ~log10(.))
```


```{r}
log_abundance %>% 
  gather(key='SampleName', value='abundance', -PathwayName) %>%
  ggplot(aes(x=SampleName, y=abundance)) + 
  geom_boxplot()
```

```{r}
rzipois = function(N, p, lambda)
{
  ifelse(runif(N)<p, 0, rpois(N, lambda=lambda))
}
```

```{r}
library(car)
qqPlot(log_abundance[[20,]])
qqPlot(rzipois(1000, 0.05, 100))

```





