---
title: "${miseq_project_prefix}: ${analysis_type}"
output:
  html_notebook:
    df_print: paged
    number_sections: yes
    theme: lumen
    toc: yes
    code_folding: hide
---

<style>
  html {
    font-size: 16pt;
  }
  
  body {
    font-size: 16pt;
  }
  
  h1 {
    font-size: 2.2rem;
  }
  
  h2 {
    font-size: 2rem;
  }
  
  h3 {
    font-size: 1.8rem;
  }
  
  h4 {
    font-size: 1.4rem;
  }
  
</style>

```{r}
source('${project_metadata_file}')
analysis_type = "${analysis_type}"
clustering_level = "${clustering_level}"
filter_by = "${filter_by}"
tables_dir = "${tables_dir}"
```

# ${miseq_project_prefix} Basic Data Prep{.tabset}

**Clustering**
Data clustered at the `r toupper('${clustering_level}')` Level.

**Subgroups**
```{r, results='asis', echo=FALSE}
for (sg_filter in c(${subgroup_filters}))
{
  cat(sg_filter, '\n\n')
}
```

**Filtering**
Data filtered by `r toupper('${filter_by}')`

* **Relative Abundance cutoff:** ${relative_abundance_cutoff}
* **Prevalence Cutoff:** ${prevalence_cutoff}
* **Min Count Cutoff:** ${min_count_cutoff}
```{r}
relative_abundance_cutoff = ${relative_abundance_cutoff}
prevalence_cutoff = ${prevalence_cutoff}
min_count_cutoff = ${min_count_cutoff}
```

## Setup

### [Close]

### Start Conda ENV
```{r}
startCondaEnv('base')
```

### Load Libraries
```{r}
library(rmarkdown)
library(tidyverse)
library(magrittr)
library(ggplot2)
library(ggbeeswarm)
library(openxlsx)

library(pheatmap)
library(kableExtra)

### Custom libraries that can be loaded from GitHub
source('${utilities_dir}/analysis-utilities/general_asv_data_analysis_utilities.R')
source('${utilities_dir}/amd_templates/setup/amd_project_utilities.R')

```


### Set Knitr Options
```{r}
${knitr_options}
```

### Load Data
```{r}
asv_table = 
  read.delim(
    '${asv_table_file}', 
    header=T, 
    sep='\t',
    stringsAsFactors=F
    )

taxonomy_table = 
  read.delim(
    '${taxonomy_table_file}', 
    header=T, 
    sep='\t',
    stringsAsFactors=F
    ) %>%
  mutate_all(as.character())

sample_data = read.xlsx('${sample_data_file}')

sample_data_columns = colnames(sample_data)
sample_names = sample_data$SampleName

```

### Extract Subgroups
```{r}
if (!is.null(c(${subgroup_filters})))
{
  subgroup_filter_string = paste0(${subgroup_filters}, collapse=" & ")
  print(sprintf("Subgroup Filter String: %s", subgroup_filter_string))
  subgroup_filter_expression = str2expression(subgroup_filter_string)
  sample_data = sample_data %>% filter(eval(subgroup_filter_expression))
  sample_names = sample_data$SampleName

  asv_table = asv_table %>% select('ASVs', sample_names)
}

```


### Set Filtering
```{r}
relative_abundance_cutoff = ${relative_abundance_cutoff}
prevalence_cutoff = ${prevalence_cutoff}
```
**relative abundance cutoff:** `r relative_abundance_cutoff`
**prevalence cutoff:** `r prevalence_cutoff`

### Get Filtered Taxa Abundance
```{r}
genus_abundance = getFilteredTaxaCounts(
  asv_table,
  taxonomy_table,
  sample_data, 
  cluster_by="Genus",
  filter_by="${filter_by}",
  relative_abundance_cutoff=relative_abundance_cutoff,
  prevalence_cutoff=prevalence_cutoff,
  id_col="SampleName",
  normalize=T
  )

asv_abundance = getFilteredTaxaCounts(
  asv_table,
  taxonomy_table,
  sample_data,
  cluster_by="ASV",
  filter_by="ASV",
  relative_abundance_cutoff=0,
  prevalence_cutoff=0,
  min_count_cutoff=0,
  id_col="SampleName",
  normalize=T
)


```

## Create Master Tables
```{r}
all_master_table = getMasterTable(
  genus_abundance,
  sample_data,
  'SampleName',
  'short_glommed_taxa'
)
```


## Collect % Prevotella
```{r}
### Prevotella genera from Silva database website https://www.arb-silva.de/search/
prevotella_genera = paste0(c(
  'Bacteroidetes_Prevotella|', 
  paste0('Bacteroidetes_Prevotella_', c(1,2,6,7,9), collapse="|")
  ),
  collapse=""
)
prevotella_genera

genus_abundance$short_glommed_taxa %>% grep("Prevotella", ., value=T) %>% print()
```

Prevotella genera from Silva database website https://www.arb-silva.de/search/:
`r prevotella_genera`

**NOTE:** Switched to matchine any genus matching string "Prevotella"

```{r}
all_master_table %<>%
  mutate(
    FractionPrevotellaAll = 
      select(., matches("prevotella")) %>% 
      rowSums(),
    FractionBacteroidetes = 
      select(., starts_with('Bacteroidetes_')) %>% 
      rowSums(),
    FractionPrevotellaBacteroidetes = FractionPrevotellaAll/FractionBacteroidetes,
    PctPrevotella_gt_1 = ifelse(FractionPrevotellaAll > 0.01, 'Y', 'N'),
    PctPrevotella_gt_5 = ifelse(FractionPrevotellaAll > 0.05, 'Y', 'N'),
    PctPrevotella_gt_10 = ifelse(FractionPrevotellaAll > 0.1, 'Y', 'N')
  )
```

## Write Table
```{r}
wb = createWorkbook()

addWorksheet(wb, sheetName="Master Table")
openxlsx::writeData(wb, sheet="Master Table", all_master_table)

all_ranks = c('Phylum', 'Class', 'Order', 'Family', 'Genus')
just_taxa = genus_abundance %>% select(all_ranks, glommed_taxa, short_glommed_taxa)
addWorksheet(wb, sheetName="${tools::toTitleCase(clustering_level)}")
openxlsx::writeData(wb, sheet="${tools::toTitleCase(clustering_level)}", just_taxa)

filename = makeDataFileName(
  'master_table.xlsx',
  "${tables_dir}",
  "${analysis_type}",
  "${miseq_project_prefix}",
  "${clustering_level}"
)
print(filename)

saveWorkbook(wb, file=filename, overwrite=T)
```


