---
title: "MiSeq-${miseq_run_number}"
output:
  html_notebook:
    df_print: paged
    number_sections: yes
    theme: lumen
    toc: yes
    code_folding: hide
---

<style>
  html {
    font-size: 16pt;
  }
  
  body {
    font-size: 16pt;
  }
  
  h1 {
    font-size: 2.2rem;
  }
  
  h2 {
    font-size: 2rem;
  }
  
  h3 {
    font-size: 1.8rem;
  }
  
  h4 {
    font-size: 1.4rem;
  }
  
</style>

```{r}
source('${project_metadata_file}')
analysis_type = "${analysis_type}"
clustering_level = "${clustering_level}"
tables_dir = "${tables_dir}"
use_allbac = ${use_allbac}
```

# MiSeq-${miseq_run_number} Basic Data Prep{.tabset}
## Clustering Level
Data clustered at the `r toupper('${clustering_level}')` Level.

## Filtering
* **Relative Abundance Cutoff:** ${relative_abundance_cutoff}
* **Prevalence Cutoff:** ${prevalence_cutoff}
* **Min Count Cutoff:** ${min_count_cutoff}


## Setup 

### [Close]

### Start Conda ENV
```{r}
startCondaEnv('base')
```

### Load Libraries
```{r}
library(rmarkdown)
library(tidyverse)
library(magrittr)
library(ggplot2)
library(ggbeeswarm)
library(openxlsx)
library(DT)
library(pheatmap)
library(kableExtra)

### Custom libraries that can be loaded from GitHub
source('~/utilities/analysis-utilities/general_asv_data_analysis_utilities.R')
source('~/utilities/amd_templates/setup/amd_project_utilities.R')
```


### Set Knitr Options
```{r}
${knitr_options}
```

### Load Data
```{r}
asv_table = 
  read.delim(
    '${asv_table_file}', 
    header=T, 
    sep='\t',
    stringsAsFactors=F
    )

taxonomy_table = 
  read.delim(
    '${taxonomy_table_file}', 
    header=T, 
    sep='\t',
    stringsAsFactors=F
    ) %>%
  mutate_all(as.character())

# sample_data = 
#   read.delim(
#     '${sample_data_file}', 
#     header=T, 
#     sep='\t',
#     stringsAsFactors=F
#     ) %>%
#   filter(SampleID!='Undetermined')

sample_data = read.xlsx('${sample_data_file}')

if ("IGA" %in% colnames(sample_data) & use_allbac)
{
  sample_data = 
    sample_data %>%
    filter(IGA=='AllBac')

  sample_names = sample_data$SampleName
  
  asv_table = 
    asv_table %>%
    select('ASVs', ends_with('AllBac'))
}

sample_data_columns = colnames(sample_data)
sample_names = sample_data$SampleName

```

### Set Filtering
```{r}
relative_abundance_cutoff = ${relative_abundance_cutoff}
prevalence_cutoff = ${prevalence_cutoff}
```
**relative abundance cutoff:** `r relative_abundance_cutoff`
**prevalence cutoff:** `r prevalence_cutoff`

### Get Filtered Taxa Abundance
```{r}
genus_abundance = getFilteredTaxaCountsDev(
  asv_table,
  taxonomy_table,
  sample_data, 
  cluster_by="Genus",
  relative_abundance_cutoff=relative_abundance_cutoff,
  prevalence_cutoff=prevalence_cutoff,
  id_col="SampleName",
  normalize=T
  )
```

## Create Master Tables
```{r}
all_master_table = getMasterTable(
  genus_abundance,
  sample_data,
  'SampleName',
  'short_glommed_taxa'
)
```


## Collect % Prevotella
```{r}
### Prevotella genera from Silva database website https://www.arb-silva.de/search/
prevotella_genera = paste0(c(
  'Bacteroidetes_Prevotella|', 
  paste0('Bacteroidetes_Prevotella_', c(1,2,6,7,9), collapse="|")
  ),
  collapse=""
)
prevotella_genera
```

Prevotella genera from Silva database website https://www.arb-silva.de/search/:
`r prevotella_genera`


```{r}
all_master_table %<>%
  mutate(
    FractionPrevotellaAll = 
      select(., matches(prevotella_genera)) %>% 
      rowSums(),
    FractionBacteroidetes = 
      select(., starts_with('Bacteroidetes_')) %>% 
      rowSums(),
    FractionPrevotellaBacteroidetes = FractionPrevotellaAll/FractionBacteroidetes,
    PctPrevotella_gt_1 = ifelse(FractionPrevotellaAll > 0.01, 'Y', 'N'),
    PctPrevotella_gt_5 = ifelse(FractionPrevotellaAll > 0.05, 'Y', 'N'),
    PctPrevotella_gt_10 = ifelse(FractionPrevotellaAll > 0.1, 'Y', 'N')
  )
```

## Write Table
```{r}
wb = createWorkbook()

addWorksheet(wb, sheetName="Master Table")
openxlsx::writeData(wb, sheet="Master Table", all_master_table)

all_ranks = c('Phylum', 'Class', 'Order', 'Family', 'Genus')
just_taxa = genus_abundance %>% select(all_ranks, glommed_taxa, short_glommed_taxa)
addWorksheet(wb, sheetName="${tools::toTitleCase(clustering_level)}")
openxlsx::writeData(wb, sheet="${tools::toTitleCase(clustering_level)}", just_taxa)

filename = makeDataFileName(
  'master_table.xlsx',
  "${tables_dir}",
  "${analysis_type}",
  "${miseq_project_prefix}",
  "${clustering_level}"
)
print(filename)

saveWorkbook(wb, file=filename, overwrite=T)
```


