---
title: "MiSeq-${miseq_run_number}"
output:
  html_notebook:
    df_print: paged
    number_sections: yes
    theme: lumen
    toc: yes
    code_folding: hide
---

<style>
  html {
    font-size: 16pt;
  }
  
  body {
    font-size: 16pt;
  }
  
  h1 {
    font-size: 2.2rem;
  }
  
  h2 {
    font-size: 2rem;
  }
  
  h3 {
    font-size: 1.8rem;
    
  }
  
  h4 {
    font-size: 1.4rem;
  }
  
</style>


```{r}
source('${project_metadata_file}')
analysis_type = "${analysis_type}"
clustering_level = "${clustering_level}"
tables_dir = "${tables_dir}"
analysis_title = "${analysis_title}"
```

# MiSeq-${miseq_run_number} `r analysis_title` Two Group Plots {.tabset}
Mann-Whitney Two Group Testing

## Clustering Level
Data clustered at the `r toupper('${clustering_level}')` Level.

## Filtering
* **Relative Abundance Cutoff:** ${relative_abundance_cutoff}
* **Prevalence Cutoff:** ${prevalence_cutoff}
* **Min Count Cutoff:** ${min_count_cutoff}

## Setup

### Start Conda ENV
```{r}
startCondaEnv('base')
```

### Load Libraries
```{r}
library(rmarkdown)
# library(knitr)


library(tidyverse)
library(magrittr)
library(DT)
library(openxlsx)
library(ggplot2)
library(ggbeeswarm)

### Custom libraries that can be loaded from GitHub
source('~/utilities/analysis-utilities/general_asv_data_analysis_utilities.R')
source('~/utilities/amd_templates/setup/amd_project_utilities.R')
```


### Set Knitr Options
```{r}
${knitr_options}

```


### Load Data
```{r}
filename = makeDataFileName(
  'master_table.xlsx', 
  "${tables_dir}",
  "${analysis_type}",
  "${miseq_project_prefix}",
  "${clustering_level}"
  )
print(sprintf("master file path: %s", filename))

all_master_table = readWorkbook(filename, sheet='Master Table')
age_lt_90_master_table = all_master_table %>% filter(Age < 90)
amd_only_master_table = all_master_table %>% filter(CaseString == 'AMD')
filtered_taxa = readWorkbook(filename, sheet="${tools::toTitleCase(clustering_level)}") %>% pull("short_glommed_taxa")


filename = makeDataFileName(
  'mann_whitney_unadjusted.xlsx', 
  "${tables_dir}",
  "${analysis_type}",
  "${miseq_project_prefix}",
  "${clustering_level}"
  )
print(sprintf("mann whitney file path: %s", filename))

two_group_stats = read.xlsx(filename)

source(file.path(
  "${project_root}", 
  "${metadata_dir}", 
  'observational_variables.R'
  ))
```

### Variables of interest
#### Observational Variables
```{r}
all_variables = observational_variables[${two_group_covariates['All']}]
age_lt_90_variables = observational_variables[${two_group_covariates['Age_lt_90']}]
amd_only_variables = observational_variables[${two_group_covariates['AMD_Only']}]
```

* **All Variables:** ${two_group_covariates['All']}
* **Age < 90 Variables:** ${two_group_covariates['Age_lt_90']}
* **AMD Only Variables:** ${two_group_covariates['AMD_Only']}

#### Raw and Calculated Experimental Variables

```{r}
raw_exp_vars = ${raw_exp_vars}
calculated_exp_vars = ${calculated_exp_vars}
```
**Calculated Experimental Variables: ** `r calculated_exp_vars`


### Set Metadata Factor Levels
```{r}
all_master_table = setFactorsLevels(
  all_master_table, 
  observational_variables
  )
age_lt_90_master_table = setFactorsLevels(
  age_lt_90_master_table, 
  observational_variables
  )
amd_only_master_table = setFactorsLevels(
  amd_only_master_table, 
  observational_variables
  )
```

### P-Value Annotationas
```{r}
pval_annotations = getPvalAnnotations(
  variables=c(raw_exp_vars, calculated_exp_vars),
  master_table=all_master_table,
  pval_height_factor=1.2,
  two_group_stats=two_group_stats
)
```


## Two Group Test Plots

### All Subgroups: AMD vs. Control
```{r}
makeBoxAndDotplot(
  master_table=all_master_table,
  indices=c(raw_exp_vars, calculated_exp_vars),
  var_data=observational_variables$CaseString,
  aesthetics=list(
    x_var='CaseString', 
    color_var='CaseString', 
    facet_var='index'
    ),
  annotation_data=pval_annotations %>% filter(TestGroup=='All'),
  title=paste(analysis_title, "All Subgroups")
)
```

### Age < 90
```{r}
makeBoxAndDotplot(
  master_table=age_lt_90_master_table,
  indices=c(raw_exp_vars, calculated_exp_vars),
  var_data=observational_variables$CaseString,
  aesthetics=list(
    x_var='CaseString', 
    color_var='CaseString', 
    facet_var='index'
    ),
  annotation_data=pval_annotations %>% filter(TestGroup=='Age_lt_90'),
  title=paste(analysis_title, "Age<90")
)
```

### AMD Only---Multiple Covariates

```{r}
plotAllTwoGroupTests(
  variables=amd_only_variables,
  master_table=amd_only_master_table,
  indices=c(raw_exp_vars, calculated_exp_vars),
  title_template=paste(analysis_title, "AMD Only: "),
  annotation_data=pval_annotations %>% filter(TestGroup=='AMD_Only')
)

```


### AMD Only, Age < 90---Multiple Covariates

```{r}
plotAllTwoGroupTests(
  variables=amd_only_variables,
  master_table=amd_only_master_table,
  indices=c(raw_exp_vars, calculated_exp_vars),
  title_template=paste(analysis_title, "AMD Only: "),
  annotation_data=pval_annotations %>% filter(TestGroup=='AMD_Only_Age_lt_90')
)

```
