---
title: "MiSeq-${miseq_run_number}"
output:
  html_notebook:
    df_print: paged
    number_sections: yes
    theme: lumen
    toc: yes
    code_folding: hide
---

<style>
  html {
    font-size: 16pt;
  }
  
  body {
    font-size: 16pt;
  }
  
  h1 {
    font-size: 2.2rem;
  }
  
  h2 {
    font-size: 2rem;
  }
  
  h3 {
    font-size: 1.8rem;
  }
  
  h4 {
    font-size: 1.4rem;
  }
  
</style>


```{r}
source('${project_metadata_file}')
analysis_type = "${analysis_type}"
clustering_level = "${clustering_level}"
tables_dir = file.path(analysis_dir, analysis_type, clustering_level, 'tables')
analysis_title = "${analysis_title}"
```

# MiSeq-${miseq_run_number} `r analysis_title` Two Group Tests {.tabset}
## Clustering Level
Data clustered at the `r toupper('${clustering_level}')` Level.

## Filtering
* **Relative Abundance Cutoff:** ${relative_abundance_cutoff}
* **Prevalence Cutoff:** ${prevalence_cutoff}
* **Min Count Cutoff:** ${min_count_cutoff}


## Setup

### Start Conda ENV
```{r}
startCondaEnv('${conda_env}')
```


### Load Libraries
```{r}
library(rmarkdown)
# library(knitr)

library(tidyverse)
library(magrittr)
library(DT)
library(openxlsx)
library(ggplot2)
library(ggbeeswarm)

### Custom libraries that can be loaded from GitHub
source('~/utilities/analysis-utilities/general_asv_data_analysis_utilities.R')
source('~/utilities/analysis-utilities/alpha_div_utilities.R')
source('~/utilities/amd_templates/setup/amd_project_utilities.R')
```

### Set Knitr Options
```{r}
${knitr_options}

```

### Load Data
```{r}
filename = makeDataFileName(
  'master_table.xlsx',
  "${tables_dir}",
  "${analysis_type}",
  "${miseq_project_prefix}",
  "${clustering_level}"
)

all_master_table = readWorkbook(filename, sheet='Master Table')
age_lt_90_master_table = all_master_table %>% filter(Age < 90)
amd_only_master_table = all_master_table %>% filter(CaseString == 'AMD')
filtered_taxa = readWorkbook(filename, sheet="${tools::toTitleCase(clustering_level)}") %>% pull("short_glommed_taxa")

source(file.path(
  "${project_root}", 
  "${metadata_dir}", 
  'observational_variables.R'
  ))
```


### Variables of interest
#### All Observational Variables
```{r}
# print(observational_variables)
```

#### AMD Only Observational Variables
```{r}
amd_only_variables = observational_variables

amd_only_excluded_variables = c(
  'Gender',
  'CaseString',
  'Age'
)

for (varname in amd_only_excluded_variables)
{
  print(sprintf('Excluding %s', varname))
  amd_only_variables[varname] = NULL
}

```

**Excluding: ** `r amd_only_excluded_variables`


#### Raw and Calculated Experimental Variables

```{r}
raw_exp_vars = ${raw_exp_vars}
calculated_exp_vars = ${calculated_exp_vars}
```
**Calculated Experimental Variables: ** `r calculated_exp_vars`


### Set Metadata Factor Levels
```{r}
all_master_table = setFactorsLevels(
  all_master_table, 
  observational_variables
  )
age_lt_90_master_table = setFactorsLevels(
  age_lt_90_master_table, 
  observational_variables
  )
amd_only_master_table = setFactorsLevels(
  amd_only_master_table, 
  observational_variables
  )
```


## TwoGroup Tests

#### Initialize Table

```{r}
contrast_names = 
  makeContrastNames(observational_variables, all_master_table) %>% 
  unlist() %>%
  unname()
```

```{r}
all_two_group_stats = data.frame(
  TestGroup=character(),
  TestVariable=character(),
  Reference=character(),
  Comparison=character()
  )

for (var in c(raw_exp_vars, calculated_exp_vars))
{
  all_two_group_stats = 
    all_two_group_stats %>%
    mutate(!!as.name(var) := numeric())
}


```

#### All
```{r}
pval_list = doTwoGroupTests(
  exp_vars = c(raw_exp_vars, calculated_exp_vars),
  test_var = observational_variables$CaseString,
  master_table = all_master_table,
  test_group = 'All'
)

all_two_group_stats %<>%
  add_row(!!!pval_list)
```

#### Age < 90
```{r}
pval_list = doTwoGroupTests(
  exp_vars = c(raw_exp_vars, calculated_exp_vars),
  test_var = observational_variables$CaseString,
  master_table = age_lt_90_master_table,
  test_group = 'Age_lt_90'
)

all_two_group_stats %<>%
  add_row(!!!pval_list)

```


#### AMD Only + Covariates
```{r}
amd_only_two_group_stats = data.frame(
  TestGroup=character(),
  TestVariable=character(),
  Reference=character(),
  Comparison=character()
  )

for (var in c(raw_exp_vars, calculated_exp_vars))
{
  amd_only_two_group_stats = 
    amd_only_two_group_stats %>%
    mutate(!!as.name(var) := numeric())
}


exp_vars = c(raw_exp_vars, calculated_exp_vars)

# test_var = amd_only_variables$AREDS
master_table = amd_only_master_table
test_group = 'AMD_Only'
  
for (test_var in amd_only_variables)
{
  
  name = test_var$covariate_of_interest
  print(name)
  
  pval_list = doTwoGroupTests(
    exp_vars = exp_vars,
    test_var = test_var,
    master_table = amd_only_master_table,
    test_group = 'AMD_Only'
  )
  
  amd_only_two_group_stats %<>%
    add_row(!!!pval_list)
}
```

## Assemble Stats Table
```{r}
all_two_group_stats =
  all_two_group_stats %>%
  add_row(!!!amd_only_two_group_stats)
```

## Write Stats Table
```{r}
filename = makeDataFileName(
  'mann_whitney_unadjusted.xlsx', 
  "${tables_dir}",
  "${analysis_type}",
  "${miseq_project_prefix}",
  "${clustering_level}"
  )

print(filename)

write.xlsx(
  all_two_group_stats,
  file=filename,
  quote=F
)
```

