---
title: "${miseq_project_prefix}: ${analysis_type}"
output:
  html_notebook:
    df_print: paged
    number_sections: yes
    theme: lumen
    toc: yes
    code_folding: hide
---

<style>
  html {
    font-size: 16pt;
  }
  
  body {
    font-size: 16pt;
  }
  
  h1 {
    font-size: 2.2rem;
  }
  
  h2 {
    font-size: 2rem;
  }
  
  h3 {
    font-size: 1.8rem;
  }
  
  h4 {
    font-size: 1.4rem;
  }
  
</style>


```{r}
source('${project_metadata_file}')
analysis_type = "${analysis_type}"
clustering_level = "${clustering_level}"
analysis_title = "${analysis_title}"
```

# ${miseq_project_prefix} `r analysis_title` Univariate Tests {.tabset}
**Clustering**
Data clustered at the `r toupper('${clustering_level}')` Level.

**Subgroups**
```{r, results='asis', echo=FALSE}
for (sg_filter in c(${subgroup_filters}))
{
  cat(sg_filter, '\n\n')
}
```

**Filtering**
* **Relative Abundance cutoff:** ${relative_abundance_cutoff}
* **Prevalence Cutoff:** ${prevalence_cutoff}
* **Min Count Cutoff:** ${min_count_cutoff}
```{r}
relative_abundance_cutoff = ${relative_abundance_cutoff}
prevalence_cutoff = ${prevalence_cutoff}
min_count_cutoff = ${min_count_cutoff}
```

## Setup

### Start Conda ENV
```{r}
startCondaEnv('base')
```


### Load Libraries
```{r}
library(rmarkdown)
# library(knitr)

library(tidyverse)
library(magrittr)

library(openxlsx)
library(ggplot2)
library(ggbeeswarm)

### Custom libraries that can be loaded from GitHub
source('${utilities_dir}/analysis-utilities/general_asv_data_analysis_utilities.R')
source('${utilities_dir}/amd_templates/setup/amd_project_utilities.R')

```

### Set Knitr Options
```{r}
${knitr_options}

```

### Load Data
```{r}
filename = makeDataFileName(
  'master_table.xlsx',
  "${tables_dir}",
  "${analysis_type}",
  "${miseq_project_prefix}",
  "${clustering_level}"
)

all_master_table = readWorkbook(filename, sheet='Master Table')
age_lt_90_master_table = all_master_table %>% filter(Age < 90)
amd_only_master_table = all_master_table %>% filter(CaseString == 'AMD')
filtered_taxa = readWorkbook(filename, sheet="${tools::toTitleCase(clustering_level)}") %>% pull("short_glommed_taxa")

source(file.path(
  "${project_root}", 
  "${metadata_dir}", 
  'observational_variables.R'
  ))
```


### Variables of interest
#### Observational Variables
```{r}
case = ${case} %>% setdiff('')
categorical_variables = ${categorical_variables} %>% setdiff('')
```

* **Categorical Variables:** ${categorical_variables}

#### Raw and Calculated Experimental Variables

```{r}
raw_exp_vars = ${raw_exp_vars} %>% setdiff("")
calculated_exp_vars = ${calculated_exp_vars} %>% setdiff("")
```
**Raw Experimental Variables: ** `r raw_exp_vars`
**Calculated Experimental Variables: ** `r calculated_exp_vars`


### Set Metadata Factor Levels
```{r}
all_master_table = setFactorsLevels(
  all_master_table, 
  observational_variables
  )
age_lt_90_master_table = setFactorsLevels(
  age_lt_90_master_table, 
  observational_variables
  )
amd_only_master_table = setFactorsLevels(
  amd_only_master_table, 
  observational_variables
  )
```


## Univariate Tests

#### Initialize Table

```{r}
contrast_names = 
  makeContrastNames(observational_variables, all_master_table) %>% 
  unlist() %>%
  unname()
```

```{r}
univariate_stats_template = data.frame(
  TestGroup=character(),
  TestVariable=character(),
  Reference=character(),
  Comparison=character()
  )

exp_vars = c(raw_exp_vars, calculated_exp_vars)

for (var in exp_vars)
{
  univariate_stats_template = 
    univariate_stats_template %>%
    mutate(!!as.name(var) := numeric())
}

univariate_stats = univariate_stats_template
```


#### All

```{r}
all_univariate_stats = univariate_stats_template

master_table = all_master_table
test_group = 'All'
  
for (var_name in categorical_variables)
{
  
  test_var = observational_variables[[var_name]]
  name = test_var$covariate_of_interest
  print(name)
  
  pval_list = doUnivariateTests(
    exp_vars = exp_vars,
    test_var = test_var,
    master_table = all_master_table,
    test_group = 'All'
  )
  
  all_univariate_stats %<>%
    add_row(!!!pval_list)
}

univariate_stats %<>% add_row(!!!all_univariate_stats)
```


#### Age < 90

```{r}
age_lt_90_univariate_stats = univariate_stats_template

# test_var = amd_only_variables$AREDS
master_table = age_lt_90_master_table
test_group = 'Age_lt_90'
  
for (var_name in categorical_variables)
{
  
  test_var = observational_variables[[var_name]]
  name = test_var$covariate_of_interest
  print(name)
  
  pval_list = doUnivariateTests(
    exp_vars = exp_vars,
    test_var = test_var,
    master_table = age_lt_90_master_table,
    test_group = 'Age_lt_90'
  )
  
  age_lt_90_univariate_stats %<>%
    add_row(!!!pval_list)
}

univariate_stats %<>% add_row(!!!age_lt_90_univariate_stats)
```


#### AMD Only 
```{r}
amd_only_univariate_stats = univariate_stats_template

# test_var = amd_only_variables$AREDS
master_table = amd_only_master_table
test_group = 'AMD_Only'
  
for (var_name in categorical_variables %>% setdiff("CaseString"))
{
  
  test_var = observational_variables[[var_name]]
  name = test_var$covariate_of_interest
  print(name)
  
  pval_list = doUnivariateTests(
    exp_vars = exp_vars,
    test_var = test_var,
    master_table = amd_only_master_table,
    test_group = 'AMD_Only'
  )
  
  amd_only_univariate_stats %<>%
    add_row(!!!pval_list)
}

univariate_stats %<>% add_row(!!!amd_only_univariate_stats)
```

#### AMD Only, Age < 90
```{r}
amd_only_age_lt_90_univariate_stats = univariate_stats_template

# test_var = amd_only_variables$AREDS
master_table = amd_only_master_table %>% filter(Age < 90)
test_group = 'AMD_Only_Age_lt_90'
  
for (var_name in categorical_variables %>% setdiff("CaseString"))
{
  
  test_var = observational_variables[[var_name]]
  name = test_var$covariate_of_interest
  print(name)
  
  pval_list = doUnivariateTests(
    exp_vars = exp_vars,
    test_var = test_var,
    master_table = master_table,
    test_group = test_group
  )
  
  amd_only_age_lt_90_univariate_stats %<>%
    add_row(!!!pval_list)
}

univariate_stats %<>% add_row(!!!amd_only_age_lt_90_univariate_stats)
```


## Write Stats Table
```{r}
filename = makeDataFileName(
  'mann_whitney_unadjusted.xlsx', 
  "${tables_dir}",
  "${analysis_type}",
  "${miseq_project_prefix}",
  "${clustering_level}"
  )

print(filename)

write.xlsx(
  univariate_stats,
  file=filename,
  quote=F
)
```

