---
title: "${miseq_project_prefix}: ${analysis_type}"
output:
  html_notebook:
    df_print: paged
    number_sections: yes
    theme: lumen
    toc: yes
    code_folding: hide
---

<style>
  html {
    font-size: 16pt;
  }
  
  body {
    font-size: 16pt;
  }
  
  h1 {
    font-size: 2.2rem;
  }
  
  h2 {
    font-size: 2rem;
  }
  
  h3 {
    font-size: 1.8rem;
  }
  
  h4 {
    font-size: 1.4rem;
  }
  
</style>

```{r}
source('${project_metadata_file}')
analysis_type = "${analysis_type}"
clustering_level = "${clustering_level}"
tables_dir = "${tables_dir}"
analysis_title = tools::toTitleCase(gsub("_", " ", analysis_type))
permanova_type = "UNWEIGHTED"
weighted=F
```

# ${miseq_project_prefix} `r permanova_type` UNIFRAC PERMANOVA {.tabset}

**Clustering**
Data clustered at the `r toupper('${clustering_level}')` Level.

**Subgroups**
```{r, results='asis', echo=FALSE}
for (sg_filter in c(${subgroup_filters}))
{
  cat(sg_filter, '\n\n')
}
```

**Filtering**
* **Relative Abundance cutoff:** ${relative_abundance_cutoff}
* **Prevalence Cutoff:** ${prevalence_cutoff}
* **Min Count Cutoff:** ${min_count_cutoff}

```{r}
relative_abundance_cutoff = ${relative_abundance_cutoff}
prevalence_cutoff = ${prevalence_cutoff}
min_count_cutoff = ${min_count_cutoff}
```

## Setup

### Start Conda ENV
```{r}
startCondaEnv('permanova', lib='~/R35')
```

### Load Libraries
```{r, echo=F}
library(tidyverse)
library(magrittr)
library(openxlsx)
library(vegan)
library(phyloseq)

### Custom libraries that can be loaded from GitHub
source('~/utilities/analysis-utilities/general_asv_data_analysis_utilities.R')
source('~/utilities/analysis-utilities/beta_diversity_utilities.R')
source('~/utilities/amd_templates/setup/amd_project_utilities.R')
```

### Set Knitr Options
```{r}
${knitr_options}
```

### Load Data
```{r}
asv_table = 
  read.delim(
    '${asv_table_file}', 
    header=T, 
    sep='\t',
    stringsAsFactors=F
    )

taxonomy_table = 
  read.delim(
    '${taxonomy_table_file}', 
    header=T, 
    sep='\t',
    stringsAsFactors=F
    ) %>%
  mutate_all(as.character())

sample_data = read.xlsx('${sample_data_file}')

sample_data %<>%
  mutate(
    CNV_Either_Eye = ifelse(CNV_Either_Eye == 1, '1', '0'),
    GA_No_CNV_Either_Eye = ifelse(GA_No_CNV_Either_Eye == 1, '1', '0')
  )

sample_data_columns = colnames(sample_data)
sample_names = sample_data$SampleName

load('${phylogenetic_tree_file}')

source(file.path(
  "${project_root}", 
  "${metadata_dir}", 
  'observational_variables.R'
  ))

```
### Extract Subgroups
```{r}
if (!is.null(c(${subgroup_filters})))
{
  subgroup_filter_string = paste0(${subgroup_filters}, collapse=" & ")
  print(sprintf("Subgroup Filter String: %s", subgroup_filter_string))
  subgroup_filter_expression = str2expression(subgroup_filter_string)
  sample_data = sample_data %>% filter(eval(subgroup_filter_expression))
  sample_names = sample_data$SampleName

  asv_table = asv_table %>% select('ASVs', sample_names)
}

```


### Make Phyloseq Object
```{r}
ps = phyloseq(
  otu_table(asv_table %>%
              remove_rownames() %>%
              column_to_rownames('ASVs') %>%
              as.matrix(),
    taxa_are_rows=T
  ),
  tax_table(taxonomy_table %>%
              remove_rownames() %>%
              column_to_rownames('ASVs') %>%
              as.matrix()
  ),
  sample_data(sample_data %>%
                remove_rownames() %>%
                column_to_rownames('SampleName')
              ),
  phy_tree(fit$tree)
)

```


### Variables of interest
#### Observational Variables
```{r}
case = ${case} %>% setdiff('')
treatments = ${treatments} %>% setdiff('')
phenotypes = ${phenotypes} %>% setdiff('')
genotypes = ${genotypes} %>% setdiff('')
key_covariates = ${key_covariates} %>% setdiff('')

main_variables = c(case, key_covariates)
other_covariates = c(treatments, phenotypes, genotypes)
all_variables = c(case, key_covariates, treatments, phenotypes, genotypes)
amd_only_variables = c(key_covariates, treatments, phenotypes, genotypes)
```

* **Case:** `r case`
* **Genotype variables:** `r genotypes`
* **Phenotype variables:** `r phenotypes`
* **Treatment variables:** `r treatments`
* **Key covariates:**  `r key_covariates`



#### Make Results Table
```{r}
all_results = data.frame(Variable=all_variables, stringsAsFactors=F)
```
## All Subgroups
### Calculate UNIFRAC
```{r}
unifrac = UniFrac(
  physeq=ps,
  weighted=weighted,
  normalized=T,
  parallel=T,
  fast=T
)
```
### Individual Variables
```{r}
pvals = list()
results_colname = 'Main_Variables_Individually'

for (var in c(main_variables))
{
  print(var)
  
  heteroskedasticity_test = betadisper(unifrac, sample_data %>% pull(var), type="median") %>% print()
  
  formula = 
    paste('unifrac', '~', var) %>%
    as.formula()
  
  print(formula)
  
  permanova = adonis(
    formula=formula, 
    data=sample_data,
    permutations=999
    )
  
  pval = permanova$aov.tab[var, 'Pr(>F)']
  pvals[[var]] = pval
  
  # all_results[var, results_colname] = pval
}

pval_df = 
  pvals %>% 
  data.frame() %>% 
  gather(key='Variable', value='pval')

print(pval_df)

all_results %<>% 
  left_join(pval_df, by='Variable') %>%
  rename(!!results_colname := pval)
```

### Main Variables
```{r}
results_colname = 'Main_Variables_Together'

formula = 
  paste('unifrac', '~', paste(main_variables, collapse=' + ')) %>%
  as.formula()

print(formula)

permanova = adonis(
  formula=formula, 
  data=sample_data,
  permutations=999
  )

results_df = 
  permanova$aov.tab %>%
  data.frame() %>%
  rownames_to_column('Variable') %>%
  mutate(pval = Pr..F.) %>%
  filter(Variable %in% main_variables) %>%
  select(Variable, pval)

results_df

all_results %<>% 
  left_join(results_df, by='Variable') %>%
  rename(!!results_colname := pval)

```

### All Variables
```{r}

results_colname = 'All_Variables_Together'

ps_no_na = 
  ps %>%
  subset_samples(
    !is.na(ARMS2rs10490924) &
    !is.na(CFHrs10737680) &
    !is.na(CFHrs1061170) &
    !is.na(SKIV2Lrs429608) &
    !is.na(GA_No_CNV_Either_Eye) &
    !is.na(CNV_Either_Eye)
    )
  
sample_data_no_na = 
  sample_data %>%
  data.frame(stringsAsFactors=F) %>%
  select(all_variables) %>%
  filter(
    !is.na(ARMS2rs10490924) &
    !is.na(CFHrs10737680) &
    !is.na(CFHrs1061170) &
    !is.na(SKIV2Lrs429608) &
    !is.na(GA_No_CNV_Either_Eye) &
    !is.na(CNV_Either_Eye)
    )


unifrac = UniFrac(
  physeq=ps_no_na,
  weighted=weighted,
  normalized=T,
  parallel=T,
  fast=T
)

u =
  unifrac %>%
  as.matrix() %>%
  data.frame()

formula = 
  paste('unifrac', '~', paste(c(all_variables), collapse=' + ')) %>%
  as.formula()

print(formula)

permanova = adonis(
  formula=formula, 
  data=sample_data_no_na,
  permutations=999
  )

results_df = 
  permanova$aov.tab %>%
  data.frame() %>%
  rownames_to_column('Variable') %>%
  mutate(pval = Pr..F.) %>%
  filter(Variable %in% all_variables) %>%
  select(Variable, pval)

results_df

all_results %<>% 
  left_join(results_df, by='Variable') %>%
  rename(!!results_colname := pval)

```

## Age < 90

### Calculate UNIFRAC
```{r}

unifrac = UniFrac(
  physeq=ps %>% subset_samples(Age < 90),
  weighted=weighted,
  normalized=T,
  parallel=T,
  fast=T
)
```
### Individual Variables
```{r}

results_colname = 'Age_lt_90_Main_Variables_Individually'

pvals = list()

for (var in main_variables)
{
  print(var)
  
  heteroskedasticity_test = 
    sample_data %>% 
    filter(Age < 90) %>%
    pull(var) %>%
    betadisper(unifrac, .) %>% 
    print()
  
  formula = 
    paste('unifrac', '~', var) %>%
    as.formula()
  
  permanova = adonis(
    formula=formula, 
    data=sample_data %>% filter(Age < 90),
    permutations=999
    )
  
  pval = permanova$aov.tab[var, 'Pr(>F)']
  pvals[[var]] = pval
}

pval_df = 
  pvals %>% 
  data.frame() %>% 
  gather(key='Variable', value='pval')

print(pval_df)

all_results %<>% 
  left_join(pval_df, by='Variable') %>%
  rename(!!results_colname := pval)
```

### Main Variables
```{r}
results_colname = 'Age_lt_90_Main_Variables_Together'

# heteroskedasticity_test = 
#   sample_data %>% 
#   filter(CaseString=='AMD') %>%
#   select(key_covariates) %>%
#   betadisper(unifrac, .) %>% 
#   print()

formula = 
  paste('unifrac', '~', paste(main_variables, collapse=' + ')) %>%
  as.formula()

print(formula)

permanova = adonis(
  formula=formula, 
  data=sample_data %>% filter(Age < 90),
  permutations=999
  )

results_df = 
  permanova$aov.tab %>%
  data.frame() %>%
  rownames_to_column('Variable') %>%
  mutate(pval = Pr..F.) %>%
  filter(Variable %in% main_variables) %>%
  select(Variable, pval)

results_df

all_results %<>% 
  left_join(results_df, by='Variable') %>%
  rename(!!results_colname := pval)
```

### All Variables
```{r}

results_colname = 'Age_lt_90_All_Variables_Together'

ps_age_lt_90_no_na = 
  ps %>%
  subset_samples(
    Age < 90 &
    !is.na(ARMS2rs10490924) &
    !is.na(CFHrs10737680) &
    !is.na(CFHrs1061170) &
    !is.na(SKIV2Lrs429608) &
    !is.na(GA_No_CNV_Either_Eye) &
    !is.na(CNV_Either_Eye)
    )
  
sample_data_age_lt_90_no_na = 
  sample_data %>%
  data.frame(stringsAsFactors=F) %>%
  select(all_variables) %>%
  filter(
    Age < 90 &
    !is.na(ARMS2rs10490924) &
    !is.na(CFHrs10737680) &
    !is.na(CFHrs1061170) &
    !is.na(SKIV2Lrs429608) &
    !is.na(GA_No_CNV_Either_Eye) &
    !is.na(CNV_Either_Eye)
    )


unifrac = UniFrac(
  physeq=ps_age_lt_90_no_na,
  weighted=weighted,
  normalized=T,
  parallel=T,
  fast=T
)

u =
  unifrac %>%
  as.matrix() %>%
  data.frame()

formula = 
  paste('unifrac', '~', paste(c(all_variables), collapse=' + ')) %>%
  as.formula()

print(formula)

permanova = adonis(
  formula=formula, 
  data=sample_data_age_lt_90_no_na,
  permutations=999
  )

results_df = 
  permanova$aov.tab %>%
  data.frame() %>%
  rownames_to_column('Variable') %>%
  mutate(pval = Pr..F.) %>%
  filter(Variable %in% all_variables) %>%
  select(Variable, pval)

results_df

all_results %<>% 
  left_join(results_df, by='Variable') %>%
  rename(!!results_colname := pval)

```

## AMD Only

### Individual Variables
```{r}
pvals = list()

results_colname = 'AMD_Only_Main_Variables_Individually'

amd_only_ps = subset_samples(ps, CaseString=='AMD')
amd_only_sample_data = 
  sample_data %>%
  filter(CaseString=='AMD')

for (var in amd_only_variables %>% setdiff('CaseString'))
{
  print(var)
  
  var_data = observational_variables[[var]]
  name = var_data$covariate_of_interest
  case = var_data$case
  control = var_data$control
  ref = var_data$labels$reference
  comp = var_data$labels$comparison
  
  if (var == 'Age')
  {
    ### Make all false
    remove_idx = amd_only_sample_data$Age > 0
  } else
  {
    remove_idx = 
      amd_only_sample_data %>%
      mutate(
        remove_idx = (!!as.name(var) %in% c(var_data$case, var_data$control))
        ) %>%
      pull(remove_idx)
  }

  
  this_ps = prune_samples(remove_idx, amd_only_ps)
  this_sample_data = amd_only_sample_data %>% filter(remove_idx)
  
  unifrac = UniFrac(
    physeq=this_ps,
    weighted=weighted,
    normalized=T,
    parallel=T,
    fast=T
  )
  
  # heteroskedasticity_test = 
  #   amd_only_sample_data %>% 
  #   pull(var) %>%
  #   betadisper(unifrac, .) %>% 
  #   print()
  
  formula = 
    paste('unifrac', '~', var) %>%
    as.formula()
  
  print(formula)
  
  permanova = adonis(
    formula=formula, 
    data=this_sample_data,
    permutations=999
    )
  
  pval = permanova$aov.tab[var, 'Pr(>F)']
  pvals[[var]] = pval
}

pval_df = 
  pvals %>% 
  data.frame() %>% 
  gather(key='Variable', value='pval')

print(pval_df)


all_results %<>% 
  left_join(pval_df, by='Variable') %>%
  rename(!!results_colname := pval)
```

### Main Variables
```{r}
results_colname = 'AMD_Only_Main_Variables_and_Single_Other'

vars_to_use = main_variables %>% setdiff("CaseString")

formula = 
  paste(
    'unifrac', '~', 
    paste(vars_to_use, collapse=' + ')
    ) %>%
  as.formula()

print(formula)

this_ps = subset_samples(ps, CaseString=='AMD')
this_sample_data = sample_data %>% filter(CaseString=='AMD')

unifrac = UniFrac(
  physeq=this_ps,
  weighted=weighted,
  normalized=T,
  parallel=T,
  fast=T
)

permanova = adonis(
  formula=formula, 
  data=this_sample_data,
  permutations=999
  )

results_df = 
  permanova$aov.tab %>%
  data.frame() %>%
  rownames_to_column('Variable') %>%
  mutate(pval = Pr..F.) %>%
  filter(Variable %in% vars_to_use) %>%
  select(Variable, pval)

all_results %<>% 
  left_join(results_df, by='Variable') %>%
  rename(!!results_colname := pval)

for (var in other_covariates)
{
  
  print(var)
  
  var_data = observational_variables[[var]]
  name = var_data$covariate_of_interest
  case = var_data$case
  control = var_data$control
  ref = var_data$labels$reference
  comp = var_data$labels$comparison
  
  if (var == 'Age')
  {
    ### Make all false
    remove_idx = amd_only_sample_data$Age > 0
  } else
  {
    remove_idx = 
      amd_only_sample_data %>%
      mutate(
        remove_idx = (!!as.name(var) %in% c(var_data$case, var_data$control))
        ) %>%
      pull(remove_idx)
  }

  
  this_ps = prune_samples(remove_idx, amd_only_ps)
  this_sample_data = amd_only_sample_data %>% filter(remove_idx)
  
  unifrac = UniFrac(
    physeq=this_ps,
    weighted=weighted,
    normalized=T,
    parallel=T,
    fast=T
  )
  
  vars_to_use = c(main_variables %>% setdiff('CaseString'), var)
  
  results_colname = paste0('AMD_Only_Main_Plus_', var)
  
  formula = 
    paste(
      'unifrac', '~', 
      paste(vars_to_use, collapse=' + ')
      ) %>%
    as.formula()

  print(formula)
  
  # unifrac = as.matrix(unifrac)
  
  filtered_sample_data = 
  
  permanova = adonis(
    formula=formula, 
    data=this_sample_data,
    permutations=999
    )
  
  results_df = 
    permanova$aov.tab %>%
    data.frame() %>%
    rownames_to_column('Variable') %>%
    mutate(pval = Pr..F.) %>%
    filter(Variable %in% c(main_variables %>% setdiff('CaseString'), var)) %>%
    select(Variable, pval)
  
  results_df
  
  all_results %<>% 
    left_join(results_df, by='Variable') %>%
    rename(!!results_colname := pval)
}

```

### All Variables
```{r}

results_colname = 'AMD_Only_All_Variables_Together'

ps_amd_only_no_na = 
  ps %>%
  subset_samples(
    CaseString == 'AMD' &
    !is.na(ARMS2rs10490924) &
    !is.na(CFHrs10737680) &
    !is.na(CFHrs1061170) &
    !is.na(SKIV2Lrs429608) &
    !is.na(GA_No_CNV_Either_Eye) &
    !is.na(CNV_Either_Eye)
    )
  
sample_data_amd_only_no_na = 
  sample_data %>%
  data.frame(stringsAsFactors=F) %>%
  select(all_variables) %>%
  filter(
    CaseString == 'AMD' &
    !is.na(ARMS2rs10490924) &
    !is.na(CFHrs10737680) &
    !is.na(CFHrs1061170) &
    !is.na(SKIV2Lrs429608) &
    !is.na(GA_No_CNV_Either_Eye) &
    !is.na(CNV_Either_Eye)
    )


unifrac = UniFrac(
  physeq=ps_amd_only_no_na,
  weighted=weighted,
  normalized=T,
  parallel=T,
  fast=T
)

u =
  unifrac %>%
  as.matrix() %>%
  data.frame()

formula = 
  paste('unifrac', '~', paste(c(amd_only_variables), collapse=' + ')) %>%
  as.formula()

print(formula)

permanova = adonis(
  formula=formula, 
  data=sample_data_amd_only_no_na,
  permutations=999
  )

results_df = 
  permanova$aov.tab %>%
  data.frame() %>%
  rownames_to_column('Variable') %>%
  mutate(pval = Pr..F.) %>%
  filter(Variable %in% amd_only_variables) %>%
  select(Variable, pval)

results_df

all_results %<>% 
  left_join(results_df, by='Variable') %>%
  rename(!!results_colname := pval)

```


## Write Results
```{r}
filename = makeDataFileName(
  'unifrac_permanova.xlsx',
  "${tables_dir}",
  "${analysis_type}",
  "${miseq_project_prefix}",
  "${clustering_level}"
)

print(filename)

if (file.exists(filename))
{
  print(sprintf('workbook %s exists', filename))
  wb = loadWorkbook(filename)
} else
{
  print(sprintf('creating workbook %s', filename))
  wb = createWorkbook()
}

sheet_name = sprintf("%s PERMANOVA", permanova_type)

if (sheet_name %in% sheets(wb))
{
  print(sprintf('removing worksheet %s', sheet_name))
  removeWorksheet(wb, sheet=sheet_name)
}

print(sprintf('creating worksheet %s', sheet_name))
addWorksheet(wb, sheetName=sheet_name)

openxlsx::writeData(wb, sheet=sheet_name, x=all_results)

saveWorkbook(wb, file=filename, overwrite=T)

```



