---
title: "${miseq_project_prefix}: ${analysis_type}"
output:
  html_notebook:
    df_print: paged
    number_sections: yes
    theme: lumen
    toc: yes
    code_folding: hide
---

<style>
  html {
    font-size: 16pt;
  }
  
  body {
    font-size: 16pt;
  }
  
  h1 {
    font-size: 2.2rem;
  }
  
  h2 {
    font-size: 2rem;
  }
  
  h3 {
    font-size: 1.8rem;
  }
  
  h4 {
    font-size: 1.4rem;
  }
  
  .dataTable {
    font-size: 9pt;
  }
  
</style>

```{r}
source('${project_metadata_file}')
analysis_type = "${analysis_type}"
clustering_level = "${clustering_level}"
tables_dir = "${tables_dir}"
analysis_title = tools::toTitleCase(gsub("_", " ", analysis_type))
cooks_cutoff = ${cooks_cutoff}
independent_filtering = ${independent_filtering}
use_age = ${use_age}
```

# ${miseq_project_prefix} DESeq2 Analysis {.tabset}

**Clustering**
Data clustered at the `r toupper('${clustering_level}')` Level.

**Subgroups**
```{r, results='asis', echo=FALSE}
for (sg_filter in c(${subgroup_filters}))
{
  cat(sg_filter, '\n\n')
}
```

**Filtering**
* **Relative Abundance cutoff:** ${relative_abundance_cutoff}
* **Prevalence Cutoff:** ${prevalence_cutoff}
* **Min Count Cutoff:** ${min_count_cutoff}

```{r}
relative_abundance_cutoff = ${relative_abundance_cutoff}
prevalence_cutoff = ${prevalence_cutoff}
min_count_cutoff = ${min_count_cutoff}
```

## Setup

### [Close]

### Start Conda ENV
```{r}
startCondaEnv('deseq', lib='~/R35')
```

### Load Libraries
```{r}
library(rmarkdown)
library(tidyverse)
library(magrittr)
library(ggplot2)
library(openxlsx)
library(DT)
library(kableExtra)
library(DESeq2)

### Custom libraries that can be loaded from GitHub
source('~/utilities/analysis-utilities/general_asv_data_analysis_utilities.R')
source('~/utilities/analysis-utilities/deseq_utilities.R')
source('~/utilities/amd_templates/setup/amd_project_utilities.R')
```

### Set Knitr Options
```{r}
${knitr_options}
```

### Load Data
```{r}
asv_table = 
  read.delim(
    '${asv_table_file}', 
    header=T, 
    sep='\t',
    stringsAsFactors=F
    )

taxonomy_table = 
  read.delim(
    '${taxonomy_table_file}', 
    header=T, 
    sep='\t',
    stringsAsFactors=F
    ) %>%
  mutate_all(as.character())

sample_data = read.xlsx('${sample_data_file}')

sample_data_columns = colnames(sample_data)
sample_names = sample_data$SampleName

source(file.path(
  "${project_root}", 
  "${metadata_dir}", 
  'observational_variables.R'
  ))

```

### Extract Subgroups
```{r}
if (!is.null(c(${subgroup_filters})))
{
  subgroup_filter_string = paste0(${subgroup_filters}, collapse=" & ")
  print(sprintf("Subgroup Filter String: %s", subgroup_filter_string))
  subgroup_filter_expression = str2expression(subgroup_filter_string)
  sample_data = sample_data %>% filter(eval(subgroup_filter_expression))
  sample_names = sample_data$SampleName

  asv_table = asv_table %>% select('ASVs', sample_names)
}

```



### Set Metadata Factor Levels
```{r}
sample_data = setFactorsLevels(sample_data, observational_variables)
# sample_data %<>%
#   mutate(AREDS=
#     ifelse(is.na(AREDS), "MISSING", AREDS) %>%
#     factor(levels=c('N', 'Y', 'MISSING'))
#   )
```

### Variables of interest
#### Observational Variables
```{r}
all_covariates = observational_variables[${linear_regression_predictors['All']}]
age_lt_90_covariates = observational_variables[${linear_regression_predictors['Age_lt_90']}]
genotypes = ${genotypes}
phenotypes = ${phenotypes}
```

#### Remove Age
```{r}
if (!use_age)
{
  all_covariates$Age = NULL
  age_lt_90_covariates$Age = NULL 
}

```

* **All Variables:** `r all_covariates %>% names()`
* **Age < 90 Variables:** `r age_lt_90_covariates %>% names()`


### Extract Subgroups
```{r}
subgroup_filter_string = paste0(${subgroup_filters}, collapse=" & ")
print(subgroup_filter_string)
subgroup_filter_expression = str2expression(subgroup_filter_string)
```
`r subgroup_filter_string`

```{r}
sample_data = sample_data %>% filter(eval(subgroup_filter_expression))
sample_names = sample_data$SampleName

asv_table = asv_table %>% select('ASVs', sample_names)
```

### Set Filtering
```{r}
relative_abundance_cutoff=${relative_abundance_cutoff}
prevalence_cutoff=${prevalence_cutoff}
```

### Set Ranks
```{r}
all_ranks = c('Phylum', 'Class', 'Order', 'Family', 'Genus')
```

## All Subgroups Case/Control Only
### Create Workbook
```{r}
deseq_results_wb = createWorkbook()
```

#### Run and add worksheets
```{r}
for (lowest_rank in all_ranks)
{
  
  taxa_counts = getFilteredTaxaCountsDev(
    asv_table=asv_table,
    taxonomy_table=taxonomy_table,
    sample_data=sample_data,
    cluster_by=lowest_rank,
    relative_abundance_cutoff=relative_abundance_cutoff,
    prevalence_cutoff=prevalence_cutoff,
    min_count_cutoff=0,
    filter_by='Taxa',
    clean_taxa=T,  ### remove NAs in lowest rank
    n_max_by_mean=F,
    id_col="SampleName", ### metadata column that containes the unique sample IDs
    add_glommed_names=T,
    normalize=F
  )
    
  results_df = runDeseqFromTables_dev(
    asv_table,
    taxonomy_table,
    sample_data,
    "CaseString",
    lowest_rank=lowest_rank,
    relative_abundance_cutoff=relative_abundance_cutoff,
    prevalence_cutoff=prevalence_cutoff,
    include_covariates = 'All',
    cooks_cutoff=cooks_cutoff,
    independent_filtering=independent_filtering
    )
  
  padj_col = "CaseString_AMD_vs_Control.padj"
  
  results_df =
    results_df %>%
    select(Taxon, !!padj_col, everything()) %>%
    arrange(!!as.name(padj_col))
  
  addWorksheet(
    wb=deseq_results_wb, 
    sheetName=lowest_rank
    )
  
  openxlsx::writeData(
    wb=deseq_results_wb, 
    sheet=lowest_rank, 
    x=results_df
    )
    
  setColWidths(
    wb=deseq_results_wb, 
    sheet=lowest_rank, 
    cols=1, 
    widths = "auto"
  )
  
  freezePane(
    deseq_results_wb, 
    sheet=lowest_rank, 
    firstCol=T,
    firstRow=T
  )
}
```

#### Save Workbook
```{r}
filename = makeDataFileName(
  'all_subgroups_case_vs_control.xlsx',
  "${tables_dir}",
  "${analysis_type}",
  "${miseq_project_prefix}",
  "rank"
)

print(filename)

saveWorkbook(deseq_results_wb, file=filename, overwrite=T)
```

## All Subgroups -- All Variables
### Create Workbook
```{r}
deseq_results_wb = createWorkbook()
```

### Run and add worksheets
```{r}
for (lowest_rank in all_ranks)
{
  
  taxa_counts = getFilteredTaxaCountsDev(
    asv_table=asv_table,
    taxonomy_table=taxonomy_table,
    sample_data=sample_data,
    cluster_by=lowest_rank,
    relative_abundance_cutoff=relative_abundance_cutoff,
    prevalence_cutoff=prevalence_cutoff,
    min_count_cutoff=0,
    filter_by='Taxa',
    clean_taxa=T,  ### remove NAs in lowest rank
    n_max_by_mean=F,
    id_col="SampleName", ### metadata column that containes the unique sample IDs
    add_glommed_names=T,
    normalize=F
  )
    
  results_df = runDeseqFromTables_dev(
    asv_table,
    taxonomy_table,
    sample_data,
    all_covariates %>% names(),
    lowest_rank=lowest_rank,
    relative_abundance_cutoff=relative_abundance_cutoff,
    prevalence_cutoff=prevalence_cutoff,
    include_covariates = 'All',
    cooks_cutoff=cooks_cutoff,
    independent_filtering=independent_filtering
    )
  
  padj_col = "CaseString_AMD_vs_Control.padj"
  
  results_df =
    results_df %>%
    select(Taxon, !!padj_col, starts_with("CaseString"), everything()) %>%
    arrange(!!as.name(padj_col))
  
  addWorksheet(
    wb=deseq_results_wb, 
    sheetName=lowest_rank
    )
  
  openxlsx::writeData(
    wb=deseq_results_wb, 
    sheet=lowest_rank, 
    x=results_df
    )
    
  setColWidths(
    wb=deseq_results_wb, 
    sheet=lowest_rank, 
    cols=1, 
    widths = "auto"
  )
  
  freezePane(
    deseq_results_wb, 
    sheet=lowest_rank, 
    firstCol=T,
    firstRow=T
  )
}
```

### Save Workbook
```{r}
filename = makeDataFileName(
  'all_subgroups_with_covariates.xlsx',
  "${tables_dir}",
  "${analysis_type}",
  "${miseq_project_prefix}",
  "rank"
)

print(filename)

saveWorkbook(deseq_results_wb, file=filename, overwrite=T)
```

## Age < 90 Case/Control Only

### Create Workbook
```{r}
deseq_results_wb = createWorkbook()
```

### Run and add worksheets
```{r}
for (lowest_rank in all_ranks)
{
  results_df = runDeseqFromTables_dev(
    asv_table,
    taxonomy_table,
    sample_data %>% filter(Age < 90),
    "CaseString",
    lowest_rank=lowest_rank,
    relative_abundance_cutoff=relative_abundance_cutoff,
    prevalence_cutoff=prevalence_cutoff,
    include_covariates = 'All',
    cooks_cutoff=cooks_cutoff,
    independent_filtering=independent_filtering
    )
  
  padj_col = "CaseString_AMD_vs_Control.padj"
  
  results_df =
    results_df %>%
    select(Taxon, !!padj_col, everything()) %>%
    arrange(!!as.name(padj_col))
  
  addWorksheet(
    wb=deseq_results_wb, 
    sheetName=lowest_rank
    )
  
  openxlsx::writeData(
    wb=deseq_results_wb, 
    sheet=lowest_rank, 
    x=results_df
    )
    
  setColWidths(
    wb=deseq_results_wb, 
    sheet=lowest_rank, 
    cols=1, 
    widths = "auto"
  )
  
  freezePane(
    deseq_results_wb, 
    sheet=lowest_rank, 
    firstCol=T,
    firstRow=T
  )
}
```

### Save Workbook
```{r}
filename = makeDataFileName(
  'age_lt_90_case_vs_control.xlsx',
  "${tables_dir}",
  "${analysis_type}",
  "${miseq_project_prefix}",
  "rank"
)

print(filename)

saveWorkbook(deseq_results_wb, file=filename, overwrite=T)
```


## Age < 90

### Create Workbook
```{r}
deseq_results_wb = createWorkbook()
```

### Run and add worksheets
```{r}
for (lowest_rank in all_ranks)
{
  results_df = runDeseqFromTables_dev(
    asv_table,
    taxonomy_table,
    sample_data %>% filter(Age < 90),
    age_lt_90_covariates %>% names(),
    lowest_rank=lowest_rank,
    relative_abundance_cutoff=relative_abundance_cutoff,
    prevalence_cutoff=prevalence_cutoff,
    include_covariates = 'All',
    cooks_cutoff=cooks_cutoff,
    independent_filtering=independent_filtering
    )
  
  padj_col = "CaseString_AMD_vs_Control.padj"
  
  results_df =
    results_df %>%
    select(Taxon, !!padj_col, starts_with("CaseString"), everything()) %>%
    arrange(!!as.name(padj_col))
  
  addWorksheet(
    wb=deseq_results_wb, 
    sheetName=lowest_rank
    )
  
  openxlsx::writeData(
    wb=deseq_results_wb, 
    sheet=lowest_rank, 
    x=results_df
    )
    
  setColWidths(
    wb=deseq_results_wb, 
    sheet=lowest_rank, 
    cols=1, 
    widths = "auto"
  )
  
  freezePane(
    deseq_results_wb, 
    sheet=lowest_rank, 
    firstCol=T,
    firstRow=T
  )
}
```

### Save Workbook
```{r}
filename = makeDataFileName(
  'age_lt_90_with_covariates.xlsx',
  "${tables_dir}",
  "${analysis_type}",
  "${miseq_project_prefix}",
  "rank"
)

print(filename)

saveWorkbook(deseq_results_wb, file=filename, overwrite=T)
```
