---
title: "${miseq_project_prefix}: ${analysis_type}"
output:
  html_notebook:
    df_print: paged
    number_sections: yes
    theme: lumen
    toc: yes
    code_folding: hide
---

<style>
  html {
    font-size: 16pt;
  }
  
  body {
    font-size: 16pt;
  }
  
  h1 {
    font-size: 2.2rem;
  }
  
  h2 {
    font-size: 2rem;
  }
  
  h3 {
    font-size: 1.8rem;
  }
  
  h4 {
    font-size: 1.4rem;
  }
  
</style>

```{r}
source('${project_metadata_file}')
analysis_type = "${analysis_type}"
clustering_level = "${clustering_level}"
tables_dir = "${tables_dir}"
```

# ${miseq_project_prefix} Percent Prevotella Data Prep {.tabset}

**Clustering**
Data clustered at the `r toupper('${clustering_level}')` Level.

**Subgroups**
```{r, results='asis', echo=FALSE}
for (sg_filter in c(${subgroup_filters}))
{
  cat(sg_filter, '\n\n')
}
```

**Filtering**
* **Relative Abundance cutoff:** ${relative_abundance_cutoff}
* **Prevalence Cutoff:** ${prevalence_cutoff}
* **Min Count Cutoff:** ${min_count_cutoff}
```{r}
relative_abundance_cutoff = ${relative_abundance_cutoff}
prevalence_cutoff = ${prevalence_cutoff}
min_count_cutoff = ${min_count_cutoff}
```

## Setup

### [Close]

### Start Conda ENV
```{r}
startCondaEnv('regression')
```

### Load Libraries
```{r}
library(rmarkdown)
library(tidyverse)
library(magrittr)
library(ggplot2)
library(ggbeeswarm)
library(openxlsx)
library(DT)
library(pheatmap)
library(kableExtra)

### Custom libraries that can be loaded from GitHub
source('~/utilities/analysis-utilities/general_asv_data_analysis_utilities.R')
source('~/utilities/amd_templates/setup/amd_project_utilities.R')
```


### Set Knitr Options
```{r}
${knitr_options}
```

### Load Data
```{r}
asv_table = 
  read.delim(
    '${asv_table_file}', 
    header=T, 
    sep='\t',
    stringsAsFactors=F
    )

taxonomy_table = 
  read.delim(
    '${taxonomy_table_file}', 
    header=T, 
    sep='\t',
    stringsAsFactors=F
    ) %>%
  mutate_all(as.character())

sample_data = read.xlsx('${sample_data_file}')

sample_data_columns = colnames(sample_data)
sample_names = sample_data$SampleName

```

### Extract Subgroups
```{r}
subgroup_filter_string = paste0(${subgroup_filters}, collapse=" & ")
print(subgroup_filter_string)
subgroup_filter_expression = str2expression(subgroup_filter_string)
```
`r subgroup_filter_string`

```{r}
sample_data = sample_data %>% filter(eval(subgroup_filter_expression))
sample_names = sample_data$SampleName

asv_table = asv_table %>% select('ASVs', sample_names)
```


### Set Filtering
```{r}
relative_abundance_cutoff = ${relative_abundance_cutoff}
prevalence_cutoff = ${prevalence_cutoff}
```
**relative abundance cutoff:** `r relative_abundance_cutoff`
**prevalence cutoff:** `r prevalence_cutoff`

### Get Filtered Taxa Abundance
```{r}
genus_abundance = getFilteredTaxaCounts(
  asv_table,
  taxonomy_table,
  sample_data, 
  cluster_by='Genus',
  relative_abundance_cutoff=relative_abundance_cutoff,
  prevalence_cutoff=prevalence_cutoff,
  id_col="SampleName",
  normalize=T
  )

fb_taxa_abundance = 
  genus_taxa_abundance %>%
  select(Phylum, !!sample_names) %>%
  filter(Phylum %in% c('Bacteroidetes', 'Firmicutes')) %>%
  group_by(Phylum) %>%
  summarise_all(sum)

phyla_abundance = getFilteredTaxaCounts(
  asv_table,
  taxonomy_table,
  sample_data,
  cluster_by='Phylum',
  relative_abundance_cutoff=relative_abundance_cutoff,
  prevalence_cutoff=prevalence_cutoff,
  id_col="SampleName",
  normalize=T
  )
```

# Firmicute and Bacteroidetes Table

## Create FB Stats Table

Final table has columns for

* Firmicutes abundance
* Bacteroidets abundance
* F to B ratio
  $$ f2br = \frac{[Firmicutes]}{[Bacteroidetes]} $$
* Log F to B ratio $logf2br$
* FB index
  $$ fbidx = \frac{[Firmicutes] - [Bacteroidetes]}{[Firmicutes] + [Bacteroidetes]}$$
* All metadata columns (`SampleName`, `CaseString`, `AREDS`, `Gender`, etc.)

Each row represents a different sample.

The three measures are different ways of characterize the relationship between Firmicute and Bacteroidetes abundance. The $fbidx$ varies from -1 (only Firmicutes present) to 1 (only Bacteroidetes present) for that sample.

```{r}
genus_master_table = getMasterTable(
  genus_taxa_abundance,
  sample_data,
  'SampleName',
  'short_glommed_taxa'
)

fb_master_table = 
  fb_taxa_abundance %>% 
  column_to_rownames('Phylum') %>% 
  t() %>% 
  data.frame(stringsAsFactors=F) %>% 
  rownames_to_column('SampleName') %>%
  inner_join(genus_master_table, by='SampleName') %>%
  mutate(
    F2B_Ratio = Firmicutes/Bacteroidetes,
    LogF2B_Ratio = log(F2B_Ratio),
    FB_Index = (Firmicutes - Bacteroidetes)/(Firmicutes + Bacteroidetes),
    FxB = Firmicutes*Bacteroidetes
  ) %>%
  select(
    SampleName, 
    Firmicutes, 
    Bacteroidetes, 
    F2B_Ratio, 
    LogF2B_Ratio, 
    FB_Index, 
    sample_data_columns,
    everything()
    )

phylum_master_table = getMasterTable(
  phyla_abundance,
  sample_data,
  key_column='SampleName',
  taxa_colnames='short_glommed_taxa'
  ) %>%
  mutate(
    F2B_Ratio = Firmicutes/Bacteroidetes,
    LogF2B_Ratio = log(F2B_Ratio),
    FB_Index = (Firmicutes - Bacteroidetes)/(Firmicutes + Bacteroidetes),
    FxB = Firmicutes*Bacteroidetes
  ) %>%
  select(
    SampleName, 
    Firmicutes, 
    Bacteroidetes, 
    F2B_Ratio, 
    LogF2B_Ratio, 
    FB_Index, 
    sample_data_columns,
    everything()
    )

all_master_table = phylum_master_table
age_lt_90_master_table = all_master_table %>% filter(Age < 90)
amd_only_master_table = all_master_table %>% filter(CaseString == 'AMD')

```

## Write Table
```{r}
FB_indices = createWorkbook()

# addWorksheet(FB_indices, sheetName="All Subgroups")
# openxlsx::writeData(FB_indices, sheet="All Subgroups", all_master_table)
# addWorksheet(FB_indices, sheetName="Age < 90")
# openxlsx::writeData(FB_indices, sheet="Age < 90", age_lt_90_master_table)
# addWorksheet(FB_indices, sheetName="AMD Only")
# openxlsx::writeData(FB_indices, sheet="AMD Only", amd_only_master_table)
# addWorksheet(FB_indices, sheetName="Taxa")
# openxlsx::writeData(FB_indices, sheet="Taxa", phyla_abundance$short_glommed_taxa)

addWorksheet(FB_indices, sheetName="Master Table")
openxlsx::writeData(FB_indices, sheet="Master Table", fb_master_table)

just_taxa = phyla_abundance %>% select(Phylum, short_glommed_taxa)
addWorksheet(FB_indices, sheetName="${tools::toTitleCase(clustering_level)}")
openxlsx::writeData(FB_indices, sheet="${tools::toTitleCase(clustering_level)}", just_taxa)

filename = makeDataFileName(
  'master_table.xlsx',
  "${tables_dir}",
  "${analysis_type}",
  "${miseq_project_prefix}",
  "${clustering_level}"
)
print(filename)

saveWorkbook(FB_indices, file=filename, overwrite=T)
```


