---
title: "MiSeq-${miseq_run_number}"
output:
  html_notebook:
    df_print: paged
    number_sections: yes
    theme: lumen
    toc: yes
    code_folding: hide
---

<style>
  html {
    font-size: 16pt;
  }
  
  body {
    font-size: 16pt;
  }
  
  h1 {
    font-size: 2.2rem;
  }
  
  h2 {
    font-size: 2rem;
  }
  
  h3 {
    font-size: 1.8rem;
  }
  
  h4 {
    font-size: 1.4rem;
  }
  
</style>

```{r}
source('${project_metadata_file}')
analysis_type = "${analysis_type}"
clustering_level = "${clustering_level}"
tables_dir = "${tables_dir}"
```

# MiSeq-${miseq_run_number} FB Analysis Data Prep{.tabset}

**Clustering**
Data clustered at the `r toupper('${clustering_level}')` Level.

**Subgroups**
```{r, results='asis', echo=FALSE}
for (sg_filter in c(${subgroup_filters}))
{
  cat(sg_filter, '\n\n')
}
```

**Filtering**
  * **Relative Abundance cutoff:** ${relative_abundance_cutoff}
* **Prevalence Cutoff:** ${prevalence_cutoff}
* **Min Count Cutoff:** ${min_count_cutoff}
```{r}
relative_abundance_cutoff = ${relative_abundance_cutoff}
prevalence_cutoff = ${prevalence_cutoff}
min_count_cutoff = ${min_count_cutoff}
```

## Setup

### [Close]

### Start Conda ENV
```{r}
startCondaEnv('regression')
```

### Load Libraries
```{r}
library(rmarkdown)
library(tidyverse)
library(magrittr)
library(ggplot2)
library(ggbeeswarm)
library(openxlsx)
library(DT)
library(pheatmap)
library(kableExtra)

### Custom libraries that can be loaded from GitHub
source('~/utilities/analysis-utilities/general_asv_data_analysis_utilities.R')
source('~/utilities/amd_templates/setup/amd_project_utilities.R')
```


### Set Knitr Options
```{r}
${knitr_options}
```

### Load Data
```{r}
sample_data = 
  read.delim(
    '${sample_data_file}', 
    header=T, 
    sep='\t',
    stringsAsFactors=F
    ) %>%
  filter(SampleID!='Undetermined')


if ("IGA" %in% colnames(sample_data) & use_allbac)
{
  sample_data = 
    sample_data %>%
    filter(IGA=='Pos')

  sample_names = sample_data$SampleName
}

sample_data_columns = colnames(sample_data)
sample_names = sample_data$SampleName


source(file.path(
  "${project_root}", 
  "${metadata_dir}", 
  'observational_variables.R'
  ))

```

### Extract Subgroups
```{r}
subgroup_filter_string = paste0(${subgroup_filters}, collapse=" & ")
print(subgroup_filter_string)
subgroup_filter_expression = str2expression(subgroup_filter_string)
```
`r subgroup_filter_string`

```{r}
sample_data = sample_data %>% filter(eval(subgroup_filter_expression))
sample_names = sample_data$SampleName

asv_table = asv_table %>% select('ASVs', sample_names)
```

## Distribution
```{r}
hist(samle_data$Pct_Iga_Pos)
```

```{r}
all_master_table = setFactorsLevels(
  all_master_table, 
  observational_variables
  )
age_lt_90_master_table = setFactorsLevels(
  age_lt_90_master_table, 
  observational_variables
  )
amd_only_master_table = setFactorsLevels(
  amd_only_master_table, 
  observational_variables
  )
```

## TwoGroup Tests

#### Initialize Table

```{r}
contrast_names = 
  makeContrastNames(observational_variables, all_master_table) %>% 
  unlist() %>%
  unname()
```

```{r}
all_two_group_stats = data.frame(
  TestGroup=character(),
  TestVariable=character(),
  Reference=character(),
  Comparison=character()
  )

for (var in c(raw_exp_vars, calculated_exp_vars))
{
  all_two_group_stats = 
    all_two_group_stats %>%
    mutate(!!as.name(var) := numeric())
}


```

#### All
```{r}
pval_list = doTwoGroupTests(
  exp_vars = c(raw_exp_vars, calculated_exp_vars),
  test_var = observational_variables$CaseString,
  master_table = all_master_table,
  test_group = 'All'
)

all_two_group_stats %<>%
  add_row(!!!pval_list)
```

#### Age < 90
```{r}
pval_list = doTwoGroupTests(
  exp_vars = c(raw_exp_vars, calculated_exp_vars),
  test_var = observational_variables$CaseString,
  master_table = age_lt_90_master_table,
  test_group = 'Age_lt_90'
)

all_two_group_stats %<>%
  add_row(!!!pval_list)

```


#### AMD Only + Covariates
```{r}
amd_only_two_group_stats = data.frame(
  TestGroup=character(),
  TestVariable=character(),
  Reference=character(),
  Comparison=character()
  )

for (var in c(raw_exp_vars, calculated_exp_vars))
{
  amd_only_two_group_stats = 
    amd_only_two_group_stats %>%
    mutate(!!as.name(var) := numeric())
}


exp_vars = c(raw_exp_vars, calculated_exp_vars)

# test_var = amd_only_variables$AREDS
master_table = amd_only_master_table
test_group = 'AMD_Only'
  
for (test_var in amd_only_variables)
{
  
  name = test_var$covariate_of_interest
  print(name)
  
  pval_list = doTwoGroupTests(
    exp_vars = exp_vars,
    test_var = test_var,
    master_table = amd_only_master_table,
    test_group = 'AMD_Only'
  )
  
  amd_only_two_group_stats %<>%
    add_row(!!!pval_list)
}
```

## Assemble Stats Table
```{r}
all_two_group_stats =
  all_two_group_stats %>%
  add_row(!!!amd_only_two_group_stats)
```

## Write Stats Table
```{r}
filename = makeDataFileName(
  'mann_whitney_unadjusted.xlsx', 
  "${tables_dir}",
  "${analysis_type}",
  "${miseq_project_prefix}",
  "${clustering_level}"
  )

print(filename)

write.xlsx(
  all_two_group_stats,
  file=filename,
  quote=F
)
```

## Linear Regression
```{r}

```

## Logistic Regression
```{r}

```


## Write Table
```{r}
wb = createWorkbook()

all_ranks = c('Phylum', 'Class', 'Order', 'Family', 'Genus')

if (clustering_level == 'ASV')
{
  ranks_to_keep = c(all_ranks, 'ASVs')
} else
{
  ranks_to_keep = all_ranks[1:match(clustering_level, all_ranks)]
}

addWorksheet(wb, sheetName="Master Table")
openxlsx::writeData(wb, sheet="Master Table", all_master_table)

just_taxa = abundance %>% select(ranks_to_keep, glommed_taxa, short_glommed_taxa)
addWorksheet(wb, sheetName="${tools::toTitleCase(clustering_level)}")
openxlsx::writeData(wb, sheet="${tools::toTitleCase(clustering_level)}", just_taxa)

filename = makeDataFileName(
  'master_table.xlsx',
  "${tables_dir}",
  "${analysis_type}",
  "${miseq_project_prefix}",
  "${clustering_level}"
)
print(filename)

saveWorkbook(wb, file=filename, overwrite=T)
```


