---
title: "${miseq_project_prefix}: ${analysis_type}"
output:
  html_notebook:
    df_print: paged
    number_sections: yes
    theme: lumen
    toc: yes
    code_folding: hide
---

<style>
  html {
    font-size: 16pt;
  }
  
  body {
    font-size: 16pt;
  }
  
  h1 {
    font-size: 2.2rem;
  }
  
  h2 {
    font-size: 2rem;
  }
  
  h3 {
    font-size: 1.8rem;
  }
  
  h4 {
    font-size: 1.4rem;
  }
  
  .dataTable {
    font-size: 9pt;
  }
  
</style>

```{r}
source('${project_metadata_file}')
analysis_type = "${analysis_type}"
clustering_level = "${clustering_level}"
tables_dir = "${tables_dir}"
analysis_title = tools::toTitleCase(gsub("_", " ", analysis_type))
```

# ${miseq_project_prefix} DESeq2 Analysis {.tabset}

**Clustering**
Data clustered at the `r toupper('${clustering_level}')` Level.

**Subgroups**
```{r, results='asis', echo=FALSE}
for (sg_filter in c(${subgroup_filters}))
{
  cat(sg_filter, '\n\n')
}
```

**Filtering**
* **Relative Abundance cutoff:** ${relative_abundance_cutoff}
* **Prevalence Cutoff:** ${prevalence_cutoff}
* **Min Count Cutoff:** ${min_count_cutoff}

```{r}
relative_abundance_cutoff = ${relative_abundance_cutoff}
prevalence_cutoff = ${prevalence_cutoff}
min_count_cutoff = ${min_count_cutoff}
```

## Setup

### [Close]

### Start Conda ENV
```{r}
startCondaEnv('deseq', lib='~/R35')
```

### Load Libraries
```{r}
library(rmarkdown)
library(tidyverse)
library(magrittr)
library(ggplot2)
library(openxlsx)
library(DT)
library(kableExtra)
library(DESeq2)

### Custom libraries that can be loaded from GitHub
source('~/utilities/analysis-utilities/general_asv_data_analysis_utilities.R')
source('~/utilities/analysis-utilities/deseq_utilities.R')
source('~/utilities/amd_templates/setup/amd_project_utilities.R')
source('~/utilities/analysis-utilities/pathway_utilities.R')
```

### Set Knitr Options
```{r}
${knitr_options}
```

### Load Data
```{r}
ko_counts_table_file = file.path(
  "${tables_dir}",
  "ko_pathway_abundance_integers.xlsx"
)

ko_counts_table = read.xlsx(
  ko_counts_table_file,
  colNames=T
  )

sample_data = read.xlsx(
  '${sample_data_file}',
  colNames=T
  )

sample_data_columns = colnames(sample_data)
sample_names = sample_data$SampleName

source(file.path(
  "${project_root}", 
  "${metadata_dir}", 
  'observational_variables.R'
  ))

```

### Extract Subgroups
```{r}
if (!is.null(c(${subgroup_filters})))
{
  subgroup_filter_string = paste0(${subgroup_filters}, collapse=" & ")
  print(sprintf("Subgroup Filter String: %s", subgroup_filter_string))
  subgroup_filter_expression = str2expression(subgroup_filter_string)
  sample_data = sample_data %>% filter(eval(subgroup_filter_expression))
  sample_names = sample_data$SampleName

  asv_table = asv_table %>% select('ASVs', sample_names)
}

```

### Set Metadata Factor Levels
```{r}
sample_data = setFactorsLevels(sample_data, observational_variables)
```

### Variables of interest
#### Observational Variables
```{r}
test_groups = ${test_groups}
```

* **Test Groups:** `r test_groups`

## All Subgroups Case/Control Only
### Create Workbook
```{r}
deseq_results_wb = createWorkbook()
```

#### Run and add worksheets
```{r}
# results_df = runDeseqFromPathwayCounts(
#   pathway_counts=pathway_counts_table,
#   sample_data,
#   variables="CaseString",
#   include_covariates = 'All'
#   )

results_df = runDeseqFromPathwayCounts(
  pathway_counts=ko_counts_table,
  sample_data,
  variables="CaseString",
  include_covariates = 'All',
  pathway_colname='KO'
  )

padj_col = "CaseString_AMD_vs_Control.padj"

results_df =
  results_df %>%
  select(KO, !!padj_col, everything()) %>%
  arrange(!!as.name(padj_col)) %>%
  mutate(PathwayName = getKEGGPathways(KO)) %>%
  select(KO, PathwayName, CaseString_AMD_vs_Control.padj, everything())
```

```{r}
final_results = 
  inner_join(results_df, df , by='Pathway') %>% 
  select(Pathway, PathwayName, CaseString_AMD_vs_Control.padj, everything())

filename = file.path('~', 'miseq-239_piphillin_deseq_case_control.xlsx')
print(filename)
write.xlsx(final_results, file=filename)
write.xlsx(final_results, 'test.xlsx')

```

```{r}
addWorksheet(
  wb=deseq_results_wb, 
  sheetName=""
  )

openxlsx::writeData(
  wb=deseq_results_wb, 
  sheet="", 
  x=results_df
  )
  
setColWidths(
  wb=deseq_results_wb, 
  sheet="", 
  cols=1, 
  widths = "auto"
)

freezePane(
  deseq_results_wb, 
  sheet="", 
  firstCol=T,
  firstRow=T
)
```

#### Save Workbook
```{r}
filename = makeDataFileName(
  'all_subgroups_case_vs_control.xlsx',
  "${tables_dir}",
  "${analysis_type}",
  "${miseq_project_prefix}",
  "rank"
)

print(filename)

saveWorkbook(deseq_results_wb, file=filename, overwrite=T)
```

## All Subgroups -- All Variables
### Create Workbook
```{r}
deseq_results_wb = createWorkbook()
```

### Run and add worksheets
```{r}
results_df = runDeseqFromPathwayCounts(
  pathway_counts=pathway_counts_table,
  sample_data,
  variables=test_groups,
  include_covariates = 'All'
  )
  
padj_col = "CaseString_AMD_vs_Control.padj"
  
results_df =
  results_df %>%
  select(pathway, !!padj_col, starts_with("CaseString"), everything()) %>%
  arrange(!!as.name(padj_col))

addWorksheet(
  wb=deseq_results_wb, 
  sheetName=""
  )

openxlsx::writeData(
  wb=deseq_results_wb, 
  sheet="", 
  x=results_df
  )
  
setColWidths(
  wb=deseq_results_wb, 
  sheet="", 
  cols=1, 
  widths = "auto"
)

freezePane(
  deseq_results_wb, 
  sheet="", 
  firstCol=T,
  firstRow=T
)
```

### Save Workbook
```{r}
filename = makeDataFileName(
  'all_subgroups_with_covariates.xlsx',
  "${tables_dir}",
  "${analysis_type}",
  "${miseq_project_prefix}",
  "rank"
)

print(filename)

saveWorkbook(deseq_results_wb, file=filename, overwrite=T)
```


## Age < 90 Case/Control Only
### Create Workbook
```{r}
deseq_results_wb = createWorkbook()
```

#### Run and add worksheets
```{r}
results_df = runDeseqFromPathwayCounts(
  counts_table=pathway_counts_table,
  sample_data %>% filter(Age < 90),
  variables="CaseString",
  include_covariates = 'All'
  )

padj_col = "CaseString_AMD_vs_Control.padj"

results_df =
  results_df %>%
  select(pathway, !!padj_col, everything()) %>%
  arrange(!!as.name(padj_col))

addWorksheet(
  wb=deseq_results_wb, 
  sheetName=""
  )

openxlsx::writeData(
  wb=deseq_results_wb, 
  sheet="", 
  x=results_df
  )
  
setColWidths(
  wb=deseq_results_wb, 
  sheet="", 
  cols=1, 
  widths = "auto"
)

freezePane(
  deseq_results_wb, 
  sheet="", 
  firstCol=T,
  firstRow=T
)
```

#### Save Workbook
```{r}
filename = makeDataFileName(
  'all_subgroups_case_vs_control.xlsx',
  "${tables_dir}",
  "${analysis_type}",
  "${miseq_project_prefix}",
  "rank"
)

print(filename)

saveWorkbook(deseq_results_wb, file=filename, overwrite=T)
```

## Age < 90 -- All Variables
### Create Workbook
```{r}
deseq_results_wb = createWorkbook()
```

### Run and add worksheets
```{r}
results_df = runDeseqFromPathwayCounts(
  counts_table=pathway_counts_table,
  sample_data %>% filter(Age < 90),
  variables=test_groups,
  include_covariates = 'All'
  )
  
padj_col = "CaseString_AMD_vs_Control.padj"
  
results_df =
  results_df %>%
  select(pathway, !!padj_col, starts_with("CaseString"), everything()) %>%
  arrange(!!as.name(padj_col))

addWorksheet(
  wb=deseq_results_wb, 
  sheetName=""
  )

openxlsx::writeData(
  wb=deseq_results_wb, 
  sheet="", 
  x=results_df
  )
  
setColWidths(
  wb=deseq_results_wb, 
  sheet="", 
  cols=1, 
  widths = "auto"
)

freezePane(
  deseq_results_wb, 
  sheet="", 
  firstCol=T,
  firstRow=T
)
```

### Save Workbook
```{r}
filename = makeDataFileName(
  'all_subgroups_with_covariates.xlsx',
  "${tables_dir}",
  "${analysis_type}",
  "${miseq_project_prefix}",
  "rank"
)

print(filename)

saveWorkbook(deseq_results_wb, file=filename, overwrite=T)
```


```{r}
kos = results_df %>% filter(CaseString_AMD_vs_Control.padj<0.1) %>% pull(Pathway)

df = data.frame(Pathway=character(), NAME=character())

for (ko in kos)
{
  print(ko)
  
  response = keggGet(ko) %>%
    .[[1]] %>%
    .[c('NAME')] %>%
    lapply(as.character) 
  response$Pathway = ko
  df %<>% add_row(!!!response)
}

df %<>% rename(PathwayName=NAME)


```

```{r}
final_results = 
  inner_join(results_df, df , by='Pathway') %>% 
  select(Pathway, PathwayName, CaseString_AMD_vs_Control.padj, everything())

filename = file.path('~', 'miseq-239_piphillin_deseq_case_control.xlsx')
print(filename)
write.xlsx(final_results, file=filename)
write.xlsx(final_results, 'test.xlsx')

```

