---
title: "${miseq_project_prefix}: ${analysis_type}"
output:
  html_notebook:
    df_print: paged
    number_sections: yes
    theme: lumen
    toc: yes
    code_folding: hide
---

<style>
  html {
    font-size: 16pt;
  }
  
  body {
    font-size: 16pt;
  }
  
  h1 {
    font-size: 2.2rem;
  }
  
  h2 {
    font-size: 2rem;
  }
  
  h3 {
    font-size: 1.8rem;
  }
  
  h4 {
    font-size: 1.4rem;
  }
  
</style>

```{r}
source('${project_metadata_file}')
analysis_type = "${analysis_type}"
clustering_level = "${clustering_level}"
tables_dir = "${tables_dir}"
```

# ${miseq_project_prefix} Prevotella Analysis {.tabset}
**Clustering**
Data clustered at the `r toupper('${clustering_level}')` Level.

**Subgroups**
```{r, results='asis', echo=FALSE}
for (sg_filter in c(${subgroup_filters}))
{
  cat(sg_filter, '\n\n')
}
```

**Filtering**
  * **Relative Abundance cutoff:** ${relative_abundance_cutoff}
* **Prevalence Cutoff:** ${prevalence_cutoff}
* **Min Count Cutoff:** ${min_count_cutoff}
```{r}
relative_abundance_cutoff = ${relative_abundance_cutoff}
prevalence_cutoff = ${prevalence_cutoff}
min_count_cutoff = ${min_count_cutoff}
```

## Setup

### [Close]

### Start Conda ENV
```{r}
startCondaEnv('regression')
```

### Load Libraries
```{r}
library(rmarkdown)
library(tidyverse)
library(magrittr)
library(ggplot2)
library(ggbeeswarm)
library(openxlsx)
library(DT)
library(pheatmap)
library(kableExtra)

### Custom libraries that can be loaded from GitHub
source('~/utilities/analysis-utilities/general_asv_data_analysis_utilities.R')
source('~/utilities/amd_templates/setup/amd_project_utilities.R')
```


### Set Knitr Options
```{r}
${knitr_options}
```

### Load Data
```{r}
asv_table = 
  read.delim(
    '${asv_table_file}', 
    header=T, 
    sep='\t',
    stringsAsFactors=F
    )

taxonomy_table = 
  read.delim(
    '${taxonomy_table_file}', 
    header=T, 
    sep='\t',
    stringsAsFactors=F
    ) %>%
  mutate_all(as.character())

sample_data = 
  read.delim(
    '${sample_data_file}', 
    header=T, 
    sep='\t',
    stringsAsFactors=F
    ) %>%
  filter(SampleID!='Undetermined')


if ("IGA" %in% colnames(sample_data) & use_allbac)
{
  sample_data = 
    sample_data %>%
    filter(IGA=='AllBac')

  sample_names = sample_data$SampleName
  
  asv_table = 
    asv_table %>%
    select('ASVs', ends_with('AllBac'))
}

sample_data_columns = colnames(sample_data)
sample_names = sample_data$SampleName

```

### Extract Subgroups
```{r}
if (!is.null(c(${subgroup_filters})))
{
  subgroup_filter_string = paste0(${subgroup_filters}, collapse=" & ")
  print(sprintf("Subgroup Filter String: %s", subgroup_filter_string))
  subgroup_filter_expression = str2expression(subgroup_filter_string)
  sample_data = sample_data %>% filter(eval(subgroup_filter_expression))
  sample_names = sample_data$SampleName

  asv_table = asv_table %>% select('ASVs', sample_names)
}

```


### Set Filtering
```{r}
relative_abundance_cutoff = ${relative_abundance_cutoff}
prevalence_cutoff = ${prevalence_cutoff}
```
**relative abundance cutoff:** `r relative_abundance_cutoff`
**prevalence cutoff:** `r prevalence_cutoff`


### Get Filtered Taxa Abundance
```{r}
taxa_abundance = getFilteredTaxaCountsDev(
  asv_table,
  taxonomy_table,
  sample_data, 
  cluster_by="${clustering_level}",
  relative_abundance_cutoff=relative_abundance_cutoff,
  prevalence_cutoff=prevalence_cutoff,
  id_col="SampleName",
  normalize=T
  )

filtered_taxa = taxa_counts$short_glommed_taxa

```


## Prevotella Analysis

### Build Master Table with Stats
```{r}
master_table = getMasterTable(
  taxa_abundance %>% filter(Genus == 'Prevotella'),
  sample_metadata,
  'SampleName',
  'short_glommed_taxa'
  ) %>%
  mutate(prevotella_gt_5pct = ifelse(pct_prevotella > 0.05, 'Y', 'N'))
  
master_table %>%
  select(prevotella, prevotella_gt_5pct) %>%
  kable() %>%
  kable_styling(fixed_thead = T, font_size=10) %>%
  scroll_box(width = "800px", height = "400px")
```

### Cross Tab
```{r}
cross_tab = master_table %>% select(CaseString, prevotella_gt_5pct) %>% table()
print(cross_tab)

cross_tab_df = 
  master_table %>% 
  select(CaseString, prevotella_gt_5pct) %>% 
  group_by(CaseString, prevotella_gt_5pct) %>% 
  summarise(n=n()) %>% 
  spread(prevotella_gt_5pct, n) %>%
  data.frame() %>%
  mutate(RowSum = select(., Y, N) %>% rowSums()) %>%
  rename(Prev_lt_5pct=N, Prev_gt_5pct=Y) %>%
  add_row(
    CaseString='ColSum', 
    Prev_lt_5pct=sum(.$Prev_lt_5pct), 
    Prev_gt_5pct=sum(.$Prev_gt_5pct), 
    RowSum=sum(.$RowSum))

kable(cross_tab_df, caption='Count Table') %>%
  kable_styling(font_size=12) %>%
  column_spec(2:3, bold=T, color='#000000') %>%
  column_spec(c(1), bold=F, color='#aaaaaa') %>%
  row_spec(1:2, bold = T, color='#000000') %>%
  row_spec(3, bold=F, color='#aaaaaa') %>%
  column_spec(c(4), bold=F, color='#aaaaaa')

```

### Percentages
```{r}
total_counts = 134

pcts_df = 
  cross_tab_df %>% 
  mutate_if(is.numeric, list(~./total_counts))

kable(pcts_df, caption='Percentage Table') %>%
  kable_styling(font_size=12) %>%
  column_spec(2:3, bold=T, color='#000000') %>%
  column_spec(c(1), bold=F, color='#aaaaaa') %>%
  row_spec(1:2, bold = T, color='#000000') %>%
  row_spec(3, bold=F, color='#aaaaaa') %>%
  column_spec(c(4), bold=F, color='#aaaaaa')

```


### $\chi^2$ Test

```{r}
cs_test = chisq.test(cross_tab)

print(cs_test)
```
**$\chi^2$ P Value = `r cs_test$p.value`**

### Welch's t
```{r}
pair_test = t.test(
  pct_prevotella ~ CaseString,
  data=master_table,
  two.sided=T
)
print(pair_test)
```

### Mann-Whitney U
```{r}
pair_test = wilcox.test(
  pct_prevotella ~ CaseString,
  data=master_table,
  two.sided=T
)
print(pair_test)
```

### Pair Plot
```{r}
pval_annotation = data.frame(
  TestVariable=c('CaseString'),
  index=c('pct_prevotella'),
  xloc=c(1.5),
  yloc=1.3*max(master_table$pct_prevotella),
  pvals=c(paste0('p = ', round(pair_test$p.value, 3)))
)

plotPairTest(
  master_table,
  indices=c('pct_prevotella'),
  var_data=observational_variables$CaseString,
  aesthetics = list(x_var='CaseString', color_var='CaseString', facet_var='index'),
  title_template='Prevotella Pct. AMD vs Control',
  annotation_data=pval_annotation
) +
  ylab("Pct. Prevotella")
```

### OLS Linear Regression
```{r}
linfit = lm(pct_prevotella ~ CaseString + Age + Gender, data=master_table)
print(summary(linfit))
```

### Rank OLS Regression
```{r}
rank_linfit = lm(rank(pct_prevotella) ~ CaseString + Age + Gender, data=master_table)
print(summary(rank_linfit))
```
```


