---
title: "MiSeq"
output:
  html_notebook:
    df_print: paged
    number_sections: yes
    theme: lumen
    toc: yes
    code_folding: hide
---

<style>
  html {
    font-size: 16pt;
  }
  
  body {
    font-size: 16pt;
  }
  
  h1 {
    font-size: 2.2rem;
  }
  
  h2 {
    font-size: 2rem;
  }
  
  h3 {
    font-size: 1.8rem;
    
  }
  
  h4 {
    font-size: 1.4rem;
  }
  
</style>


```{r}
source('${project_metadata_file}')
analysis_type = "${analysis_type}"
clustering_level = "${clustering_level}"
tables_dir = "${tables_dir}"
```

# MiSeq-${miseq_run_number} Alpha Diversity Two Group Plots {.tabset}
Mann-Whitney Two Group Testing

## Clustering Level
Data clustered at the `r toupper(clustering_level)` Level.

## Setup

### Start Conda ENV
```{r}
startCondaEnv('${conda_env}')
```

### Set Knitr Options
```{r}
library(rmarkdown)
library(knitr)

${knitr_options}

```

### Load Libraries
```{r}
library(tidyverse)
library(magrittr)
library(DT)
library(openxlsx)
library(ggplot2)
library(ggbeeswarm)

### Custom libraries that can be loaded from GitHub
source('~/utilities/analysis-utilities/general_asv_data_analysis_utilities.R')
source('~/utilities/analysis-utilities/alpha_div_utilities.R')
source('~/utilities/analysis-utilities/project_utilities.R')
```

### Load Data
```{r}
filename = makeDataFileName('master_tables.xlsx', tables_dir, analysis_type)

all_master_table = readWorkbook(filename, sheet='All Subgroups')
age_lt_90_master_table = readWorkbook(filename, sheet='Age < 90')
amd_only_master_table = readWorkbook(filename, sheet='AMD Only')

alpha_diversity = 
  all_master_table %>%
  select(SampleName, Observed, Shannon, Simpson)

filename = makeDataFileName('mann_whitney_unadjusted.xlsx', tables_dir, analysis_type)
two_group_stats = read.xlsx(filename)

source(file.path(metadata_dir, 'observational_variables.R'))
```

### Variables of interest
#### All Observational Variables
```{r}
# print(observational_variables)
```

#### AMD Only Observational Variables
```{r}
amd_only_variables = observational_variables

amd_only_excluded_variables = c(
  'Gender',
  'CaseString',
  'Age'
)

for (varname in amd_only_excluded_variables)
{
  print(sprintf('Excluding %s', varname))
  amd_only_variables[varname] = NULL
}

```

**Excluding: ** `r amd_only_excluded_variables`


#### Calculated Experimental Variables

```{r}
calculated_exp_vars = c('Observed', 'Shannon', 'Simpson')
```
**Calculated Experimental Variables: ** `r calculated_exp_vars`


### Set Metadata Factor Levels
```{r}
all_master_table = setFactorsLevels(
  all_master_table, 
  observational_variables
  )
age_lt_90_master_table = setFactorsLevels(
  age_lt_90_master_table, 
  observational_variables
  )
amd_only_master_table = setFactorsLevels(
  amd_only_master_table, 
  observational_variables
  )
```

### P-Value Annotationas
```{r}
pval_annotations = getPvalAnnotations(
  variables=c(calculated_exp_vars),
  master_table=all_master_table,
  pval_height_factor=1.2,
  two_group_stats=two_group_stats
)
```


## Two Group Test Plots

### All Subgroups: AMD vs. Control
```{r}
makeBoxAndDotplot(
  master_table=all_master_table,
  indices=calculated_exp_vars,
  var_data=observational_variables$CaseString,
  aesthetics=list(
    x_var='CaseString', 
    color_var='CaseString', 
    facet_var='index'
    ),
  annotation_data=NULL,
  title="Alpha Diversity All Subgroups"
)
```

### Age < 90
```{r}
makeBoxAndDotplot(
  master_table=age_lt_90_master_table,
  indices=calculated_exp_vars,
  var_data=observational_variables$CaseString,
  aesthetics=list(
    x_var='CaseString', 
    color_var='CaseString', 
    facet_var='index'
    ),
  annotation_data=NULL,
  title="Alpha Diversity Age<90"
)
```

### AMD Only---Multiple Covariates

```{r}
plotAllTwoGroupTests(
  variables=amd_only_variables,
  master_table=amd_only_master_table,
  indices=calculated_exp_vars,
  title_template="Alpha Diversity AMD Only: ",
  annotation_data=pval_annotations %>% filter(TestGroup=='AMD_Only')
)

```
