---
title: "MiSeq"
output:
  html_notebook:
    df_print: paged
    number_sections: yes
    theme: lumen
    toc: yes
    code_folding: hide
---

<style>
  html {
    font-size: 16pt;
  }
  
  body {
    font-size: 16pt;
  }
  
  h1 {
    font-size: 2.2rem;
  }
  
  h2 {
    font-size: 2rem;
  }
  
  h3 {
    font-size: 1.8rem;
  }
  
  h4 {
    font-size: 1.4rem;
  }
  
</style>

```{r}
source('${project_metadata_file}')
analysis_type = "${analysis_type}"
clustering_level = "${clustering_level}"
tables_dir = "${tables_dir}"
```

# MiSeq-${miseq_run_number} Alpha Diversity Analysis Data Prep{.tabset}
## Clustering Level
Data clustered at the `r toupper(clustering_level)` Level.

## Setup 

### [Close]

### Start Conda ENV
```{r}
startCondaEnv('${conda_env}')
```

### Set Knitr Options
```{r}
library(rmarkdown)
library(knitr)

${knitr_options}
```

### Load Libraries
```{r}
library(tidyverse)
library(magrittr)
library(DT)
library(openxlsx)

### Custom libraries that can be loaded from GitHub
source('~/utilities/analysis-utilities/general_asv_data_analysis_utilities.R')
source('~/utilities/analysis-utilities/alpha_div_utilities.R')
source('~/utilities/analysis-utilities/project_utilities.R')
```

### Load Data
```{r}
asv_table = 
  read.delim(
    '${asv_table_file}', 
    header=T, 
    sep='\t',
    stringsAsFactors=F
    )

taxonomy_table = 
  read.delim(
    '${taxonomy_table_file}', 
    header=T, 
    sep='\t',
    stringsAsFactors=F
    ) %>%
  mutate_all(as.character())

sample_data = 
  read.delim(
    '${sample_data_file}', 
    header=T, 
    sep='\t',
    stringsAsFactors=F
    ) %>%
  filter(SampleID!='Undetermined')

sample_data_columns = colnames(sample_data) %>% setdiff('SampleName')
sample_names = sample_data$SampleName

asv_table_cols = 
  data.frame(
    SampleID=colnames(asv_table)
  ) %>%
  inner_join(sample_data, by='SampleID') %>%
  pull(SampleName)

colnames(asv_table) = c('ASVs', asv_table_cols)

```

### Set Filtering
```{r}
relative_abundance_cutoff = ${relative_abundance_cutoff}
prevalence_cutoff = ${prevalence_cutoff}
```
**relative abundance cutoff:** `r relative_abundance_cutoff`
**prevalence cutoff:** `r prevalence_cutoff`

### Get Filtered Taxa Abundance
```{r}
genus_taxa_abundance = getFilteredTaxaCounts(
  asv_table,
  taxonomy_table,
  sample_data, 
  lowest_rank=tools::toTitleCase(clustering_level),
  relative_abundance_cutoff=relative_abundance_cutoff,
  prevalence_cutoff=prevalence_cutoff,
  id_col="SampleName",
  normalize=T
  )
```

### Create Master Table

```{r}
master_table = getMasterTable(
  genus_taxa_abundance,
  sample_data,
  'SampleName',
  'short_glommed_taxa'
)
```

## Calculate Diversity Indices {.tabset}

### [Close]  

### All
```{r}
all_alpha_diversity = calculateAlphaDiversity2(
  master_table,
  genus_taxa_abundance$short_glommed_taxa,
  c('observed', 'shannon', 'simpson')
)

amd_only_alpha_diversity = calculateAlphaDiversity2(
  master_table %>% filter(CaseString=='AMD'),
  genus_taxa_abundance$short_glommed_taxa,
  c('observed', 'shannon', 'simpson')
)

age_lt_90_alpha_diversity = calculateAlphaDiversity2(
  master_table %>% filter(Age<90),
  genus_taxa_abundance$short_glommed_taxa,
  c('observed', 'shannon', 'simpson')
)

all_master_table =
  inner_join(all_alpha_diversity, master_table, by='SampleName') %>%
  select(SampleName, Observed=observed, Shannon=shannon, Simpson=simpson, sample_data_columns, everything())

amd_only_master_table =
  inner_join(amd_only_alpha_diversity, master_table, by='SampleName') %>%
  select(SampleName, Observed=observed, Shannon=shannon, Simpson=simpson, sample_data_columns, everything())


age_lt_90_master_table =
  inner_join(age_lt_90_alpha_diversity, master_table, by='SampleName') %>%
  select(SampleName, Observed=observed, Shannon=shannon, Simpson=simpson, sample_data_columns, everything())
```

### Write Alpha Diversity Tables
```{r}
alpha_div_indices = createWorkbook()

addWorksheet(alpha_div_indices, sheetName="All Subgroups")
openxlsx::writeData(alpha_div_indices, sheet="All Subgroups", all_master_table)
addWorksheet(alpha_div_indices, sheetName="Age < 90")
openxlsx::writeData(alpha_div_indices, sheet="Age < 90", age_lt_90_master_table)
addWorksheet(alpha_div_indices, sheetName="AMD Only")
openxlsx::writeData(alpha_div_indices, sheet="AMD Only", amd_only_master_table)

filename = makeDataFileName(
  'master_tables.xlsx',
  tables_dir,
  analysis_type
)

saveWorkbook(alpha_div_indices, file=filename, overwrite=T)
```
